---
title: "Alevin QC Plots"
description: description
author:
  - "Modified by - Aaron M. Allen"
date: 'Last update: `r date()`'
output:
  html_document:
    number_sections: true
    code_folding: show
    code_download: true
    theme: cerulean
    df_print: paged
    fig_width: 8.5
    fig_height: 5
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 6
---




<style>
   tbody tr:nth-child(odd){
    background-color: #D6E0F5;
  }
</style>



```{css, echo = FALSE}
pre, code {white-space:pre !important; overflow-x:auto}
.tocify-item {white-space:pre}
```




```{r echo = FALSE, cache = FALSE}

```



```{r echo = FALSE, cache = FALSE}
ggplot2::theme_set(ggplot2::theme_grey())
ggplot2::theme_update(panel.background = ggplot2::element_rect(fill = "transparent", colour = NA),
                      plot.background = ggplot2::element_rect(fill = "transparent", colour = NA))

rmd_name <- gsub(pattern = ".Rmd", replacement = "", x = knitr::current_input(), ignore.case = TRUE)
knitr::opts_chunk$set(dev = c("png", "cairo_pdf"),
                      dev.args=list(bg="transparent"),
                      fig.align = "center",
                      fig.height = 5,
                      fig.width = 8.5,
                      dpi = 300,
                      pdf.options(encoding = "ISOLatin9.enc"),
                      fig.path = paste0("../analyses/figures/", rmd_name, "-",
                                        format(Sys.time(), "%Y%m%d_%H%M%S"), "/"),
                      fig.retina = 1,
                      warning = TRUE,
                      message = TRUE)
```


```{r echo = FALSE, cache = FALSE}
options(scipen = 9999)
```




# Setup

## Job_ID = `r commandArgs(trailingOnly = TRUE)[2]`


## Start time

```{r start-time}
set.seed(12021)
today <- format(Sys.time(), "%Y%m%d_%H%M%S")
total_start_time <- Sys.time()
total_start_time
```





## Parameters


```{r params}
params_file <- commandArgs(trailingOnly = TRUE)[3]
if (is.na(params_file)) {
    params_file <- "pipeline_R_params.R"
}
params_path <- paste0("../src/param_files/", params_file)
print(paste0("The params file is :   ", params_path))
```

```{r, class.source = 'fold-hide', code = readLines(params_path)}
```

Get the name of the machine that this is running on.  If running on my local desktop ("mentok"), then append to the `raw_path` variable.

```{r is-local}
local_path_mod <- ""
curr_machine <- Sys.info()["nodename"]
if (curr_machine == "mentok") {
    local_path_mod <- "/mnt/data2/scRNAseq/cbrg"
}
local_path_mod
```

```{r adjust-path}
raw_path <- paste0(local_path_mod, raw_path)
raw_path
```






## Libraries

```{r packages-etc, warning = FALSE, message = FALSE}
library(tidyverse)
library(alevinQC)
library(cowplot)
library(colorspace)
```


## Functions

### Timer function

```{r run-time-tibble}
process_time <- tibble()
stop_the_clock <- function(chunk_name = "unnamed") {
    temp <- tibble::tibble(Code_chunk = chunk_name,
                           Start_time = start_time,
                           Stop_time = Sys.time(),
                           Run_time = difftime(Stop_time, Start_time, units = "mins")
                           )
    process_time <- dplyr::bind_rows(process_time, temp)
    return(process_time)
}
```





### Summary table functions

```{r}
collate_summary <- function(alevinqc_list, dataset) {

    summary_table <- data.frame()
    for (i in seq_along(alevinqc_list)) {
        temp_summary <- alevinqc_list[[i]]$summaryTables[[dataset]] %>%
            as.data.frame() %>%
            rownames_to_column("Features") %>%
            rename(!!sample_names[[i]] := "V1")
        if (i == 1) {
            summary_table <- temp_summary
        } else {
            summary_table <- left_join(x = summary_table, y = temp_summary, by = c("Features"))
        }
    }
    summary_table <- summary_table %>%
        gather("Sample", "Value", -Features) %>%
        mutate(Value = as.numeric(str_remove(string = Value, pattern = "%")))
    return(summary_table)

}
```


```{r}
plot_summary <- function(summary_table, title) {

    my_variables <- unique(summary_table$Features)
    for (i in seq_along(my_variables)) {
        p1 <- summary_table %>%
            filter(Features == my_variables[[i]]) %>%
            spread("Features", "Value") %>%
            rename(my_variable = my_variables[[i]]) %>%
            ggplot(aes(x = factor(Sample, level = unlist(sample_names)),
                       y = my_variable,
                       color = my_variable,
                       fill = my_variable)) +
                geom_bar(stat = "identity") +
                ylab(my_variables[[i]]) +
                ggtitle(title) +
                theme(axis.text.x = element_text(angle = 45, hjust = 1),
                      legend.position = "none",
                      plot.title = element_text(face = "bold"),
                      axis.title.x = element_blank())
        if (grepl(pattern = "Percent mapped", x = my_variables[[i]])) {
            p1 <- p1 + ylim(0, 100)
        }
        plot(p1)
    }

}
```



### CB table functions

```{r}
collate_cb_table <- function(alevinqc_list, umi_thresh) {

    summary_table <- list()
    for (i in seq_along(alevinqc_list)) {

        summary_table[[i]] <- alevin_list[[i]]$cbTable %>%
            filter(totalUMICount >= umi_thresh) %>%
            select(-CB, -ranking, -inFirstWhiteList, -inFinalWhiteList) %>%
            summarise_all(list(mean, median), na.rm = TRUE) %>%
            gather("Feature", "Value") %>%
            mutate(Avg_Type = if_else(grepl(pattern = "_fn1", x = Feature), "mean", "median"),
                   Feature = str_remove(string = Feature, pattern = "_fn[:digit:]"),
                   Sample = sample_names[[i]])
    }
    summary_table <- bind_rows(summary_table)
    return(summary_table)

}
```



```{r}
plot_cb_summary <- function(summary_table, avg_type, umi_thresh) {

    my_variables <- unique(cb_table_summary$Feature)
    for (i in seq_along(my_variables)) {
        p1 <- cb_table_summary %>%
            filter(Feature == my_variables[[i]]) %>%
            filter(Avg_Type == avg_type) %>%
            ggplot(aes(x = factor(Sample, level = unlist(sample_names)),
                       y = Value,
                       color = Value,
                       fill = Value)) +
                geom_bar(stat = "identity") +
                ylab(my_variables[[i]]) +
                ggtitle(paste0(avg_type, ",  ", "UMI >= ", umi_thresh)) +
                theme(axis.text.x = element_text(angle = 45, hjust = 1),
                      legend.position = "none",
                      plot.title = element_text(face = "bold"),
                      axis.title.x = element_blank())
        if (grepl(pattern = "Rate", x = my_variables[[i]])) {
            p1 <- p1 + ylim(0, 1)
        }
        plot(p1)
    }

}
```


### Adjusted reads in cells function

```{r}
reads_in_cells_adj <- function(alevinqc_list, umi_thresh) {

    in_cells_table <- list()
    for (i in seq_along(alevinqc_list)) {
        in_cells_table[[i]] <- tibble(Sample = sample_names[[i]],
            total_reads = sum(alevinqc_list[[i]]$cbTable$originalFreq, na.rm = TRUE),
            read_in_cells = sum(alevinqc_list[[i]]$cbTable$originalFreq[alevinqc_list[[i]]$cbTable$totalUMICount >= umi_thresh],
                                na.rm = TRUE),
            percent_reads_in_cells = 100 * read_in_cells / total_reads
            )
    }
    in_cells_table <- bind_rows(in_cells_table)
    return(in_cells_table)

}
```




# Load the data



```{r load-data}
start_time <- Sys.time()
print("Loading Alevin data")
alevin_list <- list()
for (i in seq_along(raw_names)) {
    print(paste0("  Reading ", raw_names[[i]], " 10x data"))
    alevin_list[[i]] <- readAlevinQC(baseDir = paste0(raw_path, raw_names[[i]]))
}
Sys.time()-start_time
process_time <- stop_the_clock("load_data")
```

<!-- ```{r load-data} -->
<!-- start_time <- Sys.time() -->
<!-- print("Loading Alevin data") -->
<!-- alevin_list2 <- list() -->
<!-- alevin_list2 <- lapply(X = raw_names, function(x) readAlevinQC(baseDir = paste0(raw_path, x))) -->
<!-- Sys.time()-start_time -->
<!-- #process_time <- stop_the_clock("load_data") -->
<!-- ``` -->






# Summary table plots


## Full set of cell barcodes

```{r}
full_dataset_summary <- collate_summary(alevinqc_list = alevin_list, dataset = "fullDataset")
full_dataset_summary
```

```{r}
write_csv(x = spread(data = full_dataset_summary, key = Sample, value = Value), file = "../../proj068/analyses/alevin/fru_iso/full_dataset_summary.csv")
```


```{r}
full_dataset_summary %>% 
    spread(Sample, Value) %>% 
    knitr::kable()
```




```{r fig.width=4, fig.height=6}
plot_summary(summary_table = full_dataset_summary, title = "Full set of cell barcodes")
```


## Initial whitelist


```{r}
init_whitelist_summary <- collate_summary(alevinqc_list = alevin_list, dataset = "initialWhitelist")
init_whitelist_summary
```


```{r}
write_csv(x = spread(data = init_whitelist_summary, key = Sample, value = Value), file = "../../proj068/analyses/alevin/fru_iso/init_whitelist_summary.csv")
```


```{r}
init_whitelist_summary %>% 
    spread(Sample, Value) %>% 
    knitr::kable()
```



```{r fig.width=4, fig.height=6}
plot_summary(summary_table = init_whitelist_summary, title = "Initial whitelist")
```


## Final whitelist

```{r}
final_whitelist_summary <- collate_summary(alevinqc_list = alevin_list, dataset = "finalWhitelist")
final_whitelist_summary
```



```{r}
write_csv(x = spread(data = final_whitelist_summary, key = Sample, value = Value), file = "../../proj068/analyses/alevin/fru_iso/final_whitelist_summary.csv")
```

```{r}
final_whitelist_summary %>% 
    spread(Sample, Value) %>% 
    knitr::kable()
```



```{r fig.width=4, fig.height=6}
plot_summary(summary_table = final_whitelist_summary, title = "Final whitelist")
```








# Average feature plots


## UMI >= 0

```{r}
cb_table_summary <- collate_cb_table(alevinqc_list = alevin_list, umi_thresh = 0)
cb_table_summary
```


```{r}
write_csv(x = spread(data = cb_table_summary, key = Sample, value = Value), file = "../../proj068/analyses/alevin/fru_iso/cb_table_summary_0UMI.csv")
```



### Mean

```{r fig.width=4, fig.height=6}
plot_cb_summary(summary_table = cb_table_summary, avg_type = "mean", umi_thresh = 0)
```


```{r}
cb_table_summary %>% 
    filter(Avg_Type == "mean") %>% 
    spread(Sample, Value) %>% 
    knitr::kable()
```

### Median

```{r fig.width=4, fig.height=6}
plot_cb_summary(summary_table = cb_table_summary, avg_type = "median", umi_thresh = 0)
```




```{r}
cb_table_summary %>% 
    filter(Avg_Type == "median") %>% 
    spread(Sample, Value) %>% 
    knitr::kable()
```


## UMI >= 200

```{r}
cb_table_summary <- collate_cb_table(alevinqc_list = alevin_list, umi_thresh = 200)
cb_table_summary
```

```{r}
write_csv(x = spread(data = cb_table_summary, key = Sample, value = Value), file = "../../proj068/analyses/alevin/fru_iso/cb_table_summary_200UMI.csv")
```

### Mean

```{r fig.width=4, fig.height=6}
plot_cb_summary(summary_table = cb_table_summary, avg_type = "mean", umi_thresh = 200)
```


```{r}
cb_table_summary %>% 
    filter(Avg_Type == "mean") %>% 
    spread(Sample, Value) %>% 
    knitr::kable()
```

### Median

```{r fig.width=4, fig.height=6}
plot_cb_summary(summary_table = cb_table_summary, avg_type = "median", umi_thresh = 200)
```


```{r}
cb_table_summary %>% 
    filter(Avg_Type == "median") %>% 
    spread(Sample, Value) %>% 
    knitr::kable()
```


# Adjusted Reads in Cells


```{r}
reads_in_cells_table <- reads_in_cells_adj(alevinqc_list = alevin_list, umi_thresh = 200)
reads_in_cells_table
```

```{r}
reads_in_cells_table %>%
    gather("Features", "Value", -Sample) %>% 
    spread(Sample, Value)
```



```{r}
write_csv(x = reads_in_cells_table %>%
                gather("Features", "Value", -Sample) %>% 
                spread(Sample, Value), 
    file = "../../proj068/analyses/alevin/fru_iso/reads_in_cells_table.csv")
```



```{r}
reads_in_cells_table %>% 
    knitr::kable()
```

```{r fig.width=4, fig.height=6}
reads_in_cells_table %>%
    ggplot(aes(x = factor(Sample, level = unlist(sample_names)),
               y = percent_reads_in_cells,
               color = percent_reads_in_cells,
               fill = percent_reads_in_cells)) +
        geom_bar(stat = "identity") +
        ylab("Percent Reads in Cells (%)") +
        ggtitle(paste0("Read in Cells with UMI >= ", 200)) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              legend.position = "none",
              plot.title = element_text(face = "bold"),
              axis.title.x = element_blank())
```












# Run time

```{r run-times}
process_time
```

```{r plot-run-times, fig.width = 10, fig.height = 5}
process_time$Code_chunk <- factor(process_time$Code_chunk, levels = rev(process_time$Code_chunk))
process_time %>%
    ggplot(aes(x = Code_chunk, y = Run_time, fill = Run_time)) +
    geom_col(width = 0.8) +
    xlab("Code Chunk") +
    ylab("Run Time (mins)") +
    coord_flip()
```



```{r stop-time}
total_end_time <- Sys.time()
total_end_time
total_end_time - total_start_time
```



# CPU and RAM Usage

```{r fig.width = 12, fig.height = 4}
job_id <- commandArgs(trailingOnly = TRUE)[2]
if (is.na(job_id)) {
    job_id <- "interactive"
}
if (job_id != "local" && job_id != "interactive") {
    job_log <- paste0("../sps-", job_id, "/sps-", job_id, "-cpu.tsv")
    if (file.exists(job_log)) {
        p1 <- read_tsv(file = job_log) %>%
          gather("Process", "Usage", -`#TIME`, -REQUESTED) %>%
          group_by(`#TIME`) %>%
          ggplot(aes(x = `#TIME`, y = Usage, colour = Process)) +
              geom_line() +
              ggtitle("CPU") +
              theme(legend.position = "none", validate = TRUE)
        p2 <- read_tsv(file = paste0("../sps-", job_id, "/sps-", job_id, "-mem.tsv")) %>%
          gather("Process", "Usage", -`#TIME`, -REQUESTED) %>%
          group_by(`#TIME`) %>%
          ggplot(aes(x = `#TIME`, y = Usage, colour = Process)) +
              geom_line() +
              ggtitle("Memory")
        my_legend <- get_legend(p2)
        p2 <- p2 + theme(legend.position = "none", validate = TRUE)
        plot_grid(p1, p2, my_legend, ncol = 3, rel_widths = c(3, 3, 2))
    }
}
```


# Session info

```{r session-info}
sessionInfo()
```
