---
title: "Initial Filtering and QC of scRNA-Seq data"
description: description
author: "Aaron M. Allen"
date: 'Last update: `r date()`'
output:
  html_document:
    number_sections: true
    code_folding: show
    code_download: true
    theme: cerulean
    df_print: paged
    fig_width: 8.5
    fig_height: 5
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 6
---


```{css, echo = FALSE}
pre, code {white-space:pre !important; overflow-x:auto}
.tocify-item {white-space:pre}
```




```{r echo = FALSE, cache = FALSE}

```


```{r setup, echo = FALSE, cache = FALSE}
ggplot2::theme_set(ggplot2::theme_grey())

rmd_name <- gsub(pattern = ".Rmd", replacement = "", x = knitr::current_input(), ignore.case = TRUE)
knitr::opts_chunk$set(dev = c("png", "cairo_pdf"),
                      fig.align = "center",
                      fig.height = 5,
                      fig.width = 8.5,
                      dpi = 300,
                      pdf.options(encoding = "ISOLatin9.enc"),
                      fig.path = paste0("../analyses/figures/", rmd_name, "-",
                                        format(Sys.time(), "%Y%m%d_%H%M%S"), "/"),
                      fig.retina = 1,
                      warning = TRUE,
                      message = TRUE)
```



# Setup

## Job_ID = `r commandArgs(trailingOnly = TRUE)[2]`


## Start time

```{r start-time}
set.seed(12021)
today <- format(Sys.time(), "%Y%m%d_%H%M%S")
total_start_time <- Sys.time()
total_start_time
```




## Parameters


```{r params}
params_file <- commandArgs(trailingOnly = TRUE)[3]
params_path <- paste0("../src/param_files/", params_file)
print(paste0("The params file is :   ", params_path))
```

```{r, class.source = 'fold-hide', code = readLines(params_path)}
```

Get the name of the machine that this is running on.  If running on my local desktop ("mentok"), then append to the `raw_path` variable.

```{r is-local}
local_path_mod <- ""
curr_machine <- Sys.info()["nodename"]
if (curr_machine == "mentok") {
    local_path_mod <- "/mnt/data2/scRNAseq/cbrg"
}
local_path_mod
```

```{r adjust-path}
raw_path <- paste0(local_path_mod, raw_path)
raw_path
```

```{r run-number}
run_number_1a <- commandArgs(trailingOnly = TRUE)[4]
run_number_1a
```






## Libraries

```{r packages-etc, warning = FALSE, message = FALSE}
library(utils)
library(Seurat)
library(tidyverse)
library(cowplot)
library(colorspace)
library(ggcorrplot)
library(ggpubr)
library(corrr)
```




```{r run-time-tibble}
process_time <- tibble()
stop_the_clock <- function(chunk_name = "unnamed") {
    temp <- tibble::tibble(Code_chunk = chunk_name,
                           Start_time = start_time,
                           Stop_time = Sys.time(),
                           Run_time = difftime(Stop_time, Start_time, units = "mins")
                           )
    process_time <- dplyr::bind_rows(process_time, temp)
    return(process_time)
}
```


```{r make-dir}
dir.create(objects_path, recursive = TRUE, showWarnings = FALSE)
```



# The data

## Load data

```{r load-data}
start_time <- Sys.time()

if (run_number_1a == 1) {
    if (input_type == "cellranger") {
        print("Loading Cellranger data")
        path_modifier <- "/outs/filtered_bc_matrix/"
    }
    if (input_type == "alevin") {
        raw_list <- list()
        print("Loading Alevin data")
        path_modifier <- "/alevin/filtered_cellranger_output/"
        # path_modifier <- "/alevin/raw_cellranger_output/"
    }
    raw_list <- list()
    for (i in raw_names) {
        print(paste0("  Reading ", i, " 10x data"))
        raw_list[[i]] <- Read10X(data.dir = paste0(raw_path, i, path_modifier))
    }
}

if (run_number_1a == 2) {
    if (ambient_method == "soupx") {
        print("Loading SoupX adjusted counts data")
        raw_list <- read_rds(file = paste0(soupx_output_path,
                                           dataset, "_",
                                           input_type, "_",
                                           "soupx_adj_counts.rds")
                             )
    }
    if (ambient_method == "decontx") {
        print("Loading DecontX adjusted counts data")
        raw_list <- read_rds(file = paste0(decontx_output_path,
                                           dataset, "_",
                                           input_type, "_",
                                           "decontx_seurat_list_adj.rds")
                             )
    }
}


process_time <- stop_the_clock("load_data")
```



## Round the data

```{r round-data}
start_time <- Sys.time()
if (round_the_data) {
    round_to_int <- function(input) {
        input@x <- floor(input@x) + rbinom(length(input@x), 1, input@x - floor(input@x))
        return(input)
    }
    for (i in seq_along(raw_list)) {
        raw_list[[i]] <- round_to_int(raw_list[[i]])
    }
}
process_time <- stop_the_clock("round_data")
```



## Create Seurat objects

```{r create-seurat}
start_time <- Sys.time()
seurat_list <- list()
for (i in seq_along(raw_list)) {
    print(paste0("Creating Seurat object of ", sample_names[[i]]))
    seurat_list[[i]] <- CreateSeuratObject(counts = raw_list[[i]],
                                           min.cells = min_cells,
                                           min.features = min_features,
                                           project = sample_names[[i]])
    if (run_number_1a == 1) {
        seurat_list[[i]] <- RenameCells(object = seurat_list[[i]],
                                        add.cell.id = sample_names[[i]])
    }
    seurat_list[[i]]$orig.ident <- sample_names[[i]]
    names(seurat_list)[[i]] <- sample_names[[i]]
}
seurat_list
process_time <- stop_the_clock("create_seurat")
```


# Initial QC

## Replicate correlation

```{r counts-tibble}
if (length(seurat_list) > 1) {
    start_time <- Sys.time()
    for (i in seq_along(seurat_list)) {
        temp_counts <- tibble(gene = seurat_list[[i]]@assays[["RNA"]]@counts@Dimnames[[1]],
                              !!seurat_list[[i]]@project.name := rowSums(as.array(seurat_list[[i]]@assays[["RNA"]]@counts))
                              )
        if (i == 1) {
            raw_counts_df <- temp_counts
        } else {
            raw_counts_df <- full_join(raw_counts_df, temp_counts, by = "gene")
        }
    }
    raw_counts_df <- raw_counts_df %>% replace(is.na(.), 0)
    process_time <- stop_the_clock("counts_tibble")
}
```

```{r plot-correlation, fig.height = 6,fig.width = 6, warning = FALSE, message = FALSE}
if (length(seurat_list) > 1) {
    start_time <- Sys.time()
    combos <- combn(seq_len(length(seurat_list)), 2)
    p <- list()
    for (i in 1:dim(combos)[2]) {
        p[[i]] <- raw_counts_df %>%
              ggplot(aes(x = log1p(!!(as.name(colnames(raw_counts_df)[combos[, i][1] + 1]))),
                         y = log1p(!!(as.name(colnames(raw_counts_df)[combos[, i][2] + 1]))))) +
                geom_point() +
                coord_fixed() +
                geom_smooth(method = lm) +
                stat_cor(label.y = 15, size = 5) +
                stat_regline_equation(label.y = 14, size = 5) +
                ylim(0, 15) +
                xlim(0, 15) +
                theme_gray() +
                theme(text = element_text(size = 14)) +
                ggtitle(paste0(colnames(raw_counts_df)[combos[, i][1] + 1],
                               " vs ",
                               colnames(raw_counts_df)[combos[, i][2] + 1]
                               )
                        ) +
                theme(plot.title = element_text(face = "bold", hjust = 0.5))
        plot(p[[i]])
    }
    process_time <- stop_the_clock("plot_correlation")
}
```


## Plot percent feature class

```{r percent-feature}
start_time <- Sys.time()
for (i in seq_along(seurat_list)) {
    seurat_list[[i]][["percent_mt"]] <- PercentageFeatureSet(seurat_list[[i]], pattern = "^mt:")
    seurat_list[[i]][["percent_rRNA"]] <- PercentageFeatureSet(seurat_list[[i]], pattern = "(S|-)rRNA")
    seurat_list[[i]][["percent_rProt"]] <- PercentageFeatureSet(seurat_list[[i]], pattern = "Rp(L|S)")
    seurat_list[[i]][["percent_hsp"]] <- PercentageFeatureSet(seurat_list[[i]], pattern = "^Hsp")
}
process_time <- stop_the_clock("percent_feature")
```


```{r feature-plot-function}
qc_params_plot_function <- function(object_list,  # nolint
                                    plot_qc_param = "nCount_RNA",
                                    plot_title = NULL,
                                    plot_low_thres = 0,
                                    plot_high_thres = Inf,
                                    plot_title_mod = NULL) {

    print(paste0("Plotting values for: ", plot_qc_param))
    qc_data <- list()
    for (ii in seq_along(object_list)) {
        qc_data[[ii]] <- Seurat::FetchData(object = object_list[[ii]], vars = c("orig.ident", plot_qc_param)) %>%
            dplyr::mutate()
    }
    p <- qc_data %>%
            dplyr::bind_rows() %>%
            dplyr::rename(sample = orig.ident) %>%
            dplyr::rename(metric = plot_qc_param) %>%
            ggplot2::ggplot(ggplot2::aes(x = sample, y = metric)) +
                ggplot2::labs(x = "Samples", y = plot_qc_param) +
                ggplot2::geom_jitter(size = .01, alpha = 0.5, aes(colour = sample)) +
                ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1), legend.position = "none") +
                ggplot2::geom_boxplot(outlier.shape = NA, alpha = 0.5) +
                ggplot2::geom_hline(yintercept = c(plot_low_thres, plot_high_thres), linetype = "dashed") +
                ggplot2::ggtitle(paste0(plot_title_mod, "  -  ", plot_title)) +
                ggplot2::theme(plot.title = ggplot2::element_text(face = "bold"))
    if (plot_qc_param %in% c("nFeature_RNA", "nCount_RNA")) {
        p <- p +
            ggplot2::scale_y_continuous(trans = "log10") +
            ggplot2::annotation_logticks(sides = "l")
    }
    print(p)

}
```



```{r percent-feature-plot-raw}
start_time <- Sys.time()
options(scipen = 9999)
low_thres <- list(min_genes, min_umi, 0, 0, 0, 0)
high_thres <- list(max_genes, max_umi, max_mito, max_hsp, max_rprot, max_rrna)
titles <- list("Number of features",
                 "Number of UMIs",
                 "Percentage of mitochondrial RNA",
                  "Percentage of heat-shock protein mRNA",
                 "Percentage of ribosomal proteins",
                 "Percentage of ribosomal RNA"
                 )

for (i in seq_along(qc_params_plot)) {
    qc_params_plot_function(seurat_list,
                            plot_qc_param = qc_params_plot[[i]],
                            plot_title = titles[[i]],
                            plot_low_thres = low_thres[[i]],
                            plot_high_thres = high_thres[[i]],
                            plot_title_mod = "Raw data"
                            )
}
process_time <- stop_the_clock("percent_feature_plot_raw")
```






## Remove PolIII features

10x Chromium v3 seems to get a lot of PolIII features, which are not useful for downstream analysis. We will remove these features from the Seurat object.

```{r remove-poliii}
if (remove_poliii_genes) {
    start_time <- Sys.time()

    raw_list_backup <- raw_list
    seurat_list_backup <- seurat_list
    seurat_list <- list()

    for (i in seq_along(raw_list)) {
        print(paste0("Processing   ", sample_names[[i]], "  ..."))

        print("    Removing rRNA and tRNA genes  ...")
        r_rna_genes <- rownames(raw_list[[i]]) %>% str_subset(pattern = "rRNA")
        t_rna_genes <- rownames(raw_list[[i]]) %>% str_subset(pattern = "tRNA:")
        raw_list[[i]] <- raw_list[[i]][-(which(rownames(raw_list[[i]]) %in% c(r_rna_genes, t_rna_genes))),]

        print("    Creating Seurat object ...")
        seurat_list[[i]] <- CreateSeuratObject(counts = raw_list[[i]],
                                            min.cells = min_cells,
                                            min.features = min_features,
                                            project = sample_names[[i]])
        if (run_number_1a == 1) {
            seurat_list[[i]] <- RenameCells(object = seurat_list[[i]],
                                            add.cell.id = sample_names[[i]])
        }
        seurat_list[[i]]$orig.ident <- sample_names[[i]]
        names(seurat_list)[[i]] <- sample_names[[i]]

        print("    Adding metadata  ...")
        seurat_list[[i]][["percent_mt"]] <- PercentageFeatureSet(seurat_list[[i]], pattern = "^mt:")
        seurat_list[[i]][["percent_rRNA"]] <- PercentageFeatureSet(seurat_list[[i]], pattern = "(S|-)rRNA")
        seurat_list[[i]][["percent_rProt"]] <- PercentageFeatureSet(seurat_list[[i]], pattern = "Rp(L|S)")
        seurat_list[[i]][["percent_hsp"]] <- PercentageFeatureSet(seurat_list[[i]], pattern = "^Hsp")

    }
    seurat_list
    process_time <- stop_the_clock("remove_poliii")
}
```



### Re-Plot percent feature class - no PolIII genes

```{r percent-feature-plot-poliii}
if (remove_poliii_genes) {
    start_time <- Sys.time()
    for (i in seq_along(qc_params_plot)) {
        qc_params_plot_function(seurat_list,
                                plot_qc_param = qc_params_plot[[i]],
                                plot_title = titles[[i]],
                                plot_low_thres = low_thres[[i]],
                                plot_high_thres = high_thres[[i]],
                                plot_title_mod = "Removed PolIII genes"
                                )
    }
    process_time <- stop_the_clock("percent_feature_plot_poliii")
}
```



## Trim objects


### Removing outlier cells

```{r trim-function}
trim_data <- function(object) {
  data_trim <- object %>%
      subset(subset = nFeature_RNA >= min_genes &
                 nFeature_RNA <= max_genes &
                 nCount_RNA >= min_umi &
                 nCount_RNA <= max_umi &
                 percent_mt < max_mito &
                 percent_rRNA < max_rrna &
                 percent_rProt <= max_rprot &
                 percent_hsp <= max_hsp)
  removed <- dim(object)[2] - dim(data_trim)[2]
  perc_removed <- removed / dim(object)[2] * 100
  print(paste0("Number of cells removed in ",
                object@project.name, ": ",
                removed, " (",
                round(perc_removed, 2), "% of ",
                dim(object)[2], ")"))
  return(data_trim)
}
```

```{r trim-object}
start_time <- Sys.time()
seurat_list_trim <- list()
for (i in seq_along(seurat_list)) {
    seurat_list_trim[[i]] <- seurat_list[[i]] %>% trim_data
    seurat_list_trim[[i]]$orig.ident <- sample_names[[i]]
    names(seurat_list_trim)[[i]] <- sample_names[[i]]
}
seurat_list_trim
process_time <- stop_the_clock("trim_object")
```


### Remove Small Reps

If the number of cells is less than `min_tot_cells`, then the replicate will be removed from later analyses.

```{r remove-small-reps}
objs_to_remove <- c()
for (i in seq_along(seurat_list_trim)) {
    num_cells <- dim(seurat_list_trim[[i]])[2]
    if (num_cells < min_tot_cells) {
        message(paste0("Sample ", names(seurat_list_trim)[[i]], " has ", num_cells, " cells, and will be removed."))
        objs_to_remove <- c(objs_to_remove, names(seurat_list_trim)[[i]])
    }
}
for (i in seq_along(objs_to_remove)) {
    seurat_list_trim[[objs_to_remove[[i]]]] <- NULL
}
seurat_list_trim
```


### Re-Plot percent feature class - trimmed

```{r percent-feature-plot-trim}
start_time <- Sys.time()
for (i in seq_along(qc_params_plot)) {
    qc_params_plot_function(seurat_list_trim,
                            plot_qc_param = qc_params_plot[[i]],
                            plot_title = titles[[i]],
                            plot_low_thres = low_thres[[i]],
                            plot_high_thres = high_thres[[i]],
                            plot_title_mod = "Trimmed data"
                            )
}
process_time <- stop_the_clock("percent_feature_plot_trim")
```



## Add Tissue Metadata

```{r tissue-meta-data}
if (multi_tissue) {
    for (i in seq_along(seurat_list_trim)) {
        tissue_table <- seurat_list_trim[[i]]@meta.data %>%
            select(orig.ident) %>%
            rownames_to_column("cell_id") %>%
            mutate(tissue = str_extract(string = orig.ident, pattern = "[^_]+"))
        seurat_list_trim[[i]] <- AddMetaData(object = seurat_list_trim[[i]],
            metadata = tissue_table$tissue,
            col.name = "tissue")
    }
}
```

## Add Sex Metadata

```{r sex-meta-data}
if (insilico_sex) {
    for (i in seq_along(seurat_list_trim)) {
        seurat_list_trim[[i]] <- NormalizeData(object = seurat_list_trim[[i]])
        seurat_list_trim[[i]] <- AddModuleScore(object = seurat_list_trim[[i]],
            features = list(c("lncRNA:roX1", "lncRNA:roX2")),
            name = "modulescore_sex")
        sex_table <- seurat_list_trim[[i]]@meta.data %>%
            rownames_to_column("cell_id") %>%
            select(cell_id, modulescore_sex1) %>%
            mutate(sex = "unknown") %>%
            mutate(sex = if_else(modulescore_sex1 < 0, "female", "male"))
        seurat_list_trim[[i]] <- AddMetaData(object = seurat_list_trim[[i]],
            metadata = sex_table$sex,
            col.name = "sex")
        p1 <- data.frame(ModuleScore = seurat_list_trim[[i]]@meta.data[["modulescore_sex1"]]) %>%
            ggplot(aes(x = ModuleScore)) +
                geom_histogram(aes(y = 100 * ..count.. / sum(..count..)),
                             fill = "lightsteelblue2",
                             position = "stack",
                             bins = 60,
                             show.legend = FALSE) +
                geom_vline(xintercept = -0.4) +
                ylab("Percent") +
                theme(plot.margin = unit(c(30, 30, 30, 30), "pt")) +
                ggtitle(names(seurat_list_trim)[[i]])
        plot(p1)
    }
}
if (sample_sex) {
    for (i in seq_along(seurat_list_trim)) {
        sex_table <- seurat_list_trim[[i]]@meta.data %>%
            select(orig.ident) %>%
            rownames_to_column("cell_id") %>%
            mutate(sex = if_else(grepl("fem", orig.ident), "female", "male"))
        seurat_list_trim[[i]] <- AddMetaData(object = seurat_list_trim[[i]],
            metadata = sex_table$sex,
            col.name = "sex")
    }
}
```







## Save trimmed Seurat objects

```{r save-trimmed}
start_time <- Sys.time()
if (run_number_1a == 1) {
    write_rds(x = seurat_list,
           file = paste0(objects_path,
                         dataset, "_",
                         input_type, "_",
                         "seurat_list.rds"))
    write_rds(x = seurat_list_trim,
           file = paste0(objects_path,
                         dataset, "_",
                         input_type, "_",
                         "trimmed_seurat_list.rds")
            )
}
if (run_number_1a == 2) {
    write_rds(x = seurat_list,
           file = paste0(objects_path,
                         dataset, "_",
                         input_type, "_",
                         "doublets_removed_ambient_removed_seurat_list.rds"))
    write_rds(x = seurat_list_trim,
           file = paste0(objects_path,
                         dataset, "_",
                         input_type, "_",
                         "doublets_removed_ambient_removed_trimmed_seurat_list.rds")
            )
}
process_time <- stop_the_clock("save_trimmed")
```







# Run time

```{r run-times}
process_time
```

```{r plot-run-times, fig.width = 10, fig.height = 5}
process_time$Code_chunk <- factor(process_time$Code_chunk, levels = rev(process_time$Code_chunk))
process_time %>%
    ggplot(aes(x = Code_chunk, y = Run_time, fill = Run_time)) +
    geom_col(width = 0.8) +
    xlab("Code Chunk") +
    ylab("Run Time (mins)") +
    coord_flip()
```



```{r stop-time}
total_end_time <- Sys.time()
total_end_time
total_end_time - total_start_time
```



# CPU and RAM Usage

```{r fig.width = 12, fig.height = 4}
job_id <- commandArgs(trailingOnly = TRUE)[2]
if (job_id != "local") {
    job_log <- paste0("../sps-", job_id, "/sps-", job_id, "-cpu.tsv")
    if (file.exists(job_log)) {
        p1 <- read_tsv(file = job_log) %>%
          gather("Process", "Usage", -`#TIME`, -REQUESTED) %>%
          group_by(`#TIME`) %>%
          ggplot(aes(x = `#TIME`, y = Usage, colour = Process)) +
              geom_line() +
              ggtitle("CPU") +
              theme(legend.position = "none", validate = TRUE)
        p2 <- read_tsv(file = paste0("../sps-", job_id, "/sps-", job_id, "-mem.tsv")) %>%
          gather("Process", "Usage", -`#TIME`, -REQUESTED) %>%
          group_by(`#TIME`) %>%
          ggplot(aes(x = `#TIME`, y = Usage, colour = Process)) +
              geom_line() +
              ggtitle("Memory")
        my_legend <- get_legend(p2)
        p2 <- p2 + theme(legend.position = "none", validate = TRUE)
        plot_grid(p1, p2, my_legend, ncol = 3, rel_widths = c(3, 3, 2))
    }
}
```


# Session info

```{r session-info}
sessionInfo()
```
