---
title: "Doublet method comparison"
description: Comparing the outputs of DoubletFinder, Scrublet, and Solo
author: "Aaron M. Allen"
date: 'Last update: `r date()`'
output:
  html_document:
    number_sections: true
    code_folding: show
    code_download: true
    theme: cerulean
    df_print: paged
    fig_width: 8.5
    fig_height: 5
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 6
---



```{css, echo = FALSE}
pre, code {white-space:pre !important; overflow-x:auto}
.tocify-item {white-space:pre}
```



```{r echo = FALSE, cache = FALSE}

```


```{r setup, echo = FALSE, cache = FALSE}
ggplot2::theme_set(ggplot2::theme_grey())

rmd_name <- gsub(pattern = ".Rmd", replacement = "", x = knitr::current_input(), ignore.case = TRUE)
knitr::opts_chunk$set(dev = c("png", "cairo_pdf"),
                      fig.align = "center",
                      fig.height = 5,
                      fig.width = 8.5,
                      dpi = 300,
                      pdf.options(encoding = "ISOLatin9.enc"),
                      fig.path = paste0("../analyses/figures/", rmd_name, "-",
                                        format(Sys.time(), "%Y%m%d_%H%M%S"), "/"),
                      fig.retina = 1,
                      warning = TRUE,
                      message = TRUE)
```



# Setup

## Job_ID = `r commandArgs(trailingOnly = TRUE)[2]`


## Start time

```{r start-time}
set.seed(12021)
today <- format(Sys.time(), "%Y%m%d_%H%M%S")
total_start_time <- Sys.time()
total_start_time
```




```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(Seurat)
library(data.table)
library(Matrix)
library(eulerr)
library(ComplexHeatmap)
library(colorspace)
library(cowplot)
```



# Params



```{r params}
params_file <- commandArgs(trailingOnly = TRUE)[3]
params_path <- paste0("../src/param_files/", params_file)
print(paste0("The params file is :   ", params_path))
```

```{r, class.source = 'fold-hide', code = readLines(params_path)}
```





# Intro



To remove doublets from the Sexed brain data set I've run 3 different methods/packages to identify doublets: [DoubletFinder](https://github.com/chris-mcginnis-ucsf/DoubletFinder), [Scrublet](https://github.com/swolock/scrublet), and [Solo](https://github.com/calico/solo). I have previously made other notebooks where I run though a bunch of different parameters for each method, and then another set of notebooks where I summarize and compare the effects within each method for varying their parameters. In this notebook I'm comparing between method after taking the *"best(?)"* results from each method.

I had previously also tried [DoubletDetection](https://github.com/JonathanShor/DoubletDetection). It seemed to work well for the brain data, but in general was very sensitive to varying UMI depth. In the VNC we have a few cell types with very different UMI depth, so it was not behaving properly.

I had also briefly tried [DoubletDecon](https://github.com/EDePasquale/DoubletDecon), but after running for 3 weeks on the cluster, it still hadn't finished. So I moved on.



# Load the data


### Load the Seurat object.

I'm first loading the previously computed Seurat object.

```{r}
seurat_list_trim <- read_rds(file = paste0(objects_path,
                                          dataset, "_",
                                          input_type, "_",
                                          "trimmed_preprocessed_seurat_list.rds")
                         )
```


### Cell types and Markers

```{r load-markers}
marker_df <- read_tsv(gmt_file_path, col_names = FALSE) %>%
    select(-X2) %>%
    gather("index", "gene", -X1) %>%
    select(-index) %>%
    rename(celltype = X1) %>%
    drop_na() %>%
    group_by(celltype) %>%
    filter(celltype %in% qc_plot_cell_type) %>%
    slice(1) %>%
    arrange(celltype)
marker_df
```


Then filter
marker gene to make sure they are in the object.


```{r genes-present}
all_genes <- c()
for (i in seq_along(seurat_list_trim)) {
    all_genes <- c(all_genes, rownames(seurat_list_trim[[i]]))
}
all_genes <- unique(all_genes)
genes_not_found <- marker_df$gene[!marker_df$gene %in% all_genes]
if (length(genes_not_found) > 0) {
    warning(paste0("The following genes were not found: ", genes_not_found))
    marker_df <- marker_df %>% filter(gene != genes_not_found)
} else {
    message(paste0("All genes present in data."))
}
```





### Load DoubletFinder data.


Doublets are called as *"high"* vs. *"low"* classification to estimate the  *"heterotypic/neotypic"* vs *"homotypic/embedded"* doublets. I collapse the *"high"* vs. *"low"* doublet classification. I'm not terribly worried about a little "over-classification" as I'll be intersecting with other methods anyway.

```{r warning = F, message = F}
df_doublets <- read_csv(paste0(doublet_path,
                                 "doubletfinder/",
                                 dataset, "_",
                                 input_type, "_",
                                 "doubletfinder_calls.csv"),
                  col_names = TRUE
                  ) %>%
    mutate(DoubletFinder = if_else(DoubletFinder == "Singlet", "Singlet", "Doublet"))
head(df_doublets)
```



### Load Scrublet PC test.

The Scrublet code still allows for someone to specify a range of PCs to test. So for the sake of this script, the end user has to pick one that they want to use. I assume it is the same as the default PC, `curr_pc`, used in the rest of the pipeline.


```{r}
input_files <- Sys.glob(paste0(doublet_path, "scrublet/*_predicted_doublets.csv"))
scrub_doublets <- list()
for (i in seq_along(input_files)) {
    curr_sample_name <- input_files[[i]] %>% str_remove(paste0(doublet_path, "scrublet/")) %>% str_remove("_predicted_doublets.csv")
    scrub_doublets[[i]] <- read_csv(input_files[[i]]) %>%
        select("0", as.character(curr_pc)) %>%
        rename(c(cell_id = `0`, Scrublet = as.character(curr_pc))) %>%
        mutate(cell_id = str_c(curr_sample_name, cell_id, sep = "_") %>% str_remove("-1"),
               Scrublet = if_else(Scrublet == 1, "Doublet", "Singlet")) %>%
        select(cell_id, Scrublet)
}
scrub_doublets <- bind_rows(scrub_doublets)
head(scrub_doublets)
```





### Load Solo doublets


<!--  very old bit
Here we read in the `pred.npy` which has the raw predictions output from `solo`. On their [GitHub issue #62](https://github.com/calico/solo/issues/62), they say that the `preds.npy` are raw and need to be recalibrated, which is what the `is_doublet.npy` file is. But in our case the `is_doublet.npy` reports many fewer doublets than we would expect, and the `preds.npy` numbers are in an acceptable range. For now we will load the `preds.npy` instead of the `is_doublet.npy` until we sort this out.
 -->


<!-- Old bit of code for loading in the output from the stand alone cli solo. But I've now switched over to using solo inside scvi-tools -->
<!-- Here we read in the `is_doublet.npy` which has the raw predictions output from `solo`. -->
<!-- ```{r} -->
<!-- input_files <- Sys.glob(paste0(doublet_path,"solo/*/is_doublet.npy")) -->
<!-- solo_doublets <- list() -->
<!-- if (length(input_files) > 0) { -->
<!--     for (i in seq_along(input_files)) { -->
<!--         curr_sample_name <- input_files[[i]] %>% str_remove(paste0(doublet_path,"solo/")) %>% str_remove("/is_doublet.npy") -->
<!--         cell_ids <- read_tsv(file = paste0(doublet_path,"filtered_inputs/",curr_sample_name,"/barcodes.tsv"),col_names = FALSE) -->
<!--         doublets <- np$load(file = paste0(doublet_path,"solo/",curr_sample_name,"/is_doublet.npy")) -->
<!--         # sometimes the solo doublets file is one longer than the data ... don't know why ... -->
<!--         doublets <- doublets[1:length(cell_ids$X1)] -->
<!--         solo_doublets[[i]] <- tibble(cell_id = str_c(curr_sample_name, cell_ids$X1, sep = "_") %>% str_remove("-1"), -->
<!--                                      Solo = doublets) %>% -->
<!--             mutate(Solo = if_else(Solo == TRUE,"Doublet","Singlet")) %>% -->
<!--             select(cell_id, Solo) -->
<!--     } -->
<!--     solo_doublets <- bind_rows(solo_doublets) -->
<!-- } else { -->
<!--     solo_doublets <- data.frame(cell_id = NA, -->
<!--                                 Solo = NA) -->
<!-- } -->
<!-- head(solo_doublets) -->
<!-- ``` -->



Read in the `*_solo_doublet_predictions.csv` file.

```{r}
# input_files <- Sys.glob(paste0(doublet_path,"solo/*_solo_doublet_scores.csv"))
input_files <- Sys.glob(paste0(doublet_path, "solo/*_solo_doublet_predictions.csv"))
input_files
solo_doublets <- list()
for (i in seq_along(input_files)) {
    # curr_sample_name <- input_files[[i]] %>% str_remove(paste0(doublet_path,"solo/")) %>% str_remove("_solo_doublet_scores.csv")
    curr_sample_name <- input_files[[i]] %>% str_remove(paste0(doublet_path, "solo/")) %>% str_remove("_solo_doublet_predictions.csv")
    solo_doublets[[i]] <- read_csv(input_files[[i]]) %>%
        rename(cell_id = X1) %>%
        mutate(cell_id = str_c(curr_sample_name, cell_id, sep = "_") %>% str_remove("-1-0"),
               # Solo = if_else(doublet > 0 & singlet < 0,"Doublet","Singlet")) %>%
               Solo = if_else(prediction == "doublet", "Doublet", "Singlet")) %>%
        select(cell_id, Solo)
}
solo_doublets <- bind_rows(solo_doublets)
head(solo_doublets)
```


# Compare the methods


### Join the data

I `left_join` the data sets together.


```{r}
all_cells <- c()
for (i in seq_along(seurat_list_trim)) {
  all_cells <- c(all_cells, colnames(seurat_list_trim[[i]]))
}
```

```{r}
all_doublets <- tibble(cell_id = all_cells) %>%
    left_join(df_doublets, by = "cell_id") %>%
    left_join(scrub_doublets, by = "cell_id") %>%
    left_join(solo_doublets, by = "cell_id") %>%
    mutate(sample_rep = str_sub(cell_id, end = -18))
all_doublets[is.na(all_doublets)] <- "Singlet"
head(all_doublets)
```





### Count the doublets

Of the n cells, each method identifies a different number of doublets. This is to be expected as `Scrublet` is only supposed to identify *"heterotypic/neotypic"* cells. Since I've included the homotypic adjustment (`nExp = nExp_poi.adj`, `reuse.pANN = "pANN_0.25_X.XX_XXXX"`) for `DoubletFinder`, it should return *"homotypic/embedded"* cells as well. And `Solo` is doing its own thing...


```{r}
all_doublets %>%
    summarise(total_cells = length(cell_id),
              DoubletFinder = length(cell_id[DoubletFinder == "Doublet"]),
              Scrublet = length(cell_id[Scrublet == "Doublet"]),
              Solo = length(cell_id[Solo == "Doublet"])
              ) %>%
    gather("Method", "Number of cells")
```


### Euler plot

```{r}
fit <- euler(c("DoubletFinder" = length(all_doublets$cell_id[all_doublets$DoubletFinder == "Doublet" &
                                                          all_doublets$Scrublet == "Singlet" &
                                                          all_doublets$Solo == "Singlet"]),
               "Scrublet" = length(all_doublets$cell_id[all_doublets$DoubletFinder == "Singlet" &
                                                           all_doublets$Scrublet == "Doublet" &
                                                          all_doublets$Solo == "Singlet"]),
               "Solo" = length(all_doublets$cell_id[all_doublets$DoubletFinder == "Singlet" &
                                                           all_doublets$Scrublet == "Singlet" &
                                                          all_doublets$Solo == "Doublet"]),

               "DoubletFinder&Scrublet" = length(all_doublets$cell_id[all_doublets$DoubletFinder == "Doublet" &
                                                            all_doublets$Scrublet == "Doublet" &
                                                            all_doublets$Solo == "Singlet"]),
               "Scrublet&Solo" = length(all_doublets$cell_id[all_doublets$DoubletFinder == "Singlet" &
                                                            all_doublets$Scrublet == "Doublet" &
                                                            all_doublets$Solo == "Doublet"]),
               "DoubletFinder&Solo" = length(all_doublets$cell_id[all_doublets$DoubletFinder == "Doublet" &
                                                            all_doublets$Scrublet == "Singlet" &
                                                            all_doublets$Solo == "Doublet"]),

               "DoubletFinder&Scrublet&Solo" = length(all_doublets$cell_id[all_doublets$DoubletFinder == "Doublet" &
                                                            all_doublets$Scrublet == "Doublet" &
                                                            all_doublets$Solo == "Doublet"])
               ))

plot(fit,
     quantities = TRUE,
     fills = list(fill = colorspace::qualitative_hcl(n = 3, palette = "Set 2"), alpha = 0.5)
)
```


### Upset plots

Upset plots, if anyone prefers...

```{r fig.height = 3, fig.width = 5, eval = F}
all_doublets_bolean <- all_doublets
all_doublets_bolean[all_doublets_bolean == "Singlet"] <- "0"
all_doublets_bolean[all_doublets_bolean == "Doublet"] <- "1"
all_doublets_bolean[, 2:4] <- sapply(all_doublets_bolean[, 2:4], as.numeric)
all_doublets_bolean <- all_doublets_bolean[-rowSums(all_doublets_bolean[, 2:4]) != 0, ]
m1 <- make_comb_mat(all_doublets_bolean, mode = "distinct")
m2 <- make_comb_mat(all_doublets_bolean, mode = "intersect")
m3 <- make_comb_mat(all_doublets_bolean, mode = "union")
UpSet(m1, column_title = "distinct mode")
UpSet(m2, column_title = "intersect mode")
UpSet(m3, column_title = "union mode")
```


### Union vs intersection

Below are the numbers of identified doublets depending on whether you take the union (`any_1`), the strict intersection (`all_3`), or in-between (`any_2`).

```{r}
all_doublets %>%
    summarise(any_1 = sum(rowSums(.[, 2:4] != "Singlet") > 0),
              any_2 = sum(rowSums(.[, 2:4] != "Singlet") > 1),
              all_3 = sum(rowSums(.[, 2:4] != "Singlet") > 2)
              )
```




# Doublets in the UMAP


### Add doublet metadata

Adding meta data to the initial integrated and clustered object.

```{r}
for (i in seq_along(seurat_list_trim)) {
  seurat_list_trim[[i]] <- AddMetaData(seurat_list_trim[[i]],
                            metadata = all_doublets$DoubletFinder[all_doublets$sample_rep == unique(seurat_list_trim[[i]]$orig.ident)],
                            col.name = "DoubletFinder")
  seurat_list_trim[[i]] <- AddMetaData(seurat_list_trim[[i]],
                            metadata = all_doublets$Scrublet[all_doublets$sample_rep == unique(seurat_list_trim[[i]]$orig.ident)],
                            col.name = "Scrublet")
  seurat_list_trim[[i]] <- AddMetaData(seurat_list_trim[[i]],
                            metadata = all_doublets$Solo[all_doublets$sample_rep == unique(seurat_list_trim[[i]]$orig.ident)],
                            col.name = "Solo")
}
```



### Plot markers on UMAP

To get the lay of the land, we'll first plot out some markers.

```{r fig.width = 10, fig.height = 14}
p_list <- list()
for (ii in seq_along(seurat_list_trim)) {
    pt_size <- 4 * AutoPointSize(data = seurat_list_trim[[ii]], raster = FALSE)
    for (i in seq_along(marker_df$gene)) {
        p_list[[i]] <- FeaturePlot(object = seurat_list_trim[[ii]],
                              features = marker_df$gene[[i]],
                              pt.size = min(4, pt_size),
                              cols = c("lightsteelblue2", "black"),
                              order = TRUE) +
                      NoAxes() +
                      coord_fixed() +
                      ggtitle(paste0(marker_df$celltype[[i]], "\n(", marker_df$gene[[i]], ")")) +
                      theme(plot.title = element_text(size = 10),
                            legend.text = element_text(size = 8),
                            legend.key.width = unit(.2, "cm")
                            )
    }
    p <- plot_grid(plotlist = p_list, ncol = 2, align = "hv", axis = "tbl")
    title <- ggdraw() +
            draw_label(paste0("Sample = ", unique(seurat_list_trim[[ii]]$orig.ident)),
              fontface = "bold",
              x = 0,
              hjust = 0,
              vjust = -0.1,
              size = 12)
    plot(plot_grid(title, p, ncol = 1, align = "hv", axis = "tbl", rel_heights = c(.1, 2)))
}
```


### Plot doublets on UMAP


```{r fig.width = 14, fig.height = 6}
doublet_methods <- c("DoubletFinder", "Scrublet", "Solo")
for (i in seq_along(seurat_list_trim)) {
    p_list <- list()
    pt_size <- 4 * AutoPointSize(data = seurat_list_trim[[i]], raster = FALSE)
    for (doublet_method in doublet_methods) {
        p_list[[doublet_method]] <- DimPlot(object = seurat_list_trim[[i]],
                                          group.by = doublet_method,
                                          pt.size = min(4, pt_size),
                                          cols = c("lightsteelblue2", "black"),
                                          order = "Doublet",
                                          raster = FALSE) +
                                    NoAxes() +
                                    coord_fixed() +
                                    ggtitle(doublet_method)
        if (doublet_method != "Solo") {
            p_list[[doublet_method]] <- p_list[[doublet_method]] + NoLegend()
        }
    }
    print(paste0("Predicted doublets for: ", unique(seurat_list_trim[[i]]$orig.ident)))
    print(plot_grid(plotlist = p_list, ncol = 3, rel_widths = c(1, 1, 1.2)))
}
```




### nUMI in doublets

#### nUMI on a UMAP

Plot the nUMI and nGenes on the UMAP

```{r fig.width = 12, fig.height = 6}
p_list <- list()
my_features <- c("nFeature_RNA", "nCount_RNA")
my_max_cut <- c(Inf, 20000)
for (i in seq_along(seurat_list_trim)) {
    pt_size <- 4 * AutoPointSize(data = seurat_list_trim[[i]], raster = FALSE)
    for (ii in seq_along(my_features)) {
        p_list[[ii]] <- FeaturePlot(object = seurat_list_trim[[i]],
                                    max.cutoff = my_max_cut[[ii]],
                                    features = my_features[[ii]],
                                    pt.size = min(4, pt_size),
                                    cols = c("lightsteelblue2", "black"),
                                    order = TRUE,
                                    raster = FALSE) +
                            NoAxes() +
                            coord_fixed()
    }
    title <- ggdraw() +
        draw_label(unique(seurat_list_trim[[i]]@meta.data[["orig.ident"]]),
        fontface = "bold",
        x = 0,
        hjust = 0,
        size = 16)
    p <- plot_grid(plotlist = p_list, ncol = 2)
    print(plot_grid(title, p, ncol = 1, rel_heights = c(0.1, 1)))
}
```



#### nUMI, singlet vs doublet

First we merge the objects to make looking at all the metadata easier.

```{r}
if (length(seurat_list_trim) == 1) {
    seurat_merged <- seurat_list_trim[[1]]
} else {
    seurat_merged <- merge(seurat_list_trim[[1]],
                            unlist(seurat_list_trim[2:length(seurat_list_trim)]),
                            merge.data = TRUE)
}
```




Look at the number of UMI, number of genes, and other qc metrics, in the doublets vs singlets across the methods.


```{r}
options(scipen = 9999)
for (i in seq_along(qc_params_plot)) {
    print(paste0("Plotting values for: ", qc_params_plot[[i]]))
    doublet_methods <- c("DoubletFinder", "Scrublet", "Solo")
    for (doublet_method in doublet_methods) {
        p <- tibble(sample = seurat_merged$orig.ident,
                    metric = seurat_merged@meta.data[[qc_params_plot[[i]]]],
                    doublets = seurat_merged@meta.data[[doublet_method]]) %>%
                ggplot(aes(x = sample, y = metric, facets = doublets)) +
                    labs(x = "Samples", y = qc_params_plot[[i]]) +
                    geom_jitter(size = 0.01, alpha = 0.5, aes(colour = sample)) +
                    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
                    geom_boxplot(outlier.shape = NA, alpha = 0.5) +
                    facet_wrap(~ doublets, ncol = 2) +
                    geom_hline(yintercept = c(0, qc_params_plot_thesh[[i]]), linetype = "dashed") +
                    ggtitle(doublet_method) +
                    theme(plot.title = element_text(face = "bold"))
        if (qc_params_plot[[i]] %in% c("nFeature_RNA", "nCount_RNA")) {
            p <- p +
                scale_y_continuous(trans = "log10") +
                annotation_logticks(sides = "l")
        }
        print(p)
    }
}
```





# Doublets across cell types

First let's extract some count data for a few genes.

```{r}
my_counts <- seurat_merged@assays$RNA@counts[rownames(seurat_merged@assays$RNA@counts) %in% marker_df$gene, ] %>%
    t() %>%
    as.data.frame() %>%
    rownames_to_column("cell_id")
gene_group_table <- seurat_merged@meta.data %>%
    rownames_to_column("cell_id") %>%
    select(cell_id, orig.ident, DoubletFinder, Scrublet, Solo) %>%
    left_join(my_counts, by = "cell_id") %>%
    gather("gene", "expression", -orig.ident, -cell_id, -DoubletFinder, -Scrublet, -Solo) %>%
    gather("doublet_method", "call", -orig.ident, -cell_id, -gene, -expression)
# head(gene_group_table)
```




In this next bit we're looking at the percent of doublets that are positive for a given GOI. Ideally each method should have a similar percent for a given GOI. Or in other words, there should be minimal variation within each row.


```{r warning = F, message = F}
gene_doublet_list <- list()
for (marker in marker_df$gene) {
    gene_doublet_list[[marker]] <- gene_group_table %>%
        filter(call == "Doublet") %>%
        filter(gene == marker) %>%
        group_by(doublet_method) %>%
        summarise(`% of doublets postive for ` = round(100 * sum(expression[gene == marker] != 0) / sum(gene == marker), digits = 1)) %>%
        gather("gene_group", "percent", -doublet_method) %>%
        spread(doublet_method, percent) %>%
        mutate(gene_group = paste0(gene_group, marker))
}
gene_doublet_table <- bind_rows(gene_doublet_list)
gene_doublet_table
```



Now we flip it around and look at the % of cells positive for a given marker that are doublets. Ideally within a method, each marker should give roughly the same percent. Or in other words, there should be minimal variation within each column.


```{r warning = F, message = F}
gene_doublet_list <- list()
for (marker in marker_df$gene) {
    gene_doublet_list[[marker]] <- gene_group_table %>%
        filter(gene == marker) %>%
        group_by(doublet_method) %>%
        summarise(` cells that are doublets` = round(100 * sum(call[expression != 0 & gene == marker] == "Doublet") /
                                                sum(gene[expression != 0] == marker),
                                                digits = 1)
                                                ) %>%
        gather("gene_group", "percent", -doublet_method) %>%
        spread(doublet_method, percent) %>%
        mutate(gene_group = paste0("% of ", marker, gene_group))
}
gene_doublet_table <- bind_rows(gene_doublet_list)
gene_doublet_table
```


# Doublets by cluster


Plot the proportion of cells that are singlets vs doublets by cluster.

```{r clust_contrib_merged, fig.width = 10, fig.height = 4}
doublet_methods <- c("DoubletFinder", "Scrublet", "Solo")
for (i in seq_along(seurat_list_trim)) {
    for (doublet_method in doublet_methods) {
        p <- tibble(clust = seurat_list_trim[[i]]@active.ident,
               doublet = seurat_list_trim[[i]]@meta.data[[doublet_method]]) %>%
            ggplot(aes(x = clust, fill = doublet)) +
                geom_bar(position = "fill", width = 1) +
                scale_fill_manual(values = c("black", "lightsteelblue2")) +
                ggtitle(paste0(doublet_method, " cluster contribution for ", unique(seurat_list_trim[[i]]$orig.ident))) +
                labs(x = "Clusters", y = "Proportion of Cells") +
                theme(axis.text.x = element_blank(),
                      axis.ticks.x = element_blank(),
                      legend.title = element_blank(),
                      legend.position = "bottom",
                      plot.title = element_text(face = "bold"))
        print(p)
    }
}
```





# Remove doublets


```{r remove-doublets}
# add remove doublets and save a trimed seurat list
all_doublets <- all_doublets %>% mutate(doublet_count = rowSums(.[2:4] == "Doublet"))
singlets <- all_doublets %>% filter(doublet_count < cells_remove) %>% select(cell_id)
doublets <- all_doublets %>% filter(doublet_count >= cells_remove) %>% select(cell_id)

seurat_merged <- AddMetaData(object = seurat_merged, metadata = all_doublets$doublet_count, col.name = "doublet_count")
seurat_merged_singlets <- subset(x = seurat_merged, subset = doublet_count < cells_remove)
seurat_list_trim_singlets <- SplitObject(object = seurat_merged_singlets, split.by = "orig.ident")
```


Object list before doublet removal

```{r}
seurat_list_trim
```


Object list after doublet removal

```{r}
seurat_list_trim_singlets
```



# Save data



```{r save-data}
write_csv(all_doublets,
          paste0(doublet_path,
                 dataset, "_",
                 input_type, "_",
                 "allcells_by_method.csv"),
          col_names = TRUE)
write_csv(singlets,
          paste0(doublet_path,
                 dataset, "_",
                 input_type, "_",
                 "singlets.csv"),
          col_names = FALSE)
write_csv(doublets,
          paste0(doublet_path,
                 dataset, "_",
                 input_type, "_",
                 "doublets.csv"),
          col_names = FALSE)

write_rds(x = seurat_list_trim_singlets,
         file = paste0(objects_path,
                      dataset, "_",
                      input_type, "_",
                      "doublets_removed_trimmed_seurat_list.rds")
          )
```



```{r stop-time}
total_end_time <- Sys.time()
total_end_time
total_end_time - total_start_time
```



# CPU and RAM Usage

```{r fig.width = 12, fig.height = 4}
job_id <- commandArgs(trailingOnly = TRUE)[2]
if (job_id != "local") {
    job_log <- paste0("../sps-", job_id, "/sps-", job_id, "-cpu.tsv")
    if (file.exists(job_log)) {
        p1 <- read_tsv(file = job_log) %>%
          gather("Process", "Usage", -`#TIME`, -REQUESTED) %>%
          group_by(`#TIME`) %>%
          ggplot(aes(x = `#TIME`, y = Usage, colour = Process)) +
              geom_line() +
              ggtitle("CPU") +
              theme(legend.position = "none", validate = TRUE)
        p2 <- read_tsv(file = paste0("../sps-", job_id, "/sps-", job_id, "-mem.tsv")) %>%
          gather("Process", "Usage", -`#TIME`, -REQUESTED) %>%
          group_by(`#TIME`) %>%
          ggplot(aes(x = `#TIME`, y = Usage, colour = Process)) +
              geom_line() +
              ggtitle("Memory")
        my_legend <- get_legend(p2)
        p2 <- p2 + theme(legend.position = "none", validate = TRUE)
        plot_grid(p1, p2, my_legend, ncol = 3, rel_widths = c(3, 3, 2))
    }
}
```


# Session info

```{r session-info}
sessionInfo()
```
