---
title: "Ambient RNA removal with SoupX"
description: description
author: "Aaron M. Allen"
date: 'Last update: `r date()`'
output:
  html_document:
    number_sections: true
    code_folding: show
    code_download: true
    theme: cerulean
    df_print: paged
    fig_width: 8.5
    fig_height: 5
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 6
---


```{css, echo = FALSE}
pre, code {white-space:pre !important; overflow-x:auto}
.tocify-item {white-space:pre}
```



```{r echo = FALSE, cache = FALSE}

```


```{r setup, echo = FALSE, cache = FALSE}
ggplot2::theme_set(ggplot2::theme_grey())

rmd_name <- gsub(pattern = ".Rmd", replacement = "", x = knitr::current_input(), ignore.case = TRUE)
knitr::opts_chunk$set(dev = c("png", "cairo_pdf"),
                      fig.align = "center",
                      fig.height = 5,
                      fig.width = 8.5,
                      dpi = 300,
                      pdf.options(encoding = "ISOLatin9.enc"),
                      fig.path = paste0("../analyses/figures/", rmd_name, "-",
                                        format(Sys.time(), "%Y%m%d_%H%M%S"), "/"),
                      fig.retina = 1,
                      warning = TRUE,
                      message = TRUE)
```



# Setup

## Job_ID = `r commandArgs(trailingOnly = TRUE)[2]`


## Start time

```{r start-time}
set.seed(12021)
today <- format(Sys.time(), "%Y%m%d_%H%M%S")
total_start_time <- Sys.time()
total_start_time
```





## Parameters



```{r params}
params_file <- commandArgs(trailingOnly = TRUE)[3]
params_path <- paste0("../src/param_files/", params_file)
print(paste0("The params file is :   ", params_path))
```

```{r, class.source = 'fold-hide', code = readLines(params_path)}
```


Get the name of the machine that this is running on.  If running on my local desktop ("mentok"), then append to the `raw_path` variable.

```{r is-local}
local_path_mod <- ""
curr_machine <- Sys.info()["nodename"]
if (curr_machine == "mentok") {
    local_path_mod <- "/mnt/data2/scRNAseq/cbrg"
}
local_path_mod
```

```{r adjust-path}
raw_path <- paste0(local_path_mod, raw_path)
raw_path
```




## Libraries






```{r packages-etc, warning = FALSE, message = FALSE}
library(Seurat)
library(tidyverse)
library(Matrix)
library(SoupX)
library(cowplot)
library(colorspace)
library(scales)
library(eulerr)
library(ComplexHeatmap)
library(ggcorrplot)
library(ggpubr)
library(corrr)
library(future)
```

## other bits

```{r run-time-tibble}
process_time <- tibble()
stop_the_clock <- function(chunk_name = "unnamed") {
    temp <- tibble::tibble(Code_chunk = chunk_name,
                           Start_time = start_time,
                           Stop_time = Sys.time(),
                           Run_time = difftime(Stop_time, Start_time, units = "mins")
                           )
    process_time <- dplyr::bind_rows(process_time, temp)
    return(process_time)
}
```



```{r make-dir}
dir.create(soupx_output_path, recursive = TRUE, showWarnings = FALSE)
```




## Data

```{r load-data}
start_time <- Sys.time()
if (remove_doublets == FALSE & remove_ambient == TRUE) {
    seurat_list_pre_ambient <- read_rds(file = paste0(objects_path,
                         dataset, "_",
                         input_type, "_",
                         "trimmed_preprocessed_seurat_list.rds")
            )
}
if (remove_doublets == TRUE & remove_ambient == TRUE) {
    seurat_list_pre_ambient <- read_rds(file = paste0(objects_path,
                         dataset, "_",
                         input_type, "_",
                         "doublets_removed_trimmed_preprocessed_seurat_list.rds")
            )
}
seurat_list_pre_ambient
process_time <- stop_the_clock("load_data")
```


## Cell types and Markers

```{r load-markers}
start_time <- Sys.time()
marker_df <- read_tsv(gmt_file_path, col_names = FALSE) %>%
    select(-X2) %>%
    gather("index", "gene", -X1) %>%
    select(-index) %>%
    rename(celltype = X1) %>%
    drop_na() %>%
    group_by(celltype) %>%
    filter(celltype %in% qc_plot_cell_type) %>%
    slice(1) %>%
    arrange(celltype)
marker_df
process_time <- stop_the_clock("load_markers")
```


Load marker gene set to use for `SoupX`

```{r}
mutually_exclusive_markers_df <- read_tsv(gmt_file_path, col_names = FALSE) %>%
    select(-X2) %>%
    gather("index", "gene", -X1) %>%
    select(-index) %>%
    rename(celltype = X1) %>%
    drop_na() %>%
    group_by(celltype) %>%
    filter(celltype %in% mutually_exclusive_markers_celltypes) %>%
    arrange(celltype)
mutually_exclusive_markers_df
```

```{r}
mutually_exclusive_markers <- list()
for (i in seq_along(mutually_exclusive_markers_celltypes)) {
    mutually_exclusive_markers[[i]] <- mutually_exclusive_markers_df %>%
        filter(celltype == mutually_exclusive_markers_celltypes[[i]]) %>%
        pull(gene)
}
names(mutually_exclusive_markers) <- mutually_exclusive_markers_celltypes
mutually_exclusive_markers
```



Then filter marker gene to make sure they are in the object.

```{r genes-present}
all_genes <- c()
for (i in seq_along(seurat_list_pre_ambient)) {
    all_genes <- c(all_genes, rownames(seurat_list_pre_ambient[[i]]))
}
all_genes <- unique(all_genes)
genes_not_found <- marker_df$gene[!marker_df$gene %in% all_genes]
if (length(genes_not_found) > 0) {
    warning(paste0("The following genes were not found: ", genes_not_found))
    marker_df <- marker_df %>% filter(gene != genes_not_found)
} else {
    message(paste0("All genes present in data."))
}
```



Load a list of the singlet cell labels. We will need this for filtering the `CellRanger`/`Alevin` raw counts.

```{r read-singlets}
singlet_cells <- read_csv(paste0("../analyses/doublets/",
                               dataset, "_",
                               input_type, "_",
                               "singlets.csv"),
                          col_names = FALSE
                          )
head(singlet_cells)
```





# Extract cluster labels


```{r extract-clusters, warning = FALSE, fig.height = 10, fig.width = 10}
start_time <- Sys.time()
cluster_labels <- list()
cell_embeddings <- list()
for (i in seq_along(seurat_list_pre_ambient)) {
    Idents(seurat_list_pre_ambient[[i]]) <- seurat_list_pre_ambient[[i]]@meta.data %>%
                                        select(contains(paste0("res.", curr_res / 4))) %>%
                                        deframe()
    cluster_labels[[i]] <- Idents(seurat_list_pre_ambient[[i]])
    names(cluster_labels[[i]]) <- seurat_list_pre_ambient[[i]]@assays[["RNA"]]@data@Dimnames[[2]]
    cell_embeddings[[i]] <- as.data.frame(seurat_list_pre_ambient[[i]]@reductions[["umap"]]@cell.embeddings)
}
process_time <- stop_the_clock("extract_clusters")
```






# Load Background Droplets Data

Loading the raw counts tables for `SoupX` to estimate the "soup"


```{r table-of-droplets}
start_time <- Sys.time()
if (input_type == "cellranger") {
    path_mod <- "/outs/raw_bc_matrix/"
}
if (input_type == "alevin") {
    path_mod <- "/alevin/raw_cellranger_output/"
}

table_of_droplets <- list()
for (i in seq_along(seurat_list_pre_ambient)) {
    raw_name <- raw_names[[which(sample_names == names(seurat_list_pre_ambient)[[i]])]]
    table_of_droplets[[i]] <- readMM(paste0(raw_path, raw_name, path_mod, "matrix.mtx.gz"))
    raw_row_name <- read.table(paste0(raw_path, raw_name, path_mod, "features.tsv.gz"))
    raw_col_name <- read.table(paste0(raw_path, raw_name, path_mod, "barcodes.tsv.gz"))
    table_of_droplets[[i]]@Dimnames[[1]] <- raw_row_name$V2
    table_of_droplets[[i]]@Dimnames[[2]] <- paste0(names(seurat_list_pre_ambient)[[i]], "_", raw_col_name$V1)
}
#table_of_droplets
process_time <- stop_the_clock("table_of_droplets")
```




Load the count data for the "real" cells and remove doublets.

```{r table-of-counts}
start_time <- Sys.time()
if (input_type == "cellranger") {
    path_mod <- "/outs/filtered_bc_matrix/"
}
if (input_type == "alevin") {
    path_mod <- "/alevin/filtered_cellranger_output/"
}

table_of_counts <- list()
for (i in seq_along(seurat_list_pre_ambient)) {
    raw_name <- raw_names[[which(sample_names == names(seurat_list_pre_ambient)[[i]])]]
    table_of_counts[[i]] <- readMM(paste0(raw_path, raw_name, path_mod, "matrix.mtx.gz"))
    raw_row_name <- read.table(paste0(raw_path, raw_name, path_mod, "features.tsv.gz"))
    raw_col_name <- read.table(paste0(raw_path, raw_name, path_mod, "barcodes.tsv.gz"))
    table_of_counts[[i]]@Dimnames[[1]] <- raw_row_name$V2
    table_of_counts[[i]]@Dimnames[[2]] <- paste0(names(seurat_list_pre_ambient)[[i]], "_", raw_col_name$V1)

    singlet_strip <- singlet_cells$X1 %>%
        str_subset(names(seurat_list_pre_ambient)[[i]])
    table_of_counts[[i]] <- table_of_counts[[i]][, colnames(table_of_counts[[i]]) %in% singlet_strip]
}
#table_of_counts
process_time <- stop_the_clock("table_of_counts")
```





```{r round-data}
start_time <- Sys.time()
if (round_the_data) {
    round_to_int <- function(input) {
        input@x <- floor(input@x) + rbinom(length(input@x), 1, input@x - floor(input@x))
        return(input)
    }
    for (i in seq_along(seurat_list_pre_ambient)) {
        table_of_droplets[[i]] <- round_to_int(table_of_droplets[[i]])
        table_of_counts[[i]] <- round_to_int(table_of_counts[[i]])
    }
}
process_time <- stop_the_clock("round_data")
```




If we want to specify a custom range of UMI in the "empty" droplets, then we need to set `calcSoupProfile = FALSE` when calling `SoupChannel`, and then call `estimateSoup` separately and specify a custom range with `soupRange`.

```{r create-soup-channels}
start_time <- Sys.time()
soup_list <- list()
for (i in seq_along(seurat_list_pre_ambient)) {
    soup_list[[i]] <- SoupChannel(table_of_droplets[[i]], table_of_counts[[i]], calcSoupProfile = FALSE)
    soup_list[[i]] <- estimateSoup(soup_list[[i]], soupRange = c(soup_min, soup_max))
}
names(soup_list) <- names(seurat_list_pre_ambient)
soup_list
process_time <- stop_the_clock("create_soup_channels")
```





# Run SoupX


Running `SoupX`. I'm not using the `autoEstCont` as in my initial tests it would return values as high as 95% of all UMIs were soup. Instead I'm feeding it a list of mutually exclusively expressed genes for it to estimate the contamination level. I'm not using this initial estimation done at the cell level, but once we have a number (5-50% depending on the gene lists supplied) the contamination in the `adjustCounts` is estimated at the cluster level.

NB: by default `adjustCounts` does not return integers. These values approximate the integer counts, and if one need integers for other applications you can set `roundToInt = TRUE` (it's `FALSE` by default). I'm not sure if it makes a difference.


```{r run-soupx}
start_time <- Sys.time()
soupx_adj_counts <- list()
use_to_est <- list()
for (i in seq_along(soup_list)) {
    soup_list[[i]] <- setClusters(soup_list[[i]], cluster_labels[[i]])
    soup_list[[i]] <- setDR(soup_list[[i]], cell_embeddings[[i]])
    use_to_est[[i]] <- estimateNonExpressingCells(soup_list[[i]],
                                      nonExpressedGeneList = mutually_exclusive_markers,
                                      clusters = FALSE)
    tryCatch({
        soup_list[[i]] <- calculateContaminationFraction(soup_list[[i]],
                                                   mutually_exclusive_markers,
                                                   useToEst = use_to_est[[i]],
                                                   forceAccept = TRUE
                                                   )
    } , error = function(e) {
            message(paste0("ERROR: ", names(soup_list)[[i]], "\n"))
            message(paste0(e))
            message("Manually setting contaminated fraction to 20% ... \n\n")
            soup_list[[i]] <<- setContaminationFraction(soup_list[[i]],
                                                        contFrac = 0.2
                                                        )
    })
    soupx_adj_counts[[i]] <- adjustCounts(soup_list[[i]], roundToInt = round_the_data)
}
process_time <- stop_the_clock("run_soupx")
```




# Re-Run Seurat Analysis

Re-running the `Seurat` pipeline with the adjusted counts matrix.


```{r create-seurat-adj}
start_time <- Sys.time()
seurat_list_adj_counts <- list()
for (i in seq_along(soupx_adj_counts)) {
    seurat_list_adj_counts[[i]] <- CreateSeuratObject(counts = soupx_adj_counts[[i]],
                                                      min.features = 3,
                                                      project = "sexed_list[[i]]")
}
names(seurat_list_adj_counts) <- names(seurat_list_pre_ambient)
seurat_list_adj_counts
process_time <- stop_the_clock("create_seurat_adj")
```



```{r sct-on-adj-counts, warning = FALSE, message = FALSE}
start_time <- Sys.time()
if (parallelize) {
    plan(futures_plan, workers = futures_n_cores)
    options(future.globals.maxSize = futures_ram * 1024 * 1024^2)
}
if (normalization == "SCT") {
    for (i in seq_along(seurat_list_adj_counts)) {
        seurat_list_adj_counts[[i]] <- SCTransform(seurat_list_adj_counts[[i]],
                                                   return.only.var.genes = FALSE,
                                                   verbose = FALSE)
    }
}
if (parallelize) {
    plan("sequential")
}
process_time <- stop_the_clock("sct_on_adj_counts")
```


```{r lognorm-on-adj-counts, warning = FALSE, message = FALSE}
start_time <- Sys.time()
if (parallelize) {
    plan(futures_plan, workers = futures_n_cores)
    options(future.globals.maxSize = futures_ram * 1024 * 1024^2)
}
if (normalization == "LogNormalize") {
    for (i in seq_along(seurat_list_adj_counts)) {
        seurat_list_adj_counts[[i]] <- NormalizeData(seurat_list_adj_counts[[i]])
        seurat_list_adj_counts[[i]] <- FindVariableFeatures(object = seurat_list_adj_counts[[i]],
                                                              selection.method = var_gene_method,
                                                              nfeatures = n_var_features,
                                                              mean.cutoff = c(min_mean_cutoff, max_mean_cutoff),
                                                              dispersion.cutoff = c(min_dispersion_cutoff, max_dispersion_cutoff)
                                                          )
        seurat_list_adj_counts[[i]]@assays[["RNA"]]@var.features <- seurat_list_adj_counts[[i]]@assays[["RNA"]]@var.features[1:n_var_features]
        seurat_list_adj_counts[[i]] <- ScaleData(seurat_list_adj_counts[[i]],
                                                use.umi = use_umi,
                                                model.use = test_use)
    }
}
if (parallelize) {
    plan("sequential")
}
process_time <- stop_the_clock("lognorm_on_adj_counts")
```



```{r pca-etc-on-adj-counts}
start_time <- Sys.time()
if (parallelize) {
    plan(futures_plan, workers = futures_n_cores)
    options(future.globals.maxSize = futures_ram * 1024 * 1024^2)
}
for (i in seq_along(seurat_list_adj_counts)) {
    seurat_list_adj_counts[[i]] <- RunPCA(seurat_list_adj_counts[[i]],
                                          npcs = curr_pc,
                                          verbose = TRUE)
    seurat_list_adj_counts[[i]] <- FindNeighbors(seurat_list_adj_counts[[i]],
                                                 reduction = "pca",
                                                 dims = 1:curr_pc,
                                                 k.param = neighbours,
                                                 force.recalc = TRUE)
    seurat_list_adj_counts[[i]] <- FindClusters(seurat_list_adj_counts[[i]], resolution = curr_res / 4)
    seurat_list_adj_counts[[i]] <- RunUMAP(object = seurat_list_adj_counts[[i]], dims = 1:curr_pc)
}
seurat_list_adj_counts
if (parallelize) {
    plan("sequential")
}
process_time <- stop_the_clock("pca_etc_on_adj_counts")
```





# Compare the Adjusted Counts


## Replicate correlation


```{r plot-correlation, fig.height = 6,fig.width = 6, warning = FALSE, message = FALSE}
start_time <- Sys.time()
p <- list()
for (i in seq_along(seurat_list_adj_counts)) {
    orginal_counts <- seurat_list_pre_ambient[[i]]@assays[["RNA"]]@counts
    adjusted_counts <- seurat_list_adj_counts[[i]]@assays[["RNA"]]@counts
    common_genes <- intersect(rownames(orginal_counts), rownames(adjusted_counts))
    temp_tibble <- tibble(original = rowSums(orginal_counts[common_genes, ]),
                          adjusted = rowSums(adjusted_counts[common_genes, ])
                          )
    p[[i]] <- temp_tibble %>%
          ggplot(aes(x = log1p(original),
                     y = log1p(adjusted)
                     )
                 ) +
            geom_point() +
            coord_fixed() +
            geom_smooth(method = lm) +
            stat_cor(label.y = 15, size  = 3) +
            stat_regline_equation(label.y = 14, size  = 3) +
            ylim(0, 16) +
            xlim(0, 16) +
            theme_gray() +
            theme(text = element_text(size = 14)) +
        ggtitle(names(seurat_list_adj_counts)[[i]])
    plot(p[[i]])
}
process_time <- stop_the_clock("plot_correlation")
```



## Compare the UMAP Embeddings

Clustering post removal of the soup drastically improves clustering of each sample. Here "improves" means increased separation of the clusters and a general lack of the clump in the middle of the plot. Whether this represents something real and reflective of the biology has yet to be determined.


```{r plot-adj-umap, fig.height = 6, fig.width = 12}
start_time <- Sys.time()
for (i in seq_along(seurat_list_adj_counts)) {
    pt_size <- AutoPointSize(data = seurat_list_pre_ambient[[i]], raster = FALSE)
    p1 <- DimPlot(object = seurat_list_pre_ambient[[i]],
                  pt.size = min(4, pt_size),
                  label = TRUE)  +
        NoAxes() +
        NoLegend() +
        coord_fixed()
    p2 <- DimPlot(object = seurat_list_adj_counts[[i]],
                  pt.size = min(4, pt_size),
                  label = TRUE) +
        NoAxes() +
        NoLegend() +
        coord_fixed()
    p <- plot_grid(p1, p2, labels = c("Doublets removed", "Ambient and doublets removed"), label_size = 10)
    title <- ggdraw() + draw_label(unique(seurat_list_pre_ambient[[i]]@meta.data[["orig.ident"]]),
                                   fontface = "bold",
                                   x = 0,
                                   hjust = 0,
                                   size = 16)
    print(plot_grid(title, p, ncol = 1, rel_heights = c(0.1, 1)))
}
process_time <- stop_the_clock("plot_adj_umap")
```



## Compare cell type markers

SoupX does seem to remove most of the presumed ambient RNA from the markers.


```{r plot-marker-umap, fig.width = 8,fig.height = 24}
start_time <- Sys.time()
p_org <- list()
p_adj <- list()
for (i in seq_along(seurat_list_pre_ambient)) {
    pt_size <- AutoPointSize(data = seurat_list_pre_ambient[[i]], raster = FALSE)
    for (ii in seq_along(marker_df$gene)) {
        p_org[[ii]] <-  FeaturePlot(object = seurat_list_pre_ambient[[i]],
                                   features = marker_df$gene[[ii]],
                                   pt.size = min(4, pt_size),
                                   cols = c("lightsteelblue2", "black"),
                                   order = TRUE) +
            NoAxes() +
            coord_fixed() +
            ggtitle(paste0(marker_df$celltype[[ii]], "\n(", marker_df$gene[[ii]], ")")) +
            theme(plot.title = element_text(size = 10),
                  legend.text = element_text(size = 8),
                  legend.key.width = unit(.2, "cm")
                  )
        p_adj[[ii]] <-  FeaturePlot(object = seurat_list_adj_counts[[i]],
                                   features = marker_df$gene[[ii]],
                                   pt.size = min(4, pt_size),
                                   cols = c("lightsteelblue2", "black"),
                                   order = TRUE) +
            NoAxes() +
            coord_fixed() +
            ggtitle(paste0(marker_df$celltype[[ii]], "\n(", marker_df$gene[[ii]], ")")) +
            theme(plot.title = element_text(size = 10),
                  legend.text = element_text(size = 8),
                  legend.key.width = unit(.2, "cm")
                  )
    }
    p1 <- plot_grid(plotlist = p_org, ncol = 1)
    p2 <- plot_grid(plotlist = p_adj, ncol = 1)
    p <- plot_grid(p1, p2, ncol = 2, labels = c("Original", "Ambient Removed"), hjust = 0, vjust = 0, label_size = 14)
    title <- ggdraw() +
        draw_label(unique(seurat_list_pre_ambient[[i]]@meta.data[["orig.ident"]]),
                   fontface = "bold",
                   x = 0,
                   hjust = 0,
                   vjust = -0.1,
                   size = 20)
    print(plot_grid(title, p, ncol = 1, rel_heights = c(0.1, 2)))
}
process_time <- stop_the_clock("plot_marker_umap")
```




## Euler plots of markers

### Counts function


```{r euler-count-function}
start_time <- Sys.time()
euler_counts <- function(object, markers_list) {
    cell_lists <- list()
    for (i in seq_along(markers_list)) {
        if (length(mutually_exclusive_markers[[i]]) > 1) {
            cell_lists[[i]] <- object@assays[["RNA"]]@counts[markers_list[[i]], ] %>%
                                        t() %>%
                                        as.data.frame() %>%
                                        mutate(my_row_sum = rowSums(.)) %>%
                                        rownames_to_column("cell_id") %>%
                                        filter(my_row_sum >= coexpression_thresh) %>%
                                        pull(cell_id)
        } else {
            cell_lists[[i]] <- object@assays[["RNA"]]@counts[markers_list[[i]], ] %>%
                                        as.data.frame() %>%
                                        rownames_to_column("cell_id") %>%
                                        filter(`.` >= coexpression_thresh) %>%
                                        pull(cell_id)
        }
    }
    names(cell_lists) <- names(markers_list)
    my_counts <- c("1" = length(setdiff(setdiff(cell_lists[[1]], cell_lists[[2]]), cell_lists[[3]])),
                   "2" = length(setdiff(setdiff(cell_lists[[2]], cell_lists[[1]]), cell_lists[[3]])),
                   "3" = length(setdiff(setdiff(cell_lists[[3]], cell_lists[[1]]), cell_lists[[2]])),
                   "1&2" = length(setdiff(intersect(cell_lists[[1]], cell_lists[[2]]), cell_lists[[3]])),
                   "1&3" = length(setdiff(intersect(cell_lists[[1]], cell_lists[[3]]), cell_lists[[2]])),
                   "2&3" = length(setdiff(intersect(cell_lists[[2]], cell_lists[[3]]), cell_lists[[1]])),
                   "1&2&3" = length(intersect(cell_lists[[1]], intersect(cell_lists[[2]], cell_lists[[3]]))))
    names(my_counts) <- c(names(cell_lists)[[1]],
                          names(cell_lists)[[2]],
                          names(cell_lists)[[3]],
                          paste0(names(cell_lists)[[1]], "&", names(cell_lists)[[2]]),
                          paste0(names(cell_lists)[[1]], "&", names(cell_lists)[[3]]),
                          paste0(names(cell_lists)[[2]], "&", names(cell_lists)[[3]]),
                          paste0(names(cell_lists)[[1]], "&", names(cell_lists)[[2]], "&", names(cell_lists)[[3]])
                          )
    return(my_counts)
}
process_time <- stop_the_clock("euler_count_function")
```



```{r plot-euler-diagrams, warning = F,message = F,fig.width = 8,fig.height = 6}
start_time <- Sys.time()
for (ii in seq_along(seurat_list_adj_counts)) {

    my_counts <- euler_counts(seurat_list_pre_ambient[[ii]], mutually_exclusive_markers)
    fit <- euler(my_counts)
    p1 <- plot(fit,
               quantities = TRUE,
               fills = list(fill = colorspace::qualitative_hcl(n = 3, palette = "Set 2"), alpha = 0.8)
    )

    my_counts <- euler_counts(seurat_list_adj_counts[[ii]], mutually_exclusive_markers)
    fit <- euler(my_counts)
    p2 <- plot(fit,
               quantities = TRUE,
               fills = list(fill = colorspace::qualitative_hcl(n = 3, palette = "Set 2"), alpha = 0.8)
    )

    p <- plot_grid(p1, p2, labels = c("Original", "Ambient removed"), hjust = 0, vjust = 0, label_size = 14)
    title <- ggdraw() + draw_label(unique(seurat_list_pre_ambient[[ii]]@meta.data[["orig.ident"]]),
                                   fontface = "bold",
                                   x = 0,
                                   hjust = 0,
                                   vjust = -0.1,
                                   size = 20)
    print(plot_grid(title, p, ncol = 1, rel_heights = c(0.1, 1)))
}
process_time <- stop_the_clock("plot_euler_diagrams")
```



## Upset plots

```{r upset-plot-function}
start_time <- Sys.time()
my_upset_plot <- function(object, markers_list) {
    cell_lists <- list()
    for (i in seq_along(markers_list)) {
        if (length(markers_list[[i]]) > 1) {
            cell_lists[[i]] <- object@assays[["RNA"]]@counts[markers_list[[i]], ] %>%
                                        t() %>%
                                        as.data.frame() %>%
                                        mutate(my_group = if_else(rowSums(.) >= coexpression_thresh, 1, 0)) %>%
                                        select(my_group)
        } else {
            cell_lists[[i]] <- object@assays[["RNA"]]@counts[markers_list[[i]], ] %>%
                                        as.data.frame() %>%
                                        mutate(my_group = if_else(`.` >= coexpression_thresh, 1, 0)) %>%
                                        select(my_group)
        }
        names(cell_lists[[i]]) <- c(names(markers_list)[[i]])
    }
    all_cell_lists <- bind_cols(cell_lists)
    all_cell_lists <- all_cell_lists[-rowSums(all_cell_lists[, 1:4]) != 0, ]
    m <- make_comb_mat(all_cell_lists, mode = "distinct")

    ss <- set_size(m)
    cs <- comb_size(m)
    ht <- UpSet(m,
        set_order = order(ss),
        comb_order = order(comb_degree(m), -cs),
        top_annotation = HeatmapAnnotation(
            "Intersection size" = anno_barplot(cs,
                ylim = c(0, max(cs) * 1.1),
                border = FALSE,
                gp = gpar(fill = "black"),
                height = unit(12, "cm")
            ),
            annotation_name_side = "left",
            annotation_name_rot = 90),
        left_annotation = rowAnnotation(
            "Cells per group" = anno_barplot(-ss,
                baseline = 0,
                axis_param = list(
                    #at = c(0, -500, -1000, -1500),
                    #labels = c(0, 500, 1000, 1500),
                    labels_rot = 0),
                border = FALSE,
                gp = gpar(fill = "black"),
                width = unit(6, "cm")
            ),
            set_name = anno_text(set_name(m),
                location = 0.5,
                just = "center",
                width = max_text_width(set_name(m)) + unit(4, "mm"))
        ),
        right_annotation = NULL,
        show_row_names = FALSE)
    ht <- draw(ht) #+ ggtitle(unique(object@meta.data[["orig.ident"]])
    od <- column_order(ht)
    decorate_annotation("Intersection size", {
        grid.text(cs[od], x = seq_along(cs), y = unit(cs[od], "native") + unit(4, "pt"),
            default.units = "native", just = c("left", "bottom"),
            gp = gpar(fontsize = 10, col = "#404040"), rot = 45)
    })
}
process_time <- stop_the_clock("upset_plot_function")
```

### Original data

```{r upset-plot-original, fig.width = 10, fig.height = 6}
start_time <- Sys.time()
for (i in seq_along(seurat_list_pre_ambient)) {
    cat(paste0("\n", names(seurat_list_pre_ambient)[[i]]))
    my_upset_plot(seurat_list_pre_ambient[[i]], mutually_exclusive_markers)
}
process_time <- stop_the_clock("upset_plot_original")
```


### Adjusted data

```{r upset-plot-adjusted, fig.width = 10, fig.height = 6}
start_time <- Sys.time()
for (i in seq_along(seurat_list_adj_counts)) {
    cat(paste0("\n", names(seurat_list_adj_counts)[[i]]))
    my_upset_plot(seurat_list_adj_counts[[i]], mutually_exclusive_markers)
}
process_time <- stop_the_clock("upset_plot_adjusted")
```








# Save adjusted data

```{r save-data}
start_time <- Sys.time()

write_rds(x = soup_list,
           file = paste0(soupx_output_path,
                         dataset, "_",
                         input_type, "_",
                         "soupx_soup_list.rds")
            )
write_rds(x = soupx_adj_counts,
           file = paste0(soupx_output_path,
                         dataset, "_",
                         input_type, "_",
                         "soupx_adj_counts.rds")
            )
write_rds(x = seurat_list_adj_counts,
           file = paste0(soupx_output_path,
                         dataset, "_",
                         input_type, "_",
                         "soupx_seurat_list_adj.rds")
            )
process_time <- stop_the_clock("save_data")
```







# Run time

```{r run-times}
process_time
```

```{r plot-run-times, fig.width = 10, fig.height = 6}
process_time$Code_chunk <- factor(process_time$Code_chunk, levels = rev(process_time$Code_chunk))
process_time %>%
    ggplot(aes(x = Code_chunk, y = Run_time, fill = Run_time)) +
    geom_col(width = 0.8) +
    xlab("Code Chunk") +
    ylab("Run Time (mins)") +
    coord_flip()
```



```{r stop-time}
total_end_time <- Sys.time()
total_end_time
total_end_time - total_start_time
```



# CPU and RAM Usage

```{r fig.width = 12, fig.height = 4}
job_id <- commandArgs(trailingOnly = TRUE)[2]
if (job_id != "local") {
    job_log <- paste0("../sps-", job_id, "/sps-", job_id, "-cpu.tsv")
    if (file.exists(job_log)) {
        p1 <- read_tsv(file = job_log) %>%
          gather("Process", "Usage", -`#TIME`, -REQUESTED) %>%
          group_by(`#TIME`) %>%
          ggplot(aes(x = `#TIME`, y = Usage, colour = Process)) +
              geom_line() +
              ggtitle("CPU") +
              theme(legend.position = "none", validate = TRUE)
        p2 <- read_tsv(file = paste0("../sps-", job_id, "/sps-", job_id, "-mem.tsv")) %>%
          gather("Process", "Usage", -`#TIME`, -REQUESTED) %>%
          group_by(`#TIME`) %>%
          ggplot(aes(x = `#TIME`, y = Usage, colour = Process)) +
              geom_line() +
              ggtitle("Memory")
        my_legend <- get_legend(p2)
        p2 <- p2 + theme(legend.position = "none", validate = TRUE)
        plot_grid(p1, p2, my_legend, ncol = 3, rel_widths = c(3, 3, 2))
    }
}
```


# Session info

```{r session-info}
sessionInfo()
```
