---
title: "Normalization and Batch Correction of scRNA-Seq data"
description:
author: "Aaron M. Allen"
date: 'Last update: `r date()`'
output:
  html_document:
    number_sections: true
    code_folding: show
    code_download: true
    theme: cerulean
    df_print: paged
    fig_width: 8.5
    fig_height: 5
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 6
---



```{css, echo = FALSE}
pre, code {white-space:pre !important; overflow-x:auto}
.tocify-item {white-space:pre}
```



```{r echo = FALSE, cache = FALSE}

```


```{r setup, echo = FALSE, cache = FALSE}
ggplot2::theme_set(ggplot2::theme_grey())    # base_family = "Arial"))
options(scipen = 9999)

rmd_name <- gsub(pattern = ".Rmd", replacement = "", x = knitr::current_input(), ignore.case = TRUE)
knitr::opts_chunk$set(dev = c("png", "cairo_pdf"),
                      fig.align = "center",
                      fig.height = 5,
                      fig.width = 8.5,
                      dpi = 300,
                      pdf.options(encoding = "ISOLatin9.enc"),
                      fig.path = paste0("../analyses/figures/", rmd_name, "-",
                                        format(Sys.time(), "%Y%m%d_%H%M%S"), "/"),
                      fig.retina = 1,
                      warning = TRUE,
                      message = TRUE)
```



# Setup

## Job_ID = `r commandArgs(trailingOnly = TRUE)[2]`


## Start time

```{r start-time}
set.seed(12021)
today <- format(Sys.time(), "%Y%m%d_%H%M%S")
total_start_time <- Sys.time()
total_start_time
```







## Parameters



```{r params}
params_file <- commandArgs(trailingOnly = TRUE)[3]
if (is.na(params_file)) {
    params_file <- "pipeline_R_params.R"
}
params_path <- paste0("../src/param_files/", params_file)
print(paste0("The params file is :   ", params_path))
```

```{r, class.source = 'fold-hide', code = readLines(params_path)}
```

```{r run-number}
run_number_4a <- commandArgs(trailingOnly = TRUE)[4]
run_number_4a
```

Get the name of the machine that this is running on.  If running on my local desktop ("mentok"), then append to the `reference_object_path` variable.

```{r is-local}
local_path_mod <- ""
curr_machine <- Sys.info()["nodename"]
if (curr_machine == "mentok") {
    local_path_mod <- "/mnt/data2/scRNAseq/cbrg"
}
local_path_mod
```




## Libraries

```{r packages-etc, warning = FALSE, message = FALSE}
library(utils)
library(Seurat)
library(SingleCellExperiment)
library(harmony)
library(tidyverse)
library(cowplot)
library(pheatmap)
library(colorspace)
library(ggcorrplot)
library(ggpubr)
library(corrr)
library(ape)        # used for phylogram during cluster collapse
library(future)
```




```{r run-time-tibble}
process_time <- tibble()
stop_the_clock <- function(chunk_name = "unnamed") {
    temp <- tibble::tibble(Code_chunk = chunk_name,
                           Start_time = start_time,
                           Stop_time = Sys.time(),
                           Run_time = difftime(Stop_time, Start_time, units = "mins")
                           )
    process_time <- dplyr::bind_rows(process_time, temp)
    return(process_time)
}
```


```{r parallel-options}
# if (parallelize) {
    options(future.globals.maxSize = futures_ram * 1024 * 1024^2)
# }
```



# The data

## Load data




```{r load-data}
start_time <- Sys.time()
if (remove_doublets == FALSE && remove_ambient == FALSE) {
    path_modifier <- ""
}
if (remove_doublets == TRUE && remove_ambient == FALSE) {
    path_modifier <- "_doublets_removed"
}
if (remove_doublets == FALSE && remove_ambient == TRUE) {
    path_modifier <- "_ambient_removed"
}
if (remove_doublets == TRUE && remove_ambient == TRUE) {
    path_modifier <- "_doublets_removed_ambient_removed"
}

if (run_number_4a == 1) {
    seurat_list_trim <- read_rds(file = paste0(objects_path,
                             dataset, "_",
                             input_type,
                             path_modifier, "_",
                             "trimmed_seurat_list.rds")
                )
    if (user_cells_remove) {
        cells_to_keep <- read_csv(file = user_cells_remove_file) %>% pull(cell_id)
        for (i in seq_along(seurat_list_trim)) {
            seurat_list_trim[[i]] <- subset(x = seurat_list_trim[[i]], cells = cells_to_keep)
        }
    }
}
if (run_number_4a == 2) {
    # for each string in "annotation_cell_type" vector, we need to pass the arguement into the script for subclustering
    subcluster_celltype <- commandArgs(trailingOnly = TRUE)[5]
    print(paste0("Loading for subclustering: ", subcluster_celltype))
    seurat_list_trim <- read_rds(file = paste0(objects_path,
                         dataset, "_",
                         input_type, "_",
                         "seurat_list",
                         path_modifier, "_",
                         subcluster_celltype, "_",
                         "subclustering.rds")
                     )
     curr_pc <- subcluster_pcs[which(annotation_cell_type == subcluster_celltype)]
     curr_res <- subcluster_res[which(annotation_cell_type == subcluster_celltype)]
}
seurat_list_trim
process_time <- stop_the_clock("load_data")
```







## Cell types and Markers

```{r load-markers}
start_time <- Sys.time()
marker_df <- read_tsv(paste0(gmt_file_path), col_names = FALSE) %>%
    select(-X2) %>%
    gather("index", "gene", -X1) %>%
    select(-index) %>%
    rename(celltype = X1) %>%
    drop_na() %>%
    group_by(celltype) %>%
    filter(celltype %in% qc_plot_cell_type) %>%
    slice(1) %>%
    arrange(celltype)
marker_df
process_time <- stop_the_clock("load_markers")
```


Then filter marker gene to make sure they are in the object.


```{r genes-present}
all_genes <- c()
for (i in seq_along(seurat_list_trim)) {
    all_genes <- c(all_genes, rownames(seurat_list_trim[[i]]))
}
all_genes <- unique(all_genes)
genes_not_found <- marker_df$gene[!marker_df$gene %in% all_genes]
if (length(genes_not_found) > 0) {
    warning(paste0("The following genes were not found: ", genes_not_found))
    marker_df <- marker_df %>% filter(gene != genes_not_found)
} else {
    message(paste0("All genes present in data."))
}
```





# Normalize

```{r normalize, warning = FALSE, message = FALSE}
start_time <- Sys.time()
if (normalization == "SCT") {
    print(paste0("Normalization set to ", normalization))
     for (i in seq_along(seurat_list_trim)) {
        print(paste0("SCTransform-ing ", seurat_list_trim[[i]]@project.name))
        seurat_list_trim[[i]] <- SCTransform(seurat_list_trim[[i]],
                                            return.only.var.genes = FALSE,
                                            conserve.memory = FALSE,
                                            vars.to.regress = sct_vars_to_regress,
                                            verbose = FALSE,
                                            variable.features.n = n_var_features)
    }
}
if (normalization == "LogNormalize") {
    print(paste0("Normalization RNA slot"))
    for (i in seq_along(seurat_list_trim)) {
        print(paste0("Normalizing RNA slot of ", seurat_list_trim[[i]]@project.name))
        seurat_list_trim[[i]] <- NormalizeData(object = seurat_list_trim[[i]], assay = "RNA")
    }
}
if (normalization == "LogNormalize") {
    for (i in seq_along(seurat_list_trim)) {
        print(paste0("Finding variable features in ",
                     seurat_list_trim[[i]]@project.name,
                     " with 'FindVariableFeatures'"))

        seurat_list_trim[[i]] <- FindVariableFeatures(object = seurat_list_trim[[i]],
                                                      selection.method = var_gene_method,
                                                      nfeatures = n_var_features,
                                                      mean.cutoff = c(min_mean_cutoff, max_mean_cutoff),
                                                      dispersion.cutoff = c(min_dispersion_cutoff, max_dispersion_cutoff)
                                                      )

        seurat_list_trim[[i]]@assays[["RNA"]]@var.features <- seurat_list_trim[[i]]@assays[["RNA"]]@var.features[1:n_var_features]
    }
}
process_time <- stop_the_clock("normalize")
```



# Integrate



```{r integrate-data}
start_time <- Sys.time()

if (reduce_by_var_genes) {
    var_features <- SelectIntegrationFeatures(object.list = seurat_list_trim,
                                              nfeatures = n_var_features,
                                              selection.method = var_gene_method,
                                              mean.cutoff = c(min_mean_cutoff, max_mean_cutoff),
                                              dispersion.cutoff = c(min_dispersion_cutoff, max_dispersion_cutoff)
                                              )
} 


if (integration == "none") {
    integrated <- seurat_list_trim[[1]]
} else {
    if (integration == "cca") {
        if (normalization == "SCT") {
            seurat_list_trim <- PrepSCTIntegration(object.list = seurat_list_trim,
                                              anchor.features = var_features,
                                              verbose = FALSE)
        }
        if (parallelize) {
            plan(futures_plan, workers = futures_n_cores)
            options(future.globals.maxSize = futures_ram * 1024 * 1024^2)
        }
        anchors <- FindIntegrationAnchors(object.list = seurat_list_trim,
                                              dims = 1:max_pc,
                                              scale = TRUE,
                                              normalization.method = normalization,
                                              reduction = "cca",
                                              anchor.features = var_features,
                                              nn.method = nn_method,
                                              n.trees = nn_trees,
                                              eps = nn_eps,
                                              verbose = FALSE)
        if (parallelize) {
            plan("sequential")
        }
        all_features <- lapply(seurat_list_trim, row.names) %>% Reduce(intersect, .)
        integrated <- IntegrateData(anchorset = anchors,
                                 normalization.method = normalization,
                                 features.to.integrate = all_features,
                                 verbose = FALSE)
    }

    if (integration == "harmony") {
        integrated <- merge(seurat_list_trim[[1]],
                            unlist(seurat_list_trim[2:length(seurat_list_trim)]),
                            merge.data = TRUE)
    }
}




# if variable genes haven't been selected, set to all genes
if (!reduce_by_var_genes) {
    var_features <- rownames(integrated)
}
# if reducing by user supplied gene list (such as TFs), subset var_features to only these genes
if (reduce_by_user_gene_list) {
    user_genes <- read_tsv(file = user_gene_list_file, col_names = "gene")
    var_features <- dplyr::intersect(var_features, user_genes$gene)
}
# remove gene-specific isoforms from variable genes
var_features <- setdiff(var_features, iso_genes_remove)

# if there are any specific genes you want to make sure are used, add
# them here (i.e. FAN markers)
if (user_genes_add) {
    var_features <- unique(dplyr::union(var_features, genes_add))
}

VariableFeatures(integrated) <- var_features

process_time <- stop_the_clock("integrate_data")
```


# Run PCA

```{r scale-data}
start_time <- Sys.time()
# if (parallelize) {
#     plan(futures_plan, workers = futures_n_cores)
    options(future.globals.maxSize = futures_ram * 1024 * 1024^2)
# }
if (normalization == "LogNormalize") {
    if (integration == "cca") {
        DefaultAssay(object = integrated) <- "integrated"
        assay_use <- "integrated"
    }
    if (integration == "harmony" || integration == "none") {
        DefaultAssay(object = integrated) <- "RNA"
        assay_use <- "RNA"
    }
    if (scale_all_features) {
       features_to_scale <- rownames(integrated)
    } else {
       features_to_scale <- VariableFeatures(integrated)
    }
    integrated <- ScaleData(object = integrated,
                            assay = assay_use,
                            features = features_to_scale,
                            vars.to.regress = vars_to_regress)
}
# if (parallelize) {
#     plan("sequential")
# }
integrated
process_time <- stop_the_clock("scale_data")
```


```{r run-pca}
# if (parallelize) {
    plan(futures_plan, workers = futures_n_cores)
    options(future.globals.maxSize = futures_ram * 1024 * 1024^2)
# }

integrated <- RunPCA(object = integrated, features = var_features, npcs = max_pc)

# if (parallelize) {
#     plan("sequential")
# }


invisible(gc())
process_time <- stop_the_clock("run_pca")
```



# Run Harmony

```{r run-harmony}
start_time <- Sys.time()
if (integration == "harmony") {
    if (normalization == "SCT") {
        assay_use <- "SCT"
    } else if (normalization == "LogNormalize") {
        assay_use <- "RNA"
    }
    integrated <- RunHarmony(integrated,
                            group.by.vars = harmony_group,
                            assay.use = assay_use,
                            theta = harmony_theta,      # Diversity clustering penalty parameter
                            lambda = harmony_lambda,    # Ridge regression penalty parameter
                            nclust = harmony_nclust,    # Number of clusters in model
                            tau = harmony_tau)          # expected number of cells per cluster
}
invisible(gc())
process_time <- stop_the_clock("run_harmony")
```



# Visualize

## Plot PCA



```{r elbow-plot, fig.width = 12, fig.height = 6}
start_time <- Sys.time()
test_pcs <- seq(min_pc, max_pc, step_pc)
ElbowPlot(integrated, reduction = "pca", ndims = max_pc) +
    geom_vline(xintercept = seq(5, max_pc, 5), linetype = "dashed", color = "grey40") +
    ylim(0, 10)
process_time <- stop_the_clock("elbow_plot")
```


```{r plot-pca-heatmap, fig.width = 12, fig.height = 20, warning = FALSE, message = FALSE}
start_time <- Sys.time()
if (integration == "cca" || integration == "none") {
    set_reduction <- "pca"
} else if (integration == "harmony") {
    set_reduction <- "harmony"
}
p_list <- list()
for (i in 1:min(max_pc, 10)) {
    p_list[[i]] <- DimHeatmap(integrated,
                            reduction = set_reduction,
                            dims = i,
                            cells = 500,
                            balanced = TRUE,
                            fast = FALSE) +
        scale_fill_gradientn(colors = diverging_hcl(10, palette = "Vik"))
}
plot_grid(plotlist = p_list, ncol = 2)
rm(p_list)
invisible(gc())
process_time <- stop_the_clock("plot_pca_heatmap")
```


## Run UMAP


```{r run-umap}
start_time <- Sys.time()
if (integration == "cca" || integration == "none") {
    set_reduction <- "pca"
}
if (integration == "harmony") {
    set_reduction <- "harmony"
}

for (test_pc in test_pcs) {
    integrated <- RunUMAP(integrated,
        reduction = set_reduction,
        dims = 1:test_pc,
        n.neighbors = neighbours,
        n.epochs = umap_n_epochs,
        min.dist = umap_min_dist,
        spread = umap_spread,
        fast = FALSE,
        reduction.name = paste0("umap_", test_pc, "pcs"),
        reduction.key = paste0("umap_", test_pc, "pcs_"))
    #integrated@reductions[[paste0("umap_",test_pc,"pcs")]] <- integrated@reductions[["umap"]]
}
process_time <- stop_the_clock("run_umap")
```




### Plot UMAP Embeddings

```{r plot-umap, fig.width = 12, fig.height = 12}
start_time <- Sys.time()
if (normalization == "SCT") {
    default_assay <- "SCT"
}
if (normalization == "LogNormalize") {
    default_assay <- "RNA"
}

DefaultAssay(object = integrated) <- default_assay
DefaultAssay(object = integrated)
pt_size <- 4 * AutoPointSize(data = integrated, raster = FALSE)

for (test_pc in test_pcs) {
    set_reduction <- paste0("umap_", test_pc, "pcs")
    col_list <- scales::hue_pal()(length(unique(integrated@meta.data[["experiment"]])))
    print(paste0("PC = ", test_pc))
    print(paste0("Tissue = ", dataset,
                 ", Normalization = ", normalization,
                 ", Integration = ", integration
                 )
             )
    for (i in seq_along(unique(integrated@meta.data[["experiment"]]))) {
      my_cells <- WhichCells(object = integrated,
                             expression = experiment == unique(integrated@meta.data[["experiment"]])[i])
      print(DimPlot(object = integrated,
                        reduction = set_reduction,
                        cells.highlight = my_cells,
                        cols = "lightgrey",
                        cols.highlight = col_list[[i]],
                        sizes.highlight = 1.2 * min(4, pt_size),
                        pt.size = min(4, pt_size),
                        raster = FALSE) +
                    NoLegend() +
                    NoAxes() +
                    coord_fixed() +
                    ggtitle(paste0(unique(integrated@meta.data[["experiment"]])[i],
                            "\n", length(my_cells), " cells")
                            )
            )
    }
    for (i in seq_along(marker_df$gene)) {
      print(FeaturePlot(object = integrated,
                            features = marker_df$gene[[i]],
                            reduction = set_reduction,
                            pt.size = min(4, pt_size),
                            cols = c("lightsteelblue2", "black"),
                            order = TRUE,
                            raster = FALSE) +
                    NoAxes() +
                    coord_fixed() +
                    ggtitle(paste0(marker_df$celltype[[i]], "\n(", marker_df$gene[[i]], ")")) +
                    theme(plot.title = element_text(size = 10),
                          legend.text = element_text(size = 8),
                          legend.key.width = unit(.2, "cm")
                        )
              )
    }
}
integrated@reductions[["umap"]] <- integrated@reductions[[paste0("umap_", curr_pc, "pcs")]]
process_time <- stop_the_clock("plot_umap")
```



## Save integrated data


```{r save-end}
start_time <- Sys.time()
if (run_number_4a == 1) {
    write_rds(x = integrated,
              file = paste0(objects_path,
                          dataset, "_",
                          input_type, "_",
                          normalization, "_",
                          integration, "_",
                          "integrated_seurat",
                          path_modifier,
                          ".rds")
              )
}
if (run_number_4a == 2) {
    # for each string in "annotation_cell_type" vector, we need to pass the arguement into the script for subclustering
    write_rds(x = integrated,
            file = paste0(objects_path,
                          dataset, "_",
                          input_type, "_",
                          normalization, "_",
                          integration, "_",
                          "integrated_seurat",
                          path_modifier, "_",
                          subcluster_celltype, "_",
                          "subclusting.rds")
                  )
}
process_time <- stop_the_clock("save_end")
```




```{r}
integrated_diet <- DietSeurat(object = integrated, 
                              counts = TRUE, 
                              data = TRUE, 
                              scale.data = FALSE, 
                              features = NULL, 
                              assays = "RNA", 
                              dimreducs = names(integrated@reductions), 
                              graphs = NULL
                              )

write_rds(x = integrated_diet,
          file = paste0(objects_path,
                        dataset, "_",
                        input_type, "_",
                        normalization, "_",
                        integration, "_",
                        "integrated_seurat",
                        path_modifier,
                        "_diet.rds")
          )

if (dim(integrated)[2] > 10^6) {
    integrated_diet_split <- SplitObject(object = integrated_diet,
                                         split.by = "sex")
    write_rds(x = integrated_diet_split[["female"]],
              file = paste0(objects_path,
                            dataset, "_", "female_",
                            input_type, "_",
                            normalization, "_",
                            integration, "_",
                            "integrated_seurat",
                            path_modifier,
                            "_diet.rds")
            )
    write_rds(x = integrated_diet_split[["male"]],
              file = paste0(objects_path,
                            dataset, "_", "male_",
                            input_type, "_",
                            normalization, "_",
                            integration, "_",
                            "integrated_seurat",
                            path_modifier,
                            "_diet.rds")
            )
}
```





# Run time

```{r run-times}
process_time
```

```{r plot-run-times, fig.width = 10, fig.height = 4}
process_time$Code_chunk <- factor(process_time$Code_chunk, levels = rev(process_time$Code_chunk))
process_time %>%
    ggplot(aes(x = Code_chunk, y = Run_time, fill = Run_time)) +
    geom_col(width = 0.8) +
    xlab("Code Chunk") +
    ylab("Run Time (mins)") +
    coord_flip()
```



```{r stop-time}
total_end_time <- Sys.time()
total_end_time
total_end_time - total_start_time
```



# CPU and RAM Usage

```{r fig.width = 12, fig.height = 4}
job_id <- commandArgs(trailingOnly = TRUE)[2]
if (is.na(job_id)) {
    job_id <- "interactive"
}
if (job_id != "local" && job_id != "interactive") {
    job_log <- paste0("../sps-", job_id, "/sps-", job_id, "-cpu.tsv")
    if (file.exists(job_log)) {
        p1 <- read_tsv(file = job_log) %>%
          gather("Process", "Usage", -`#TIME`, -REQUESTED) %>%
          group_by(`#TIME`) %>%
          ggplot(aes(x = `#TIME`, y = Usage, colour = Process)) +
              geom_line() +
              ggtitle("CPU") +
              theme(legend.position = "none", validate = TRUE)
        p2 <- read_tsv(file = paste0("../sps-", job_id, "/sps-", job_id, "-mem.tsv")) %>%
          gather("Process", "Usage", -`#TIME`, -REQUESTED) %>%
          group_by(`#TIME`) %>%
          ggplot(aes(x = `#TIME`, y = Usage, colour = Process)) +
              geom_line() +
              ggtitle("Memory")
        my_legend <- get_legend(p2)
        p2 <- p2 + theme(legend.position = "none", validate = TRUE)
        plot_grid(p1, p2, my_legend, ncol = 3, rel_widths = c(3, 3, 2))
    }
}
```





# Session info

```{r session-info}
sessionInfo()
```
