---
title: "Run tSNE"
description: description
author: "Aaron M. Allen"
date: 'Last update: `r date()`'
output:
  html_document:
    number_sections: true
    code_folding: show
    code_download: true
    theme: cerulean
    df_print: paged
    fig_width: 8.5
    fig_height: 5
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 6
---



```{css, echo = FALSE}
pre, code {white-space:pre !important; overflow-x:auto}
.tocify-item {white-space:pre}
```



```{r echo = FALSE, cache = FALSE}

```


```{r setup, echo = FALSE, cache = FALSE}
ggplot2::theme_set(ggplot2::theme_grey())    # base_family = "Arial"))
options(scipen = 9999)

rmd_name <- gsub(pattern = ".Rmd", replacement = "", x = knitr::current_input(), ignore.case = TRUE)
knitr::opts_chunk$set(dev = c("png", "cairo_pdf"),
                      fig.align = "center",
                      fig.height = 5,
                      fig.width = 8.5,
                      dpi = 300,
                      pdf.options(encoding = "ISOLatin9.enc"),
                      fig.path = paste0("../analyses/figures/", rmd_name, "-",
                                        format(Sys.time(), "%Y%m%d_%H%M%S"), "/"),
                      fig.retina = 1,
                      warning = TRUE,
                      message = TRUE)
```



# Setup

## Job_ID = `r commandArgs(trailingOnly = TRUE)[2]`


## Start time

```{r start-time}
set.seed(12021)
today <- format(Sys.time(), "%Y%m%d_%H%M%S")
total_start_time <- Sys.time()
total_start_time
```






## Parameters



```{r params}
params_file <- commandArgs(trailingOnly = TRUE)[3]
if (is.na(params_file)) {
    params_file <- "pipeline_R_params.R"
}
params_path <- paste0("../src/param_files/", params_file)
print(paste0("The params file is :   ", params_path))
```

```{r, class.source = 'fold-hide', code = readLines(params_path)}
```

```{r run-number}
run_number_4b <- commandArgs(trailingOnly = TRUE)[4]
run_number_4b
```

Get the name of the machine that this is running on.  If running on my local desktop ("mentok"), then append to the `reference_object_path` variable.

```{r is-local}
local_path_mod <- ""
curr_machine <- Sys.info()["nodename"]
if (curr_machine == "mentok") {
    local_path_mod <- "/mnt/data2/scRNAseq/cbrg"
}
local_path_mod
```




## Libraries

```{r packages-etc, warning = FALSE, message = FALSE}
library(utils)
library(Seurat)
library(SingleCellExperiment)
library(harmony)
library(tidyverse)
library(cowplot)
library(pheatmap)
library(colorspace)
library(ggcorrplot)
library(ggpubr)
library(corrr)
library(ape)        # used for phylogram during cluster collapse
library(future)
```




```{r run-time-tibble}
process_time <- tibble()
stop_the_clock <- function(chunk_name = "unnamed") {
    temp <- tibble::tibble(Code_chunk = chunk_name,
                           Start_time = start_time,
                           Stop_time = Sys.time(),
                           Run_time = difftime(Stop_time, Start_time, units = "mins")
                           )
    process_time <- dplyr::bind_rows(process_time, temp)
    return(process_time)
}
```





# The data

## Load data

```{r load-data}
start_time <- Sys.time()
if (remove_doublets == FALSE && remove_ambient == FALSE) {
    path_modifier <- ""
}
if (remove_doublets == TRUE && remove_ambient == FALSE) {
    path_modifier <- "_doublets_removed"
}
if (remove_doublets == FALSE && remove_ambient == TRUE) {
    path_modifier <- "_ambient_removed"
}
if (remove_doublets == TRUE && remove_ambient == TRUE) {
    path_modifier <- "_doublets_removed_ambient_removed"
}

if (run_number_4b == 1) {
    integrated <- read_rds(file = paste0(objects_path,
                                      dataset, "_",
                                      input_type, "_",
                                      normalization, "_",
                                      integration, "_",
                                      "integrated_seurat",
                                      path_modifier,
                                      ".rds")
                              )
}
if (run_number_4b == 2) {
# for each string in "annotation_cell_type" vector, we need to pass the arguement into the script for subclustering
    subcluster_celltype <- commandArgs(trailingOnly = TRUE)[5]
    print(paste0("Loading for subclustering: ", subcluster_celltype))
    integrated <- read_rds(file = paste0(objects_path,
                                      dataset, "_",
                                      input_type, "_",
                                      normalization, "_",
                                      integration, "_",
                                      "integrated_seurat",
                                      path_modifier, "_",
                                      subcluster_celltype, "_",
                                      "subclusting.rds")
    )
    curr_pc <- subcluster_pcs[which(annotation_cell_type == subcluster_celltype)]
    curr_res <- subcluster_res[which(annotation_cell_type == subcluster_celltype)]
}
process_time <- stop_the_clock("load_data")
```

```{r load-markers}
start_time <- Sys.time()
marker_df <- read_tsv(paste0(local_path_mod,gmt_file_path), col_names = FALSE) %>%
    select(-X2) %>%
    gather("index", "gene", -X1) %>%
    select(-index) %>%
    rename(celltype = X1) %>%
    drop_na() %>%
    group_by(celltype) %>%
    filter(celltype %in% qc_plot_cell_type) %>%
    slice(1) %>%
    arrange(celltype)
marker_df
process_time <- stop_the_clock("load_markers")
```


Then filter marker gene to make sure they are in the object.

```{r genes-present}
all_genes <- rownames(integrated)
genes_not_found <- marker_df$gene[!marker_df$gene %in% all_genes]
if (length(genes_not_found) > 0) {
    warning(paste0("The following genes were not found: ", genes_not_found))
    marker_df <- marker_df %>% filter(gene != genes_not_found)
} else {
    message(paste0("All genes present in data."))
}
```





## Run tSNE


```{r run-tsne, fig.width = 12, fig.height = 12}
start_time <- Sys.time()
if (integration == "cca") {
    set_reduction <- "pca"
}
if (integration == "harmony") {
    set_reduction <- "harmony"
}

test_pcs <- seq(min_pc, max_pc, step_pc)

tsne_embeddings <- list()
for (test_pc in test_pcs) {
    integrated <- RunTSNE(integrated,
                     reduction = set_reduction,
                     num_threads = tsne_n_cores,
                     verbose = FALSE,
                     dims = 1:test_pc,
                     do.fast = tSNE_do_fast,
                     check_duplicates = tSNE_check_duplicates,
                     max_iter = tSNE_max_iter,
                     theta = tSNE_theta,
                     # perplexity = min(floor((dim(integrated)[2] - 1) / 3), floor(dim(integrated)[2]/100), 500),
                     reduction.name = paste0("tsne_", test_pc, "pcs"),
                     reduction.key = paste0("tsne", test_pc, "pcs_"))
    #integrated@reductions[[paste0("tsne_",test_pc,"pcs")]] <- integrated@reductions[["tsne"]]
    tsne_embeddings[[paste0("tsne_", test_pc, "pcs")]] <- integrated@reductions[[paste0("tsne_", test_pc, "pcs")]]

}
process_time <- stop_the_clock("run_tsne")
```




### Plot tSNE Embeddings

```{r plot-tsne, fig.width = 12, fig.height = 12}
start_time <- Sys.time()
if (normalization == "SCT") {
    default_assay <- "SCT"
}
if (normalization == "LogNormalize") {
    default_assay <- "RNA"
}

DefaultAssay(object = integrated) <- default_assay
DefaultAssay(object = integrated)
pt_size <- 4 * AutoPointSize(data = integrated, raster = FALSE)

for (test_pc in test_pcs) {
    print(paste0("PC = ", test_pc))
    print(paste0("Tissue = ", dataset,
                 ", Normalization = ", normalization,
                 ", Integration = ", integration
                 )
             )
    set_reduction <- paste0("tsne_", test_pc, "pcs")
    col_list <- scales::hue_pal()(length(unique(integrated@meta.data[["experiment"]])))
    for (i in seq_along(unique(integrated@meta.data[["experiment"]]))) {
      my_cells <- WhichCells(object = integrated,
                             expression = experiment == unique(integrated@meta.data[["experiment"]])[i])
      print(DimPlot(object = integrated,
                        reduction = set_reduction,
                        cells.highlight = my_cells,
                        cols = "lightgrey",
                        cols.highlight = col_list[[i]],
                        sizes.highlight = 1.2 * min(4, pt_size),
                        pt.size = min(4, pt_size),
                        raster = FALSE) +
                    NoLegend() +
                    NoAxes() +
                    coord_fixed() +
                    ggtitle(unique(integrated@meta.data[["experiment"]])[i])
                )
    }
    for (i in seq_along(marker_df$gene)) {
      print(FeaturePlot(object = integrated,
                            features = marker_df$gene[[i]],
                            reduction = set_reduction,
                            pt.size = min(4, pt_size),
                            cols = c("lightsteelblue2", "black"),
                            order = TRUE,
                            raster = FALSE) +
                    NoAxes() +
                    coord_fixed() +
                    ggtitle(paste0(marker_df$celltype[[i]], "\n(", marker_df$gene[[i]], ")")) +
                    theme(plot.title = element_text(size = 10),
                          legend.text = element_text(size = 8),
                          legend.key.width = unit(.2, "cm")
                          )
              )
    }
}
integrated@reductions[["tsne"]] <- integrated@reductions[[paste0("tsne_", curr_pc, "pcs")]]
process_time <- stop_the_clock("plot_tsne")
```







## Save integrated data




## Save clustered data

```{r save-end}
start_time <- Sys.time()
if (run_number_4b == 1) {
    write_rds(x = integrated,
              file = paste0(objects_path,
                          dataset, "_",
                          input_type, "_",
                          normalization, "_",
                          integration, "_",
                          "integrated_tsne_embeddings_",
                          path_modifier,
                          ".rds")
              )
}
if (run_number_4b == 2) {
# for each string in "annotation_cell_type" vector, we need to pass the arguement into the script for subclustering
    write_rds(x = integrated,
        file = paste0(objects_path,
                      dataset, "_",
                      input_type, "_",
                      normalization, "_",
                      integration, "_",
                      "integrated_tsne_embeddings",
                      path_modifier, "_",
                      subcluster_celltype, "_",
                      "subclusting.rds")
              )
}
process_time <- stop_the_clock("save_end")
```







# Run time

```{r run-times}
process_time
```

```{r plot-run-times, fig.width = 10, fig.height = 3}
process_time$Code_chunk <- factor(process_time$Code_chunk, levels = rev(process_time$Code_chunk))
process_time %>%
    ggplot(aes(x = Code_chunk, y = Run_time, fill = Run_time)) +
    geom_col(width = 0.8) +
    xlab("Code Chunk") +
    ylab("Run Time (mins)") +
    coord_flip()
```



```{r stop-time}
total_end_time <- Sys.time()
total_end_time
total_end_time - total_start_time
```



# CPU and RAM Usage

```{r fig.width = 12, fig.height = 4}
job_id <- commandArgs(trailingOnly = TRUE)[2]
if (is.na(job_id)) {
    job_id <- "interactive"
}
if (job_id != "local" && job_id != "interactive") {
    job_log <- paste0("../sps-", job_id, "/sps-", job_id, "-cpu.tsv")
    if (file.exists(job_log)) {
        p1 <- read_tsv(file = job_log) %>%
          gather("Process", "Usage", -`#TIME`, -REQUESTED) %>%
          group_by(`#TIME`) %>%
          ggplot(aes(x = `#TIME`, y = Usage, colour = Process)) +
              geom_line() +
              ggtitle("CPU") +
              theme(legend.position = "none", validate = TRUE)
        p2 <- read_tsv(file = paste0("../sps-", job_id, "/sps-", job_id, "-mem.tsv")) %>%
          gather("Process", "Usage", -`#TIME`, -REQUESTED) %>%
          group_by(`#TIME`) %>%
          ggplot(aes(x = `#TIME`, y = Usage, colour = Process)) +
              geom_line() +
              ggtitle("Memory")
        my_legend <- get_legend(p2)
        p2 <- p2 + theme(legend.position = "none", validate = TRUE)
        plot_grid(p1, p2, my_legend, ncol = 3, rel_widths = c(3, 3, 2))
    }
}
```


# Session info

```{r session-info}
sessionInfo()
```
