---
title: "Clustering of scRNA-Seq data"
description: description
author: "Aaron M. Allen"
date: 'Last update: `r date()`'
output:
  html_document:
    number_sections: true
    code_folding: show
    code_download: true
    theme: cerulean
    df_print: paged
    fig_width: 8.5
    fig_height: 5
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 6
---



```{css, echo = FALSE}
pre, code {white-space:pre !important; overflow-x:auto}
.tocify-item {white-space:pre}
```



```{r echo = FALSE, cache = FALSE}

```


```{r setup, echo = FALSE, cache = FALSE}
ggplot2::theme_set(ggplot2::theme_grey())    # base_family = "Arial"))
options(scipen = 9999)

rmd_name <- gsub(pattern = ".Rmd", replacement = "", x = knitr::current_input(), ignore.case = TRUE)
knitr::opts_chunk$set(dev = c("png", "cairo_pdf"),
                      fig.align = "center",
                      fig.height = 5,
                      fig.width = 8.5,
                      dpi = 300,
                      pdf.options(encoding = "ISOLatin9.enc"),
                      fig.path = paste0("../analyses/figures/", rmd_name, "-",
                                        format(Sys.time(), "%Y%m%d_%H%M%S"), "/"),
                      fig.retina = 1,
                      warning = TRUE,
                      message = TRUE)
```



# Setup

## Job_ID = `r commandArgs(trailingOnly = TRUE)[2]`


## Start time

```{r start-time}
set.seed(12021)
today <- format(Sys.time(), "%Y%m%d_%H%M%S")
total_start_time <- Sys.time()
total_start_time
```






## Parameters



```{r params}
params_file <- commandArgs(trailingOnly = TRUE)[3]
if (is.na(params_file)) {
    params_file <- "pipeline_R_params.R"
}
params_path <- paste0("../src/param_files/", params_file)
print(paste0("The params file is :   ", params_path))
```

```{r, class.source = 'fold-hide', code = readLines(params_path)}
```

```{r run-number}
run_number_4c <- commandArgs(trailingOnly = TRUE)[4]
run_number_4c
```

Get the name of the machine that this is running on.  If running on my local desktop ("mentok"), then append to the `reference_object_path` variable.

```{r is-local}
local_path_mod <- ""
curr_machine <- Sys.info()["nodename"]
if (curr_machine == "mentok") {
    local_path_mod <- "/mnt/data2/scRNAseq/cbrg"
}
local_path_mod
```




## Libraries

```{r packages-etc, warning = FALSE, message = FALSE}
library(utils)
library(Seurat)
library(SingleCellExperiment)
library(tidyverse)
library(cowplot)
library(pheatmap)
library(colorspace)
library(ggcorrplot)
library(ggpubr)
library(corrr)
library(ape)        # used for phylogram during cluster collapse
library(future)
```




```{r run-time-tibble}
process_time <- tibble()
stop_the_clock <- function(chunk_name = "unnamed") {
    temp <- tibble::tibble(Code_chunk = chunk_name,
                           Start_time = start_time,
                           Stop_time = Sys.time(),
                           Run_time = difftime(Stop_time, Start_time, units = "mins")
                           )
    process_time <- dplyr::bind_rows(process_time, temp)
    return(process_time)
}
```



```{r parallel-options}
if (parallelize) {
    options(future.globals.maxSize = futures_ram * 1024 * 1024^2)
}
```



# The data

## Load data

```{r}
warning(paste0("Dataset to be loaded is: ", dataset))
```

```{r load-data}
start_time <- Sys.time()
if (remove_doublets == FALSE && remove_ambient == FALSE) {
    path_modifier <- ""
}
if (remove_doublets == TRUE && remove_ambient == FALSE) {
    path_modifier <- "_doublets_removed"
}
if (remove_doublets == FALSE && remove_ambient == TRUE) {
    path_modifier <- "_ambient_removed"
}
if (remove_doublets == TRUE && remove_ambient == TRUE) {
    path_modifier <- "_doublets_removed_ambient_removed"
}

if (run_number_4c == 1) {
    integrated <- read_rds(file = paste0(objects_path,
                                      dataset, "_",
                                      input_type, "_",
                                      normalization, "_",
                                      integration, "_",
                                      "integrated_seurat",
                                      path_modifier,
                                      ".rds")
                              )
}
if (run_number_4c == 2) {
# for each string in "annotation_cell_type" vector, we need to pass the arguement into the script for subclustering
    subcluster_celltype <- commandArgs(trailingOnly = TRUE)[5]
    print(paste0("Loading for subclustering: ", subcluster_celltype))
    integrated <- read_rds(file = paste0(objects_path,
                                      dataset, "_",
                                      input_type, "_",
                                      normalization, "_",
                                      integration, "_",
                                      "integrated_seurat",
                                      path_modifier, "_",
                                      subcluster_celltype, "_",
                                      "subclusting.rds")
    )
    curr_pc <- subcluster_pcs[which(annotation_cell_type == subcluster_celltype)]
    curr_res <- subcluster_res[which(annotation_cell_type == subcluster_celltype)]
}
integrated
process_time <- stop_the_clock("load_data")
```



```{r load-markers}
start_time <- Sys.time()
marker_df <- read_tsv(paste0(gmt_file_path), col_names = FALSE) %>%
    select(-X2) %>%
    gather("index", "gene", -X1) %>%
    select(-index) %>%
    rename(celltype = X1) %>%
    drop_na() %>%
    group_by(celltype) %>%
    filter(celltype %in% qc_plot_cell_type) %>%
    slice(1) %>%
    arrange(celltype)
marker_df
process_time <- stop_the_clock("load_markers")
```


Then filter marker gene to make sure they are in the object.

```{r genes-present}
all_genes <- rownames(integrated)
genes_not_found <- marker_df$gene[!marker_df$gene %in% all_genes]
if (length(genes_not_found) > 0) {
    warning(paste0("The following genes were not found: ", genes_not_found))
    marker_df <- marker_df %>% filter(gene != genes_not_found)
} else {
    message(paste0("All genes present in data."))
}
```








# Clustering

```{r cluster-setup}
start_time <- Sys.time()
if (integration == "harmony") {
    set_reduction <- "harmony"
} else {
    set_reduction <- "pca"
}
if (integration == "cca") {
    default_assay <- "integrated"
} else {
    if (normalization == "SCT") {
        default_assay <- "SCT"
    }
    if (normalization == "LogNormalize") {
        default_assay <- "RNA"
    }
}
graph_name <- paste0(default_assay, "_snn")

print(paste0("Normalization is ", normalization))
print(paste0("Integration is ", integration))
print(paste0("Graph name is ", graph_name))


DefaultAssay(object = integrated) <- default_assay
DefaultAssay(object = integrated)
```


```{r neighbours}
if (parallelize) {
    # plan(futures_plan, workers = futures_n_cores)
    options(future.globals.maxSize = futures_ram * 1024 * 1024^2)
}
gc()
integrated <- FindNeighbors(integrated,
                           reduction = set_reduction,
                           dims = 1:curr_pc,
                           k.param = neighbours,
                           # nn.method = nn_method,
                           # nn.eps = nn_eps,
                           # n.trees = nn_trees,
                           force.recalc = TRUE,
                           graph.name = graph_name,
                           verbose = TRUE)
gc()
```


```{r clustering}
gc()
integrated <- FindClusters(integrated,
                          resolution = resolution_test,
                          graph.name = graph_name,
                          method = leiden_graph_method,
                          algorithm = clustering_algorithm,
                          group.singletons = group_singletons,
                          verbose = TRUE)
gc()

### note: leiden clusterings start at 1, but louvain starts at 0
###       to avoid issues with the cluster collapse function I'm decrementing the Idents by 1 if leiden is run
if (clustering_algorithm == 4) {
  clusters <- names(integrated@meta.data) %>% str_subset(graph_name)
  for (my_cluster in clusters) {
      integrated@meta.data[[my_cluster]] <- as.factor(as.numeric(integrated@meta.data[[my_cluster]]) - 1)
  }
}

if (parallelize) {
    plan("sequential")
}
process_time <- stop_the_clock("clustering")
```


```{r plot-num-clusters-func}
plot_number_of_clusters <- function(.data) { # nolint
    p <- .data %>%
            tibble::rownames_to_column("cell") %>%
            tidyr::gather(resolution, cluster, -cell) %>%
            dplyr::mutate(cluster = as.double(cluster),
                resolution = as.numeric(sub("^[^_]*_snn_res.", "", resolution))) %>%
            dplyr::group_by(resolution) %>%
            dplyr::summarize(ncluster = max(cluster, na.rm = TRUE)) %>%
            dplyr::arrange(resolution) %>%
            ggplot2::ggplot(aes(x = as.factor(resolution), y = ncluster)) +
                ggplot2::geom_col() +
                ggplot2::labs(x = "Resolution", y = "Number of clusters")
    return(p)
}
```


## Number of clusters per resolution

```{r plot-num-clusters, warning = FALSE, message = FALSE}
integrated@meta.data[, grep("*_snn", names(integrated@meta.data))] %>% plot_number_of_clusters
```



## DimPlot cluster resolutions

```{r plot-clusters, fig.width = 12, fig.height = 12}
start_time <- Sys.time()
ident_meta_positions <- grep("*_snn", names(integrated@meta.data))
pt_size <- 4 * AutoPointSize(data = integrated, raster = FALSE)

for (i in seq_along(ident_meta_positions)) {
    Idents(integrated) <- integrated@meta.data[, ident_meta_positions[[i]]]
    print(DimPlot(integrated,
                  reduction = paste0("umap_", curr_pc, "pcs"),
                  label = TRUE,
                  pt.size = min(4, pt_size),
                  raster = FALSE) +
            NoAxes() +
            NoLegend() +
            coord_fixed() +
            theme(plot.title = element_text(size = 10, hjust = .5)) +
            ggtitle(paste("Resolution = ", resolution_test[[i]]))
        )
}
process_time <- stop_the_clock("plot_clusters")
```




## Plot single res

```{r plot-clustering-dimplot, fig.asp = 1}
start_time <- Sys.time()
res_name <- paste0(graph_name, "_res.", curr_res)

Idents(integrated) <- integrated$seurat_clusters <- integrated@meta.data[[res_name]]
DimPlot(integrated, reduction = paste0("umap_", curr_pc, "pcs"), label = TRUE, pt.size = min(4, pt_size), raster = FALSE) +
  NoAxes() +
  NoLegend() +
  coord_fixed() +
  ggtitle(paste0("Tissue = ", dataset),
          subtitle = paste0("PC = ", curr_pc,
                            ", Resolution = ", curr_res,
                            ", Normalization = ", normalization,
                            ", Integration = ", integration
                            )
          )
process_time <- stop_the_clock("plot_clustering_dimplot")
```


```{r clust-contrib, fig.width = 12, fig.height = 6}
start_time <- Sys.time()
tibble(clust = integrated@active.ident,
           samp = integrated$experiment) %>%
    ggplot(aes(x = clust, fill = samp)) +
        geom_bar(position = "fill", width = 1) +
        ggtitle("Cluster contribution") +
        theme_void() +
        theme(axis.text.x = element_blank(),
              axis.ticks.x = element_blank(),
              legend.position = "bottom")
process_time <- stop_the_clock("clust_contrib")
```




## Filter QC on UMAP


```{r qc-cluster, fig.height = 12, fig.width = 10}
start_time <- Sys.time()
p_list <- list()
for (i in seq_along(qc_params_plot)) {
    p_list[[i]] <- FeaturePlot(object = integrated,
            features = qc_params_plot[[i]],
            max.cutoff = qc_params_plot_thesh[[i]],
            reduction = paste0("umap_", curr_pc, "pcs"),
            cols = c("lightsteelblue2", "black"),
            raster = FALSE) +
    NoAxes() +
    coord_fixed()
}
plot_grid(plotlist = p_list, ncol = 2)
process_time <- stop_the_clock("qc_cluster")
```



## Number of cells per cluster

```{r cells-per-cluster, fig.height = 4, fig.width = 10}
start_time <- Sys.time()
as.data.frame(table(Idents(integrated))) %>%
  as_tibble() %>%
  left_join(tibble(nUMI = integrated$nCount_RNA, cluster = Idents(integrated)) %>%
  dplyr::group_by(cluster) %>%
  dplyr::summarize(mean(nUMI)), by = c("Var1" = "cluster")) %>%
  arrange(Freq) %>%
  mutate(Var1 = factor(Var1, levels = Var1)) %>%
  ggplot(aes(x = Var1, y = Freq, color = `mean(nUMI)`, fill  = `mean(nUMI)`)) +
    geom_col(width = 0.4, size = 2) +
    # scale_y_log10() +
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
    labs(x = "Clusters", y = "Number of cells")
process_time <- stop_the_clock("cells_per_cluster")
```

## Male cells

```{r roX1-expression, fig.asp = 1}
start_time <- Sys.time()
if (normalization == "SCT") {
    default_assay <- "SCT"
} else {
    default_assay <- "RNA"
}
DefaultAssay(object = integrated) <- default_assay
temp_split <- SplitObject(object = integrated, split.by = "experiment")
for (i in seq_along(temp_split)) {
    print(FeaturePlot(object = temp_split[[i]],
                    features = "lncRNA:roX1",
                    order = TRUE,
                    reduction = paste0("umap_", curr_pc, "pcs"),
                    cols = c("lightsteelblue2", "black"),
                    raster = FALSE) +
              ggtitle(unique(temp_split[[i]]$experiment)) +
        coord_fixed()
        )
}
process_time <- stop_the_clock("roX1_expression")
```





## Save clustered data


```{r make-dir}
dir.create(paste0(objects_path, "../markers/"), recursive = TRUE, showWarnings = FALSE)
```


```{r save-end}
start_time <- Sys.time()
if (run_number_4c == 1) {
    write_rds(x = integrated,
              file = paste0(objects_path,
                          dataset, "_",
                          input_type, "_",
                          normalization, "_",
                          integration, "_",
                          "integrated_seurat",
                          path_modifier,
                          ".rds")
              )
}
if (run_number_4c == 2) {
# for each string in "annotation_cell_type" vector, we need to pass the arguement into the script for subclustering
    write_rds(x = integrated,
        file = paste0(objects_path,
                      dataset, "_",
                      input_type, "_",
                      normalization, "_",
                      integration, "_",
                      "integrated_seurat",
                      path_modifier, "_",
                      subcluster_celltype, "_",
                      "subclusting.rds")
              )
}
process_time <- stop_the_clock("save_end")
```







# Run time

```{r run-times}
process_time
```

```{r plot-run-times, fig.width = 10, fig.height = 5}
process_time$Code_chunk <- factor(process_time$Code_chunk, levels = rev(process_time$Code_chunk))
process_time %>%
    ggplot(aes(x = Code_chunk, y = Run_time, fill = Run_time)) +
    geom_col(width = 0.8) +
    xlab("Code Chunk") +
    ylab("Run Time (mins)") +
    coord_flip()
```



```{r stop-time}
total_end_time <- Sys.time()
total_end_time
total_end_time - total_start_time
```



# CPU and RAM Usage

```{r fig.width = 12, fig.height = 4}
job_id <- commandArgs(trailingOnly = TRUE)[2]
if (is.na(job_id)) {
    job_id <- "interactive"
}
if (job_id != "local" && job_id != "interactive") {
    job_log <- paste0("../sps-", job_id, "/sps-", job_id, "-cpu.tsv")
    if (file.exists(job_log)) {
        p1 <- read_tsv(file = job_log) %>%
          gather("Process", "Usage", -`#TIME`, -REQUESTED) %>%
          group_by(`#TIME`) %>%
          ggplot(aes(x = `#TIME`, y = Usage, colour = Process)) +
              geom_line() +
              ggtitle("CPU") +
              theme(legend.position = "none", validate = TRUE)
        p2 <- read_tsv(file = paste0("../sps-", job_id, "/sps-", job_id, "-mem.tsv")) %>%
          gather("Process", "Usage", -`#TIME`, -REQUESTED) %>%
          group_by(`#TIME`) %>%
          ggplot(aes(x = `#TIME`, y = Usage, colour = Process)) +
              geom_line() +
              ggtitle("Memory")
        my_legend <- get_legend(p2)
        p2 <- p2 + theme(legend.position = "none", validate = TRUE)
        plot_grid(p1, p2, my_legend, ncol = 3, rel_widths = c(3, 3, 2))
    }
}
```


# Session info

```{r session-info}
sessionInfo()
```
