---
title: "Annotate Broad Cell Types"
description:
author: "Aaron M. Allen"
date: 'Last update: `r date()`'
output:
  html_document:
    number_sections: true
    code_folding: show
    code_download: true
    theme: cerulean
    df_print: paged
    fig_width: 8.5
    fig_height: 5
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 6
---



```{css, echo = FALSE}
pre, code {white-space:pre !important; overflow-x:auto}
.tocify-item {white-space:pre}
```



```{r echo = FALSE, cache = FALSE}

```


```{r setup, echo = FALSE, cache = FALSE}
ggplot2::theme_set(ggplot2::theme_grey())

rmd_name <- gsub(pattern = ".Rmd", replacement = "", x = knitr::current_input(), ignore.case = TRUE)
knitr::opts_chunk$set(dev = c("png", "cairo_pdf"),
                      fig.align = "center",
                      fig.height = 5,
                      fig.width = 8.5,
                      dpi = 300,
                      pdf.options(encoding = "ISOLatin9.enc"),
                      fig.path = paste0("../analyses/figures/", rmd_name, "-",
                                        format(Sys.time(), "%Y%m%d_%H%M%S"), "/"),
                      fig.retina = 1,
                      warning = TRUE,
                      message = TRUE)
```



# Setup

## Job_ID = `r commandArgs(trailingOnly = TRUE)[2]`


## Start time

```{r start-time}
set.seed(12021)
today <- format(Sys.time(), "%Y%m%d_%H%M%S")
total_start_time <- Sys.time()
total_start_time
```






## Parameters



```{r params}
params_file <- commandArgs(trailingOnly = TRUE)[3]
params_path <- paste0("../src/param_files/", params_file)
print(paste0("The params file is :   ", params_path))
```

```{r, class.source = 'fold-hide', code = readLines(params_path)}
```





## Libraries


```{r load-libraries}
library(tidyverse)
library(Seurat)
library(pheatmap)
library(colorspace)
library(cowplot)
```




## other bits

```{r run-time-tibble}
process_time <- tibble()
stop_the_clock <- function(chunk_name = "unnamed") {
    temp <- tibble::tibble(Code_chunk = chunk_name,
                           Start_time = start_time,
                           Stop_time = Sys.time(),
                           Run_time = difftime(Stop_time, Start_time, units = "mins")
                           )
    process_time <- dplyr::bind_rows(process_time, temp)
    return(process_time)
}
```





## Data

```{r load-data}
start_time <- Sys.time()
if (remove_doublets == FALSE && remove_ambient == FALSE) {
    path_modifier <- ""
}
if (remove_doublets == TRUE && remove_ambient == FALSE) {
    path_modifier <- "_doublets_removed"
}
if (remove_doublets == FALSE && remove_ambient == TRUE) {
    path_modifier <- "_ambient_removed"
}
if (remove_doublets == TRUE && remove_ambient == TRUE) {
    path_modifier <- "_doublets_removed_ambient_removed"
}



core_object_path <- paste0(objects_path,
                                  dataset, "_",
                                  input_type, "_",
                                  normalization, "_",
                                  integration, "_")
diet_object <- paste0(core_object_path, "integrated_seurat", path_modifier, "_diet.rds")
regular_object <- paste0(core_object_path, "integrated_seurat", path_modifier, ".rds")
if (file.exists(diet_object)) {
   object_to_load <- diet_object
} else {
    object_to_load <- regular_object
}
integrated <- read_rds(file = object_to_load)
pt_size <- 4 * AutoPointSize(data = integrated, raster = FALSE)
integrated


process_time <- stop_the_clock("load_data")
```



# Define Markers



## Cell types and Markers

```{r load-markers}
start_time <- Sys.time()
marker_df <- read_tsv(gmt_file_path, col_names = FALSE) %>%
    select(-X2) %>%
    gather("index", "gene", -X1) %>%
    select(-index) %>%
    rename(celltype = X1) %>%
    drop_na() %>%
    group_by(celltype) %>%
    arrange(celltype)
marker_df
process_time <- stop_the_clock("load_markers")
```


```{r genes-present}
all_genes <- integrated@assays[["RNA"]]@counts@Dimnames[[1]]
genes_not_found <- marker_df %>%
    filter(!gene %in% all_genes) %>%
    pull(gene)
if (length(genes_not_found) > 0) {
    warning(paste0("The following genes were not found: ", genes_not_found))
    marker_df <- marker_df %>%
        filter(gene %in% all_genes)
} else {
    message(paste0("All genes present in data."))
}
```




## AddModuleScore


```{r add-modulescore}
start_time <- Sys.time()

if (integration == "cca") {
    assay_use <- "integrated"
} else if (normalization == "SCT") {
    assay_use <- "SCT"
} else {
    assay_use <- "RNA"
}

for (my_celltype in unique(marker_df$celltype)) {
    markers <- marker_df %>% filter(celltype == my_celltype) %>% pull(gene)
    integrated <- AddModuleScore(object = integrated,
                                 features = list(markers),
                                 name = paste0("modulescore_", my_celltype),
                                 assay = assay_use,
                                 ctrl = n_crt_features,
                                 seed = 12021)
}
process_time <- stop_the_clock("add_modulescore")
```



## Normalize Module Score

To compare across module scores, we take the l2-norm of all the module scores.

```{r l2norm}
start_time <- Sys.time()
l2norm <- function(x) {
    x / sqrt(sum(x^2))
}
module_scores <- colnames(integrated@meta.data) %>% str_subset("modulescore") %>% str_subset("_l2norm", negate = TRUE)
for (i in seq_along(module_scores)) {
    if (sum(integrated@meta.data[[module_scores[[i]]]]) != 0) {
        integrated <- AddMetaData(object = integrated,
                                  metadata = l2norm(integrated@meta.data[[module_scores[[i]]]]),
                                  col.name = paste0(module_scores[[i]], "_l2norm"))
    } else {
        integrated <- AddMetaData(object = integrated,
                              metadata = 0,
                              col.name = paste0(module_scores[[i]], "_l2norm"))
    }
}
process_time <- stop_the_clock("l2norm")
```




## Threshold Module Score

```{r kmeans-threshold}
start_time <- Sys.time()
k_means_thresholds_list <- list()
for (my_celltype in unique(marker_df$celltype)) {
    module_scores_norm <- paste0("modulescore_", my_celltype, "1_l2norm")
    temp <- data.frame(ModuleScore = integrated@meta.data[[module_scores_norm]])
    tryCatch({
        km <- kmeans(temp$ModuleScore, centers = my_k, iter.max = my_k_itr)
        temp$clust <- as.factor(km$cluster)
        k_thresh <- temp %>%
                    group_by(clust) %>%
                    summarise(min = min(ModuleScore),
                              max = max(ModuleScore)) %>%
                    gather("bound", "value", -clust) %>%
                    # ungroup() %>%
                    arrange(value) %>%
                    pull(value) %>%
                    .[my_k * 2 - 1]
    } , error = function(e) {
            message(paste0(e))
            message("Setting k_thresh to 0 ... \n\n")
            k_thresh <<- 0
    })
    k_means_thresholds_list[[my_celltype]] <- tibble(Celltype = my_celltype,
                                                     K_thresh = k_thresh)
}
k_means_thresholds <- bind_rows(k_means_thresholds_list)
k_means_thresholds
process_time <- stop_the_clock("kmeans_threshold")
```


## Plot Module Scores

### Define plot function

```{r plot-function}
start_time <- Sys.time()
plot_module_thresh_kmeans <- function(object, celltype, thresh_table) {
    my_module <- paste0("modulescore_", celltype, "1")
    my_module_norm <- paste0(my_module, "_l2norm")

    p1 <- Seurat::FeaturePlot(object = object,
               features = my_module,
               reduction = paste0("umap_", curr_pc, "pcs"),
               cols = c("lightsteelblue2", "black"),
               pt.size = min(4, pt_size),
               order = TRUE,
               raster = FALSE
               ) +
      Seurat::NoAxes() +
      ggplot2::coord_fixed()
    p2 <- Seurat::FeaturePlot(object = object,
               features = my_module_norm,
               reduction = paste0("umap_", curr_pc, "pcs"),
               cols = c("lightsteelblue2", "black"),
               pt.size = min(4, pt_size),
               order = TRUE,
               raster = FALSE
               ) +
      Seurat::NoAxes() +
      ggplot2::coord_fixed()

    k_thresh <- thresh_table %>%
        filter(celltype == Celltype) %>%
        dplyr::pull(K_thresh)

    my_label <- paste0("\nk-", my_k, " slpit = ", round(x = k_thresh, digits = 4), "    ")
    p3 <- data.frame(ModuleScore = object@meta.data[[my_module_norm]]) %>%
            ggplot2::ggplot(aes(x = ModuleScore)) +
                ggplot2::geom_histogram(aes(y = 100 * ..count.. / sum(..count..)),
                             fill = "lightsteelblue2",
                             position = "stack",
                             bins = 60,
                             show.legend = FALSE) +
                ggplot2::geom_vline(aes(xintercept = k_thresh), linetype = "dotted", size = 1, colour = "black") +
                ggplot2::ylab("Percent") +
                ggplot2::annotate("text", x = Inf, y = Inf, label = my_label, vjust = 1, hjust = 1, size = 4) +
                ggplot2::theme(plot.margin = ggplot2::unit(c(30, 30, 30, 30), "pt"))

    cells_highlight <- as.data.frame(integrated@meta.data) %>%
                            tibble::rownames_to_column("cellid") %>%
                            dplyr::select(cellid, my_module_norm) %>%
                            dplyr::filter(.[2] > k_thresh &  .[2] > 0) %>%
                            dplyr::pull(cellid)
    p4 <- Seurat::DimPlot(object = integrated,
            reduction = paste0("umap_", curr_pc, "pcs"),
            pt.size = min(4, pt_size),
            label = FALSE,
            cells.highlight = cells_highlight,
            cols.highlight = "black",
            sizes.highlight = 0.1,
            cols = "lightsteelblue2",
            na.value = "lightsteelblue2",
            raster = FALSE
            ) +
        ggplot2::ggtitle(paste0(celltype, "?")) +
        Seurat::NoAxes() +
        Seurat::NoLegend() +
        ggplot2::coord_fixed()

    p_all <- plot_grid(p1, p2, p3, p4, ncol = 2)
    title <- ggdraw() + draw_label(celltype, fontface = "bold", x = 0, hjust = 0, vjust = -0.1, size = 20)
    plot_grid(title, p_all, ncol = 1, rel_heights = c(0.1, 1))
}
process_time <- stop_the_clock("plot_function")
```



### Plot the module scores

```{r plot-modulescores, fig.width = 10, fig.height = 10}
start_time <- Sys.time()
for (my_celltype in unique(marker_df$celltype)) {
    cat(paste0("\n", my_celltype))
    print(plot_module_thresh_kmeans(object = integrated, celltype = my_celltype, thresh_table = k_means_thresholds))
}
process_time <- stop_the_clock("plot_modulescores")
```




## Plot Clusters

For these course annotations, we want to keep the cluster resolution low and keep the cell types broad.

```{r plot-curr-res-label, fig.width = 10, fig.height = 10}
start_time <- Sys.time()

if (integration == "cca") {
    default_assay <- "integrated"
} else {
    if (normalization == "SCT") {
        default_assay <- "SCT"
    }
    if (normalization == "LogNormalize") {
        default_assay <- "RNA"
    }
}
graph_name <- paste0(default_assay, "_snn")

curr_res_label <- paste0(graph_name, "_res.", curr_res)
DefaultAssay(object = integrated) <- default_assay
DefaultAssay(object = integrated)


DimPlot(object = integrated,
        reduction = paste0("umap_", curr_pc, "pcs"),
        group.by = curr_res_label,
        pt.size = min(4, pt_size),
        raster = FALSE,
        label = TRUE,
        label.box = TRUE,
        repel = FALSE) +
    NoAxes() +
    NoLegend() +
    coord_fixed()
process_time <- stop_the_clock("plot_curr_res_label")
```



# Annotate Cells


## Max Module Score

### Extract Module Scores

```{r extract-module-scores}
start_time <- Sys.time()
old_colnames <- colnames(as.data.frame(integrated@meta.data)) %>% str_subset("_l2norm")
new_colnames <- old_colnames %>% str_remove("modulescore_") %>% str_remove("1_l2norm")

module_score_df <- as.data.frame(integrated@meta.data) %>%
    select(contains("_l2norm"))
colnames(module_score_df) <- new_colnames
head(module_score_df)
process_time <- stop_the_clock("extract_module_scores")
```

### Annotate

```{r annotate-by-cell}
start_time <- Sys.time()
annotation_broad_df <- module_score_df %>%
    select(annotation_cell_type)
top_annotation_broad <- colnames(annotation_broad_df)[max.col(annotation_broad_df, ties.method = "first")]
my_colnames <- colnames(annotation_broad_df)
annotation_broad_df <- annotation_broad_df %>%
    mutate(max_value = pmax(!!!rlang::syms(my_colnames)),
           broad = top_annotation_broad
           ) %>%
    rownames_to_column("cell_id")

if (manual_thresh) {
    k_means_thresholds <- tibble(CellType = annotation_cell_type,
                                 K_thresh = manual_thresholds)
}
annotation_broad_df <- annotation_broad_df %>%
    left_join(k_means_thresholds, by = c("broad" = "Celltype"))

if (annotate_by_K) {
    annotation_broad_df <- annotation_broad_df %>%
        mutate(broad = if_else(max_value > K_thresh & max_value > 0, broad, "unknown"))
} else {
    annotation_broad_df <- annotation_broad_df %>%
        mutate(broad = if_else(max_value > 0, broad, "unknown"))
}
annotation_broad_df[annotation_broad_df == "unknown"] <- NA
head(annotation_broad_df)
process_time <- stop_the_clock("annotate_by_cell")
```



## Plot Annotations


```{r plot-annotation, fig.width = 10, fig.height = 10}
start_time <- Sys.time()
integrated <- AddMetaData(integrated, metadata = annotation_broad_df$broad, col.name = "annotation_broad_cell_type")
DimPlot(object = integrated,
        reduction = paste0("umap_", curr_pc, "pcs"),
        group.by = "annotation_broad_cell_type",
        pt.size = min(4, pt_size),
        raster = FALSE) +
    NoAxes() +
    coord_fixed()
process_time <- stop_the_clock("plot_annotation")
```


## Cluster contributions

Look at the relative cluster-wise contributions of the annotations.

```{r fig.width = 10, fig.height = 5}
tibble(clust = integrated@meta.data[[curr_res_label]],
        annotation = integrated@meta.data[["annotation_broad_cell_type"]]) %>%
        ggplot(aes(x = clust, fill = annotation)) +
            geom_bar(position = "fill", width = 1, colour = "white", size = 0.4) +
            # scale_fill_manual(values = colorspace::qualitative_hcl(n = 7,palette = "Set 2"),na.value = "azure3") +
            # scale_fill_manual(values = scales::hue_pal()(length(unique(integrated$annotation_broad_cell_type))-1),
            #                   na.value = "grey80") +
            scale_y_continuous(labels = scales::percent) +
            ggtitle(paste0("Annotation cluster contribution")) +
            labs(x = "Clusters", y = "Percent of Cells") +
            theme(#axis.text.x = element_blank(),
                  axis.ticks.x = element_blank(),
                  legend.title = element_blank(),
                  legend.position = "bottom",
                  plot.title = element_text(face = "bold"))
```


### Donut Chart

```{r plot-ann-pie, fig.width = 6, fig.height = 6}
ann_use <- "annotation_broad_cell_type"
tibble(Annotations = integrated@meta.data[[ann_use]]) %>%
    group_by(Annotations) %>%
    dplyr::summarise(count = n()) %>%
    ggplot(aes(x = 1, y = count, fill = Annotations)) +
        geom_bar(width = 1, stat = "identity", colour = "white", size = 1) +
        coord_polar("y", direction = 1) +
        xlim(c(-0.5, 1.5)) +
        ggtitle(ann_use) +
        theme_void() +
        theme(legend.title = element_blank(),
            axis.title = element_blank(),
            axis.ticks = element_blank(),
            plot.title = element_text(face = "bold"))
```






## Extrapolating most common

### Find most common annotation

```{r get-top-ann}
start_time <- Sys.time()
cluster_level_ann <- as.data.frame(integrated@meta.data) %>%
    select(contains(curr_res_label), annotation_broad_cell_type) %>%
    group_by(.[1]) %>%
    table() %>%
    as.data.frame() %>%
    group_by(.[1]) %>%
    arrange(.[1], desc(Freq)) %>%
    mutate(Perc = 100 * Freq / sum(Freq)) %>%
    slice(1)
head(cluster_level_ann)
process_time <- stop_the_clock("get_top_ann")
```



### Get labels per cell

```{r top-ann-by-cell}
start_time <- Sys.time()
cluster_level_ann_by_cell <- as.data.frame(integrated@meta.data) %>%
    rownames_to_column("cell_id") %>%
    select(cell_id, contains(curr_res_label)) %>%
    mutate(RNA_snn_res.20 = as.factor(RNA_snn_res.20)) %>%
    left_join(cluster_level_ann)
head(cluster_level_ann_by_cell)
process_time <- stop_the_clock("top_ann_by_cell")
```



### Annotate and Plot

```{r plot-annotation-ext, fig.width = 10, fig.height = 10}
start_time <- Sys.time()
integrated <- AddMetaData(object = integrated,
                          metadata = cluster_level_ann_by_cell$annotation_broad_cell_type,
                          col.name = "annotation_broad_cell_type_ext")
DimPlot(object = integrated,
        reduction = paste0("umap_", curr_pc, "pcs"),
        group.by = "annotation_broad_cell_type_ext",
        pt.size = min(4, pt_size),
        raster = FALSE) +
    NoAxes() +
    coord_fixed()
process_time <- stop_the_clock("plot_annotation_ext")
```


### Donut chart

```{r plot-ann-pie-3, fig.width = 6, fig.height = 6}
ann_use <- "annotation_broad_cell_type_ext"
tibble(Annotations = integrated@meta.data[[ann_use]]) %>%
    group_by(Annotations) %>%
    dplyr::summarise(count = n()) %>%
    ggplot(aes(x = 1, y = count, fill = Annotations)) +
        geom_bar(width = 1, stat = "identity", colour = "white", size = 1) +
        coord_polar("y", direction = 1) +
        xlim(c(-0.5, 1.5)) +
        ggtitle(ann_use) +
        theme_void() +
        theme(legend.title = element_blank(),
            axis.title = element_blank(),
            axis.ticks = element_blank(),
            plot.title = element_text(face = "bold"))
```



# Save data

```{r save-data}
start_time <- Sys.time()
write_rds(x = integrated, file = object_to_load)
process_time <- stop_the_clock("save_data")
```


```{r}
modulescores <- colnames(integrated@meta.data) %>% str_subset("modulescore")
annotations_metadata <- FetchData(object = integrated, 
                                  vars = c("experiment",
                                           all_of(modulescores),
                                           "annotation_broad_cell_type",
                                           "annotation_broad_cell_type_ext")
                                  ) %>% 
    rownames_to_column("cell_id")

write_csv(x = annotations_metadata,
          file = paste0(objects_path,
                        dataset, "_",
                        input_type, "_",
                        normalization, "_",
                        integration, "_",
                        "integrated_seurat", 
                        path_modifier,
                        "_annotations.csv")
          )
```








# Run time

```{r run-times}
process_time
```

```{r plot-run-times, fig.width = 10, fig.height = 8}
process_time$Code_chunk <- factor(process_time$Code_chunk, levels = rev(process_time$Code_chunk))
process_time %>%
    ggplot(aes(x = Code_chunk, y = Run_time, fill = Run_time)) +
    geom_col(width = 0.8) +
    xlab("Code Chunk") +
    ylab("Run Time (mins)") +
    coord_flip()
```



```{r stop-time}
total_end_time <- Sys.time()
total_end_time
total_end_time - total_start_time
```



# CPU and RAM Usage

```{r fig.width = 12, fig.height = 4}
job_id <- commandArgs(trailingOnly = TRUE)[2]
if (job_id != "local") {
    job_log <- paste0("../sps-", job_id, "/sps-", job_id, "-cpu.tsv")
    if (file.exists(job_log)) {
        p1 <- read_tsv(file = job_log) %>%
          gather("Process", "Usage", -`#TIME`, -REQUESTED) %>%
          group_by(`#TIME`) %>%
          ggplot(aes(x = `#TIME`, y = Usage, colour = Process)) +
              geom_line() +
              ggtitle("CPU") +
              theme(legend.position = "none", validate = TRUE)
        p2 <- read_tsv(file = paste0("../sps-", job_id, "/sps-", job_id, "-mem.tsv")) %>%
          gather("Process", "Usage", -`#TIME`, -REQUESTED) %>%
          group_by(`#TIME`) %>%
          ggplot(aes(x = `#TIME`, y = Usage, colour = Process)) +
              geom_line() +
              ggtitle("Memory")
        my_legend <- get_legend(p2)
        p2 <- p2 + theme(legend.position = "none", validate = TRUE)
        plot_grid(p1, p2, my_legend, ncol = 3, rel_widths = c(3, 3, 2))
    }
}
```


# Session info

```{r session-info}
sessionInfo()
```
