---
title: "Correlate from bulk"
description:
author: "Aaron M. Allen"
date: 'Last update: `r date()`'
output:
  html_document:
    number_sections: true
    code_folding: show
    code_download: true
    theme: cerulean
    df_print: paged
    fig_width: 8.5
    fig_height: 5
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 6
---



```{css, echo = FALSE}
pre, code {white-space:pre !important; overflow-x:auto}
.tocify-item {white-space:pre}
```



```{r echo = FALSE, cache = FALSE}

```


```{r setup, echo = FALSE, cache = FALSE}
ggplot2::theme_set(ggplot2::theme_grey())

rmd_name <- gsub(pattern = ".Rmd", replacement = "", x = knitr::current_input(), ignore.case = TRUE)
knitr::opts_chunk$set(dev = c("png", "cairo_pdf"),
                      fig.align = "center",
                      fig.height = 5,
                      fig.width = 8.5,
                      dpi = 300,
                      pdf.options(encoding = "ISOLatin9.enc"),
                      fig.path = paste0("../analyses/figures/", rmd_name, "-",
                                        format(Sys.time(), "%Y%m%d_%H%M%S"), "/"),
                      fig.retina = 1,
                      warning = TRUE,
                      message = TRUE)
```



# Setup

## Job_ID = `r commandArgs(trailingOnly = TRUE)[2]`


## Start time

```{r start-time}
set.seed(12021)
today <- format(Sys.time(), "%Y%m%d_%H%M%S")
total_start_time <- Sys.time()
total_start_time
```






## Parameters



```{r params}
params_file <- commandArgs(trailingOnly = TRUE)[3]
params_path <- paste0("../src/param_files/", params_file)
print(paste0("The params file is :   ", params_path))
```

```{r, class.source = 'fold-hide', code = readLines(params_path)}
```


Get the name of the machine that this is running on.  If running on my local desktop ("mentok"), then append to the `raw_path` variable.

```{r is-local}
local_path_mod <- ""
curr_machine <- Sys.info()["nodename"]
if (curr_machine == "mentok") {
    local_path_mod <- "/mnt/data2/scRNAseq/cbrg"
}
local_path_mod
```






## Libraries

```{r packages-etc, warning = FALSE, message = FALSE}
library(Seurat)
library(tidyverse)
library(cowplot)
library(pheatmap)
library(colorspace)
library(ggrepel)
library(corrr)
library(readxl)
```



## The Functions


```{r}
top_n_avg_exp <- function(object, markers_df, ident, n = 10, assay_use = "RNA", slot_use = "data", ...) { # nolint
    Seurat::Idents(object) <- object@meta.data[[ident]]
    top_n_markers <- markers_df %>%
        dplyr::filter(p_val_adj < 0.05) %>%
        dplyr::group_by(cluster) %>%
        dplyr::top_n(n = n, wt = avg_log2FC) %>%
        dplyr::pull(gene) %>%
        unique()
    avg_exp_by_cluster <- Seurat::AverageExpression(object = object,
                                            assays = assay_use,
                                            slot = slot_use,
                                            features = top_n_markers)
    return(avg_exp_by_cluster)
}
```


```{r}
corr_bulk_sc <- function(avg_exp_mat, bulk_df, assay_use = "RNA", ...) { # nolint
    slim_corr <- avg_exp_mat[[assay_use]] %>%
        as.data.frame() %>%
        tibble::rownames_to_column("gene") %>%
        dplyr::left_join(bulk_df, by = "gene") %>%
        tibble::column_to_rownames("gene") %>%
        corrr::correlate(method = "spearman") %>%
        corrr::shave() %>%
        corrr::stretch() %>%
        dplyr::filter(y %in% colnames(bulk_df) &
               !x %in% colnames(bulk_df))
    return(slim_corr)
}
```


```{r}
plot_corr_gap <- function(object, ident, corr_df, cell_type, curr_pc, y_limits = NULL, ...) { # nolint
    Seurat::Idents(object) <- object@meta.data[[ident]]
    p1 <- corr_df %>%
        dplyr::filter(y == cell_type) %>%
        dplyr::mutate(x = forcats::fct_reorder(x, dplyr::desc(r))) %>%
        ggplot2::ggplot(ggplot2::aes(x = x, y = r)) +
            ggplot2::geom_point() +
            ggplot2::expand_limits(x = c(-5, length(levels(corr_df$x)) + 5)) +
            ggplot2::coord_cartesian(clip = "off") +
            ggplot2::ggtitle(cell_type) +
            ggplot2::ylab("Spearman Correlation") +
            ggplot2::xlab("Clusters") +
            ggplot2::theme_classic() +
            ggplot2::theme(
                plot.title = ggplot2::element_text(size = 24, face = "bold.italic"),
                axis.title.x = ggplot2::element_text(size = 14, face = "bold"),
                axis.title.y = ggplot2::element_text(size = 14, face = "bold"),
                axis.text.x = ggplot2::element_blank(),
                axis.ticks.x = ggplot2::element_blank()) +
            ggrepel::geom_label_repel(aes(label = x),
                  box.padding   = 0.35,
                  point.padding = 0.5,
                  segment.color = "grey50")
    if (!is.null(y_limits)) {
        p1 <- p1 + ggplot2::ylim(y_limits)
    }
    cluster_highlight <- corr_df %>%
        dplyr::filter(y == cell_type) %>%
        dplyr::top_n(n = 1, wt = r) %>%
        dplyr::pull(x)
    my_cells <- Seurat::WhichCells(object = object, idents = cluster_highlight)
    p2 <- Seurat::DimPlot(object = object,
                  cells.highlight = my_cells,
                  cols = "lightsteelblue2",
                  cols.highlight = "black",
                  sizes.highlight = 0.4,
                  pt.size = 0.01,
                  raster = FALSE,
                  reduction = paste0("tsne_", curr_pc, "pcs")) +
        ggplot2::ggtitle(paste0("Cluster = ", cluster_highlight)) +
        Seurat::NoLegend() +
        ggplot2::coord_fixed()
    p3 <- cowplot::plot_grid(p1, p2, ncol = 2, rel_widths = c(1, 2))
    return(p3)
}
```










## Load the data



```{r load-data}
if (remove_doublets == FALSE && remove_ambient == FALSE) {
    path_modifier <- ""
}
if (remove_doublets == TRUE && remove_ambient == FALSE) {
    path_modifier <- "_doublets_removed"
}
if (remove_doublets == FALSE && remove_ambient == TRUE) {
    path_modifier <- "_ambient_removed"
}
if (remove_doublets == TRUE && remove_ambient == TRUE) {
    path_modifier <- "_doublets_removed_ambient_removed"
}
integrated <- read_rds(file = paste0(objects_path,
                                  dataset, "_",
                                  input_type, "_",
                                  normalization, "_",
                                  integration, "_",
                                  "integrated_seurat",
                                  path_modifier,
                                  ".rds")
                          )
```




```{r,fig.width=16, fig.height=10}
pt_size <- AutoPointSize(data = integrated, raster = FALSE)
DimPlot(object = integrated,
        reduction = paste0("umap_", curr_pc, "pcs"),
        group.by = "RNA_snn_res.10",
        pt.size = pt_size,
        repel = TRUE,
        label = TRUE,
        label.size = 5) +
  NoLegend() +
  coord_fixed()
```




# Cluster Markers


```{r}
list.files("../../proj136/analyses/markers/", pattern = ".csv")
```

```{r}
cluster_markers <- read_csv(file = "../../proj136/analyses/markers/New_Meta_cbNeuron_PUBnofacs_v1_alevin_LogNormalize_harmony_RNA_snn_res.10_RNA_cluster_markers_doublets_removed.csv") %>% 
    filter(p_val_adj < 0.05) %>% 
    filter(avg_log2FC > 0.8) %>%
    filter(pct.1 > 0.25)
```




# Turner-Evans et al 2021 - Central Complex

```{r}
central_complex <- read_xlsx(path = paste0(local_path_mod, turner_evans_path), skip = 3)
head(central_complex)
```


```{r}
fbgn_gene <- read_csv(paste0(local_path_mod, fbgn_gene_path))
head(fbgn_gene)
```


```{r}
central_complex_log <- fbgn_gene %>%
    select(FBgn, gene_symbol_r6.25) %>%
    unique() %>%
    drop_na() %>%
    inner_join(central_complex, by = c("FBgn" = "cell type")) %>%
    select(-FBgn, -contains("Whole Brain")) %>%
    rename("gene" = "gene_symbol_r6.25") %>%
    gather("cell_type", "expression",  -gene) %>%
    mutate(cell_type = str_remove(cell_type, pattern = "\\...[:digit:][:digit:]")) %>%
    mutate(cell_type = str_remove(cell_type, pattern = "\\...[:digit:]")) %>%
    group_by(cell_type, gene) %>%
    summarise(avg_exp = mean(expression)) %>%
    spread("cell_type", "avg_exp") %>%
    column_to_rownames("gene") %>%
    log1p() %>%
    rownames_to_column("gene")
head(central_complex_log)
```



## Correlate



```{r}
merged_avg_exp_by_cluster <- top_n_avg_exp(object = integrated,
                                           markers_df = cluster_markers,
                                           ident = "RNA_snn_res.10",
                                           n = 10,
                                           assay_use = "RNA",
                                           slot_use = "data")
merged_cx_slim_corr <- corr_bulk_sc(avg_exp_mat = merged_avg_exp_by_cluster, bulk_df = central_complex_log, assay_use = "RNA")
```


## Plot

```{r fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
Idents(integrated) <- integrated@meta.data[["RNA_snn_res.10"]]
cell_types <- unique(merged_cx_slim_corr$y)
for (i in seq_along(cell_types)) {
    p1 <- plot_corr_gap(object = integrated,
                        ident = "RNA_snn_res.10",
                        curr_pc = 240,
                        corr_df = merged_cx_slim_corr,
                        cell_type = cell_types[[i]],
                        y_limits = c(0, 1))
    plot(p1)
}
```









# Aso et al 2019 - pam dan, kc




```{r}
my_files <- list.files("../../../aallen/reference_datasets/bulk/GSE139889_Aso_2019_KC_PAMDAN/data/") %>% str_subset(".csv")
my_data <- list() 
for (i in seq_along(my_files)) {
    my_name <- my_files[[i]] %>% str_remove(".csv")
    my_data[[my_name]] <- read_csv(paste0("../../../aallen/reference_datasets/bulk/GSE139889_Aso_2019_KC_PAMDAN/data/", my_files[[i]]), locale = readr::locale(encoding = "UTF-8"))    
}
my_data[["transcriptID_x_counts"]] <- my_data[["transcriptID_x_counts"]] %>% 
    filter(!grepl(pattern = "DQ", x = transcript_id) & 
               !grepl(pattern = "EF", x = transcript_id))
my_data[["transcriptID_x_gene"]] <- my_data[["transcriptID_x_gene"]] %>% 
    filter(!grepl(pattern = "DQ", x = transcript_id) & 
               !grepl(pattern = "EF", x = transcript_id)) %>% 
    mutate(gene_symbol = if_else(gene_symbol == "01-Dec", "dec-1",gene_symbol)) %>% 
    mutate(gene_symbol = if_else(gene_symbol == "01-Sep", "Sep1",gene_symbol)) %>% 
    mutate(gene_symbol = if_else(gene_symbol == "02-Sep", "Sep2",gene_symbol)) %>% 
    mutate(gene_symbol = if_else(gene_symbol == "04-Sep", "Sep4",gene_symbol)) %>% 
    mutate(gene_symbol = if_else(gene_symbol == "05-Sep", "Sep5",gene_symbol)) 
my_data[["sampleID_x_celltype"]]$`Cell type`[my_data[["sampleID_x_celltype"]]$`Cell type` == "PMA-β1, PAM-β2"]
my_data[["sampleID_x_celltype"]] <- my_data[["sampleID_x_celltype"]] %>% 
    mutate(`Cell type` = str_replace(string = `Cell type`,
                                     pattern = "PMA-β1, PAM-β2",
                                     replacement = "PAM-β1, PAM-β2"),
           Sex = str_replace(string = as.character(Sex), 
                             pattern = "FALSE", 
                             replacement = "female")
           )
sum_by_gene <- my_data[["transcriptID_x_gene"]] %>% 
    mutate(gene_symbol = as.factor(gene_symbol)) %>% 
    select(transcript_id, gene_symbol) %>% 
    left_join(my_data[["transcriptID_x_counts"]], by = "transcript_id") %>% 
    gather("Sample_ID", "value", -transcript_id, -gene_symbol) %>% 
    left_join(select(my_data[["sampleID_x_celltype"]], Sample_ID, `Cell type`), by = "Sample_ID") %>% 
    group_by(gene_symbol, `Cell type`, Sample_ID) %>%
    drop_na() %>% 
    summarise(value = sum(value))
aso_etal_log <- sum_by_gene %>% 
    dplyr::select(-Sample_ID) %>% 
    dplyr::group_by(`Cell type`, gene_symbol) %>% 
    dplyr::summarise(avg_exp = mean(value)) %>% 
    tidyr::spread ("Cell type", "avg_exp") %>% 
    column_to_rownames("gene_symbol") %>% 
    log1p() %>% 
    tibble::rownames_to_column("gene_symbol") %>% 
    dplyr::rename("gene" = "gene_symbol")
```


## Correlate


```{r}
merged_avg_exp_by_cluster <- top_n_avg_exp(object = integrated, markers_df = cluster_markers, ident = "RNA_snn_res.10", n = 10, assay_use = "RNA", slot_use = "data")
merged_aso_etal_slim_corr <- corr_bulk_sc(avg_exp_mat = merged_avg_exp_by_cluster, bulk_df = aso_etal_log, assay_use = "RNA")
```


## Plot

```{r fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
Idents(integrated) <- integrated@meta.data[["RNA_snn_res.10"]]
cell_types <- unique(merged_aso_etal_slim_corr$y)
for (i in seq_along(cell_types)) {
    p1 <- plot_corr_gap(object = integrated,
                        ident = "RNA_snn_res.10",
                        curr_pc = 240,
                        corr_df = merged_cx_slim_corr,
                        cell_type = cell_types[[i]],
                        y_limits = c(0, 1))
    plot(p1)
}
```




















# Davis et al 2020

```{r}
davis_etal <- read_tsv(paste0(local_path_mod, davis_path)) %>%
    rename("gene" = "X1")
head(davis_etal)
```


```{r}
colnames(davis_etal)
```


```{r}
davis_slim_log <- davis_etal %>%
    column_to_rownames("gene") %>%
    log1p() %>%
    rownames_to_column("gene")
head(davis_slim_log)
```



## Correlate



```{r}
merged_avg_exp_by_cluster <- top_n_avg_exp(object = integrated,
                                           markers_df = cluster_markers,
                                           ident = "RNA_snn_res.10",
                                           n = 10,
                                           assay_use = "RNA",
                                           slot_use = "data")
merged_davis_slim_corr <- corr_bulk_sc(avg_exp_mat = merged_avg_exp_by_cluster, bulk_df = davis_slim_log, assay_use = "RNA")
```


## Plot

```{r fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
Idents(integrated) <- integrated@meta.data[["RNA_snn_res.10"]]
cell_types <- unique(merged_davis_slim_corr$y)
for (i in seq_along(cell_types)) {
    p1 <- plot_corr_gap(object = integrated,
                        ident = "RNA_snn_res.10",
                        curr_pc = 240,
                        corr_df = merged_davis_slim_corr,
                        cell_type = cell_types[[i]],
                        y_limits = c(0, 1))
    plot(p1)
}
```














# Konstantinides 2018 bulk

```{r}
konstantinides_bulk <- read_tsv(paste0(local_path_mod, konstantinides_path))
head(konstantinides_bulk)
```


```{r}
fbgn_gene <- read_csv(paste0(local_path_mod, fbgn_gene_path))
head(fbgn_gene)
```

```{r}
konstantinides_bulk <- fbgn_gene %>%
    select(FBgn, gene_symbol_r6.25) %>%
    unique() %>%
    drop_na() %>%
    inner_join(konstantinides_bulk, by = "FBgn") %>%
    select(-FBgn) %>%
    column_to_rownames("gene_symbol_r6.25") %>%
    log1p() %>%
    rownames_to_column("gene")
head(konstantinides_bulk)
```



## Correlate


```{r}
avg_exp_by_cluster <- top_n_avg_exp(object = integrated, markers = cluster_markers, ident = "RNA_snn_res.10", n = 10)
niko_slim_corr <- corr_bulk_sc(avg_exp_mat = avg_exp_by_cluster, bulk_df = konstantinides_bulk)
```



## Plot

```{r fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
Idents(integrated) <- integrated@meta.data[["RNA_snn_res.10"]]
cell_types <- unique(niko_slim_corr$y)
for (i in seq_along(cell_types)) {
    p1 <- plot_corr_gap(object = integrated,
                        ident = "RNA_snn_res.10",
                        curr_pc = 240,
                        corr_df = niko_slim_corr,
                        cell_type = cell_types[[i]],
                        y_limits = c(0, 1))
    plot(p1)
}
```







# Run time


```{r stop-time}
total_end_time <- Sys.time()
total_end_time
total_end_time - total_start_time
```



# Session info

```{r session-info}
sessionInfo()
```
