---
title: "Allen_2025b_Analysis__Figure_S2_DE_QC.Rmd"
description: description
author: "Aaron M. Allen"
date: 'Last update: `r date()`'
output:
  html_document:
    number_sections: true
    code_folding: show
    code_download: true
    theme: cerulean
    df_print: paged
    fig_width: 8.5
    fig_height: 5
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 6
---



```{css, echo = FALSE}
pre, code {white-space:pre !important; overflow-x:auto}
.tocify-item {white-space:pre}
```



```{r echo = FALSE, cache = FALSE}

```


```{r setup, echo = FALSE, cache = FALSE}
ggplot2::theme_set(ggplot2::theme_grey())
ggplot2::theme_update(panel.background = ggplot2::element_rect(fill = "transparent", colour = NA),
                      plot.background = ggplot2::element_rect(fill = "transparent", colour = NA))

rmd_name <- gsub(pattern = ".Rmd", replacement = "", x = knitr::current_input(), ignore.case = TRUE)
knitr::opts_chunk$set(dev = c("cairo_pdf"),
                      dev.args=list(bg="transparent"),
                      fig.align = "center",
                      fig.height = 5,
                      fig.width = 8.5,
                      dpi = 100,
                      pdf.options(encoding = "ISOLatin9.enc"),
                      fig.path = paste0("../analyses/figures/", rmd_name, "-",
                                        format(Sys.time(), "%Y%m%d_%H%M%S"), "/"),
                      fig.retina = 1,
                      warning = TRUE,
                      message = TRUE)
```



# Setup


## Start time

```{r start-time}
set.seed(12021)
today <- format(Sys.time(), "%Y%m%d_%H%M%S")
total_start_time <- Sys.time()
total_start_time
```





## Libraries

```{r packages-etc, warning = FALSE, message = FALSE}
library(Seurat)
library(cowplot)
library(colorspace)
library(tidyverse)
```



# Functions

## plot_sex_bias_across_celltypes_multi


```{r}

plot_sex_bias_across_celltypes_multi <- function(
    object,
    feature = "fru",
    grouping_condition = "subcluster_clusters",
    comparison_condition = "sex",
    data_slot = "counts",
    facet_plot = FALSE,
    facet_by = "experiment",
    y_limit = NULL,
    umi_count_df
    ){
    
    
    temp_data <- FetchData(object, vars = c(feature, grouping_condition, comparison_condition, facet_by), slot = data_slot) %>% 
        dplyr::rename("my_feature" = (!!feature)) %>% 
        dplyr::rename("my_grouping" = (!!grouping_condition)) %>% 
        dplyr::rename("my_comparison" = (!!comparison_condition)) %>% 
        dplyr::rename("my_facet" = (!!facet_by)) %>%
        filter(!my_facet %in% c("sleep", "species"))
    

    temp_result <- temp_data %>%
        group_by(my_grouping, my_comparison) %>%
        summarize(total_feature = sum(my_feature, na.rm = TRUE), .groups = "drop") %>%
        pivot_wider(names_from = my_comparison, values_from = total_feature, values_fill = list(total_feature = 0)) %>%
        mutate(log2_ratio = log2((female + 0.1) / (male + 0.1)))
    
    temp_result_1 <- temp_result %>%
        arrange(log2_ratio) %>%
        mutate(my_grouping = factor(my_grouping, levels = unique(my_grouping))) 
    
    temp_result_2a <- temp_data %>% 
        dplyr::group_by(my_grouping) %>% 
        dplyr::summarise(mean_feature = mean(my_feature), .groups = "drop") %>% 
        dplyr::mutate(log_mean_feature = log1p(mean_feature))
    temp_result_2a$my_grouping <- factor(temp_result_2a$my_grouping, levels = temp_result_1$my_grouping)
    
    temp_result_2b <- temp_data %>% 
        dplyr::group_by(my_grouping, my_facet) %>% 
        dplyr::summarise(sum_feature = sum(my_feature), .groups = "drop")
    temp_result_2b$my_grouping <- factor(temp_result_2b$my_grouping, levels = temp_result_1$my_grouping)
    
    x_closest_to_zero <- temp_result_1 %>%
        dplyr::slice(which.min(abs(log2_ratio))) %>%
        pull(my_grouping)
    # y_limit_calc <- ceiling(max(abs(temp_result_1$log2_ratio)))
    y_limit_calc <- temp_result_1 %>% 
        dplyr::filter(!my_grouping %in% c("Achl_003", "Achl_008", "Achl_058", "Glut_058", "GABA_012")) %>% 
        pull(log2_ratio) %>% abs() %>% max() %>% ceiling()
    if (is.null(y_limit)) {
        y_limit <- max(2, y_limit_calc)
    }
    
    
    temp_result_3 <- temp_data %>%
        group_by(my_grouping, my_facet, my_comparison) %>%
        summarize(total_feature = sum(my_feature, na.rm = TRUE), .groups = "drop") %>%
        pivot_wider(names_from = my_comparison, values_from = total_feature, values_fill = list(total_feature = 0)) %>%
        mutate(log2_ratio = log2((female + 0.1) / (male + 0.1)))

    p1 <- ggplot(temp_result_1, aes(x = my_grouping, y = log2_ratio, colour = log2_ratio)) +
        geom_hline(yintercept = 0, linetype = "dashed") +
        # geom_hline(yintercept = 0.5, linetype = "dashed") +
        # geom_hline(yintercept = -0.5, linetype = "dashed") +
        geom_vline(xintercept = x_closest_to_zero, linetype = "dashed") +
        geom_point(size = 2) +
        scale_y_continuous(labels = scales::number_format(accuracy = 0.01), limits = c(-y_limit, y_limit)) +
        # ylim(-y_limit, y_limit) +
        scale_color_gradientn(
            colours = c("#00bfc4", "black", "#c77cff"), 
            limits = c(-1,1), 
            oob = scales::squish
        ) +
        labs(
            x = grouping_condition, 
            y = paste0("log2(Female ", feature, " / Male ", feature, ")")
        ) +
        theme_classic() +
        theme(
            axis.title.x=element_blank(),
            axis.text.x=element_blank(),
            axis.ticks.x=element_blank()
        ) +
        NoLegend()
 
    p2a <- ggplot(temp_result_2a, aes(x = my_grouping, y = log_mean_feature)) +
        geom_bar(stat = "identity", position = "stack", fill = "#bcd2ee", colour = "white", width = 1) +
        scale_y_continuous(labels = scales::number_format(accuracy = 0.001)) +
        scale_fill_discrete_qualitative(palette = "Set 2") +
        # scale_y_continuous(trans = "log1p") +
        labs(
            y = paste0("log1p( Mean ", feature, " UMI )")
        ) +
        theme_classic() +
        theme(
            axis.title.x=element_blank(),
            axis.text.x=element_blank(),
            axis.ticks.x=element_blank()
        ) +
        NoLegend()

 
    p2b <- ggplot(temp_result_2b, aes(x = my_grouping, y = sum_feature, fill = my_facet)) +
        geom_bar(stat = "identity", position = "fill", colour = "white", width = 1) +
        scale_y_continuous(labels = scales::number_format(accuracy = 0.001)) +
        scale_fill_discrete_qualitative(palette = "Set 2") +
        labs(
            y = paste0("Proportion ", feature, " UMI")
        ) +
        theme_classic() +
        theme(
            axis.title.x=element_blank(),
            axis.text.x=element_blank(),
            axis.ticks.x=element_blank()
        ) +
        NoLegend()

    
    p3 <- ggplot(temp_result, aes(y = log2_ratio)) +
        geom_hline(yintercept = 0, linetype = "dashed") +
        geom_boxplot(outlier.size = -1) +
        ylim(-y_limit, y_limit) +
        labs(
            x = grouping_condition, 
            y = paste0("log2(Female ",  feature, " / Male ", feature, ")")
        ) +
        theme_classic() +
        theme(
            axis.title.x=element_blank(),
            axis.text.x=element_blank(),
            axis.ticks.x=element_blank()
        )
     

    p4 <- ggplot(temp_result_3, aes(x = my_facet, y = log2_ratio, colour = log2_ratio)) +
        geom_hline(yintercept = 0, linetype = "dashed") +
        geom_boxplot(aes(fill = my_facet), outlier.size = -1) +
        ylim(-y_limit, y_limit) +
        scale_fill_discrete_qualitative(palette = "Set 2") +
        labs(
            x = grouping_condition, 
            y = paste0("log2(Female ",  feature, " / Male ", feature, ")")
        ) +
        theme_classic() +
        theme(
            axis.title.x=element_blank(),
            axis.text.x=element_blank(),
            axis.ticks.x=element_blank()
        ) +
        NoLegend()
    
    
    temp_result_4 <- umi_count_df %>% 
        dplyr::filter(RowName == feature) %>% 
        dplyr::filter(!Dataset %in% c("sleep", "species")) %>% 
        dplyr::mutate(expected = overall_percent * sum(RowSum) / 100,
                      difference = log2(RowSum/expected)
                      ) 
    # log_norm_y_limit <- max(y_limit, max(abs(temp_result_4$difference)))
    log_norm_y_limit <- max(4, ceiling(max(abs(temp_result_4$difference))))
    
    p5 <- ggplot(temp_result_4, aes(x = Dataset, y = difference, fill = Dataset)) +
        geom_hline(yintercept = 0, linetype = "dashed") +
        geom_bar(stat = "identity") +
        ylim(-log_norm_y_limit, log_norm_y_limit) +
        scale_fill_discrete_qualitative(palette = "Set 2") +
        labs(
            y = "log2( UMI(obs) / UMI(exp) )"
        ) +
        theme_classic() +
        theme(
            axis.title.x=element_blank(),
            axis.text.x=element_blank(),
            axis.ticks.x=element_blank()
        ) +
        NoLegend()


    ptitle <- ggdraw() + draw_label(paste0("Log2 Ratio of Female to Male ", feature), 
                                   fontface = "bold", x = 0, hjust = 0, vjust = -0.1, size = 14)
    p6 <- plot_grid(p1, p2a, p2b, ncol = 1, rel_heights = c(2,1,1))
    p7 <- plot_grid(p6, p3, p4, p5, ncol = 4, rel_widths = c(8,1,1,1))
    # p8 <- plot_grid(p7, p6, ncol = 1, rel_heights = c(1,1))
    p9 <- plot_grid(ptitle, p7, ncol = 1, rel_heights = c(1, 10))
    
    return(p9)  
}


```



## plot_sex_bias_volcanoplots

```{r}
plot_sex_bias_volcanoplots <- function(
    clusterOI = NULL,
    deg_df = NULL,
    x_lims = NULL,
    y_lims = NULL,
    p_cutoff = 0.01,
    fc_cutoff = 1,
    is_zinb = TRUE
){
    clusterOI_deg_gene_filtered <- deg_df %>% 
        filter(Cluster == clusterOI) %>% 
        dplyr::filter(!Gene %in% c("lncRNA:roX1", "lncRNA:roX2")) %>%
        arrange(desc(abs(log2FoldChange)))
    # clusterOI_deg_gene_filtered
    
    if (is_zinb) {
        x_data <- "log2FoldChange"
        y_data <- "padj"
    } else {
        x_data <- "avg_log2FC"
        y_data <- "p_val_adj"
    }
    if (is.null(x_lims)) {
        x_lims <- ceiling(max(-log10(clusterOI_deg_gene_filtered[[x_data]]), na.rm = TRUE))
    }
    if (is.null(y_lims)) {
        y_lims <- ceiling(max(-log10(clusterOI_deg_gene_filtered[[y_data]]), na.rm = TRUE))
    }

    p1 <- EnhancedVolcano::EnhancedVolcano(
        toptable = clusterOI_deg_gene_filtered,
        lab = clusterOI_deg_gene_filtered$Gene,
        xlim = c(-x_lims, x_lims),
        ylim = c(0, y_lims),
        x = x_data,
        y = y_data,
        pCutoff = p_cutoff,
        FCcutoff = fc_cutoff,
        pointSize = 2,
        raster = TRUE,
        boxedLabels = FALSE,
        drawConnectors = TRUE,
        subtitle = paste0("Cluster = ", clusterOI,
                        ", test = ", "zinb-deseq"),
        caption = paste0("FC cutoff = ", fc_cutoff,
                       "; p-value cutoff = ", p_cutoff)
    )
    return(p1)
}

```



## plot_sex_bias_GO_analysis

```{r}

filter_and_run_sex_bias_GO_analysis <- function(
        de_df = NULL,                        # data frame of differential expression
        pval_thres = 0.05,                   # threshold cutoff of adjusted p-value
        log_fc_thres = 0.5,                  # threshold cutoff for log fold change
        which_sex = "female",                # 
        # log_fc_negative = FALSE,             # TRUE for negative logFC gene, ie male-biased
        # genes_to_keep = NULL,                # specified vector of genes to remove   
        # fg_genes = NULL,                    # vector of differentially expressed genes
        bg_genes = NULL,                     # vector of background genes
        use_bg_genes = FALSE,                # whether to use background genes or not                 
        ontology_catagory = "BP",            # one of "BP", "MF", "CC"
        org_db = "org.Dm.eg.db",             # drosophila melanogaster
        key_type = "SYMBOL"                  # using gene symbol for now
    ){
    
    library(clusterProfiler)
    library(org.Dm.eg.db)
    library(AnnotationDbi)
    
    sig_symbol <- de_df %>% 
        dplyr::filter(padj < pval_thres) %>%
        dplyr::mutate(log2FoldChange = if (which_sex == "male") -log2FoldChange else log2FoldChange) %>%
        dplyr::filter(log2FoldChange > log_fc_thres) %>% 
    	dplyr::select(Gene) %>% 
    	unique()
    # sig_symbol    
    
    go_results_bp_deg <- clusterProfiler::enrichGO(gene = sig_symbol$Gene, OrgDb = org_db, universe = bg_genes, keyType = key_type, ont = ontology_catagory)
    return(go_results_bp_deg)
    
}



plot_sex_bias_GO_analysis <- function(
        de_df = NULL,                        # data frame of differential expression
        pval_thres = 0.05,                   # threshold cutoff of adjusted p-value
        log_fc_thres = 0.5,                  # threshold cutoff for log fold change
        which_sex = "female",                #
        # log_fc_negative = FALSE,             # TRUE for negative logFC gene, ie male-biased
        # genes_to_keep = NULL,                # specified vector of genes to remove
        # fg_genes = NULL,                    # vector of differentially expressed genes
        bg_genes = NULL,                     # vector of background genes
        use_bg_genes = FALSE,                # whether to use background genes or not
        ontology_catagory = "BP",            # one of "BP", "MF", "CC"
        org_db = "org.Dm.eg.db",             # drosophila melanogaster
        key_type = "SYMBOL"                  # using gene symbol for now
    ){

    library(clusterProfiler)
    library(org.Dm.eg.db)
    library(AnnotationDbi)

    sig_symbol <- de_df %>%
        dplyr::filter(padj < pval_thres) %>%
        dplyr::mutate(log2FoldChange = if (which_sex == "male") -log2FoldChange else log2FoldChange) %>%
        dplyr::filter(log2FoldChange > log_fc_thres) %>%
    	dplyr::select(Gene) %>%
    	unique()
    # sig_symbol

    go_results_bp_deg <- clusterProfiler::enrichGO(gene = sig_symbol$Gene, OrgDb = org_db, universe = bg_genes, keyType = key_type, ont = ontology_catagory)
    go_results_bp_deg_df <- as.data.frame(go_results_bp_deg) %>% dplyr::arrange(qvalue)
    p1 <- dotplot(go_results_bp_deg, orderBy = "x", showCategory = 20) +
        labs(
            title = paste0(which_sex, " -- ", ontology_catagory),
            subtitle = paste0(
                "FC cutoff = ", log_fc_thres,
                "; p-value cutoff = ", pval_thres
            ),
            caption = paste0(
                "Use bg gene = ", use_bg_genes
            )
        )

    # return(go_results_bp_deg)
    print(go_results_bp_deg_df)
    plot(p1)

}
```





# Load data



```{r}
meta_cb <- read_rds("../../proj136/analyses/rds_files/New_Meta_cbNeuron_PUBnofacs_noPolIII_v1_alevin_LogNormalize_harmony_integrated_seurat_doublets_removed.rds")
DefaultAssay(meta_cb) <- "RNA"
Idents(meta_cb) <- "subcluster_clusters"
meta_cb
```



```{r}
options(scipen = 1)
deg_list <- list()
deg_list[["zinb-deseq"]] <- read_csv(file = "../../proj136/analyses/deg_sexed/New_Meta_cbNeuron_PUBnofacs_noPolIII_v1_alevin_LogNormalize_harmony_integrated_seurat_doublets_removed_bothsexes_DEG_sexed_subcluster_clusters_RNA_zinb_deseq_df_2.csv", col_types = cols())
```




# Gene Filtering



## iso and trans genes

```{r}
iso_genes_remove <- c("fru-plus", "fru-dP1", "fru-dP1-plus", "fruA", "fruB", "fruC",
					  "fruD1", "fruD2", "fruABint", "fruBCint", "fruCint", "fruPr1Ex1",
					  "fruPr1Ex2", "fruCOM", "dsxM", "dsxF", "dsxCOM",
					  "EGFP", "GFP", "mCherry", "Gal4", "GAL4"
)
```



## dataset specific genes


### load pre-calc. counts

```{r}
all_gene_max_sum__by_exp_together_wPercent <- read_csv(file = "../../proj136/analyses/all_gene_max_sum__by_exp_wPercent.csv")
head(all_gene_max_sum__by_exp_together_wPercent)
```


```{r}
all_gene_max_sum <- read_csv(file = "../../proj136/analyses/all_gene_max_sum.csv")
head(all_gene_max_sum)
```



### filter candidates

```{r}
genes_adfca_gt_30p <- all_gene_max_sum__by_exp_together_wPercent %>%
    dplyr::filter(Dataset == "adfca") %>%
    dplyr::filter(percent_dataset > 30) %>%
    pull(RowName)
length(genes_adfca_gt_30p)
```


```{r}
genes_any_gt_60p <- all_gene_max_sum__by_exp_together_wPercent %>%
    dplyr::filter(percent_dataset > 60) %>% 
    pull(RowName) %>% 
    unique()
length(genes_any_gt_60p)
```


```{r}
genes_low_exp <- all_gene_max_sum %>%
    dplyr::filter(RowSum < 400) %>% 
    dplyr::filter(RowMax < 4) %>% 
    pull(RowName) %>% 
    unique()
length(genes_low_exp)
```


```{r}
genes_remove <- sort(unique(c(genes_adfca_gt_30p, genes_any_gt_60p, genes_low_exp)))
length(genes_remove)
```


```{r}
genes_sexed_dataset <- all_gene_max_sum__by_exp_together_wPercent %>% 
    dplyr::filter(!Dataset %in% c("sleep", "species")) %>%
    dplyr::group_by(RowName) %>% 
    dplyr::summarise(sexed_dataset_counts = sum(percent_dataset)) %>% 
    dplyr::filter(sexed_dataset_counts > 45) %>% 
    pull(RowName)
length(genes_sexed_dataset)
```

```{r}
str_subset(string = genes_any_gt_60p, pattern = "whe")
```

```{r}
str_subset(string = genes_sexed_dataset, pattern = "Eip93F")
```

```{r}
all_gene_max_sum__by_exp_together_wPercent %>% 
    dplyr::filter(!Dataset %in% c("sleep", "species")) %>%
    dplyr::group_by(RowName) %>% 
    dplyr::summarise(sexed_dataset_counts = sum(percent_dataset)) %>% 
    dplyr::filter(RowName == "Eip93F")
```


## filter deg list

```{r}
deg_gene_filtered <- deg_list[["zinb-deseq"]] %>%
    dplyr::filter(!Gene %in% iso_genes_remove) %>%
    dplyr::filter(!Gene %in% genes_remove) %>%
    dplyr::filter(Gene %in% genes_sexed_dataset)
deg_gene_filtered
```




# Plot repeat offenders

```{r fig.height=8, fig.width=6}
deg_gene_filtered %>%
    dplyr::filter(padj < 0.01) %>% 
    dplyr::count(Gene) %>% 
    dplyr::arrange(n) %>%
    dplyr::mutate(Gene = factor(Gene, levels = unique(Gene))) %>% 
    dplyr::filter(n > 20) %>% 
    ggplot(aes(x = n, y = Gene)) +
        geom_bar(stat = "identity", fill = "#bcd2ee") +
        theme_classic() +
        theme(
            axis.title.x = element_blank(),
            axis.title.y = element_blank()
        ) +
        ggtitle("Number of cell types each gene is Sig Diff -- stringent")
```


```{r fig.height=12, fig.width=6}
deg_gene_filtered %>%
    dplyr::filter(padj < 0.05) %>% 
    dplyr::count(Gene) %>% 
    dplyr::arrange(n) %>%
    dplyr::mutate(Gene = factor(Gene, levels = unique(Gene))) %>% 
    dplyr::filter(n > 20) %>% 
    ggplot(aes(x = n, y = Gene)) +
        geom_bar(stat = "identity", fill = "#bcd2ee") +
        theme_classic() +
        theme(
            axis.title.x = element_blank(),
            axis.title.y = element_blank()
        ) +
        ggtitle("Number of cell types each gene is Sig Diff -- relaxed")
```




## Calc Normalized UMI


```{r}
overall_dataset_percent <- FetchData(object = meta_cb, vars = c("nCount_RNA", "experiment")) %>% 
    dplyr::filter(!experiment %in% c("sleep", "species")) %>%
    dplyr::group_by(experiment) %>% 
    dplyr::summarise(total_umi = sum(nCount_RNA)) %>% 
    dplyr::mutate(overall_percent = 100 * total_umi / sum(total_umi)) #%>% 
    # dplyr::select(experiment, overall_percent)
head(overall_dataset_percent)
```



```{r}
all_gene_max_sum__by_exp_together_wNormCont <- all_gene_max_sum__by_exp_together_wPercent %>% 
    dplyr::filter(!Dataset %in% c("sleep", "species")) %>%
    dplyr::group_by(RowName) %>% 
    dplyr::arrange(RowName) %>% 
    dplyr::mutate(percent_dataset = 100 * RowSum / sum(RowSum)) %>% 
    dplyr::left_join(overall_dataset_percent, by = c("Dataset" = "experiment")) %>% 
    dplyr::mutate(normalized_contribution = percent_dataset / overall_percent) 
head(all_gene_max_sum__by_exp_together_wNormCont)
```








## Plot log across cell types


```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=6}

genes <- deg_gene_filtered %>%
    dplyr::filter(padj < 0.05) %>% 
    dplyr::count(Gene) %>% 
    dplyr::arrange(desc(n)) %>% 
    dplyr::filter(n > 20) %>% 
    pull(Gene)

for (i in seq_along(genes)) {
    p1 <- plot_sex_bias_across_celltypes_multi(
        object = meta_cb, 
        feature = genes[[i]], 
        grouping_condition = "subcluster_clusters", 
        comparison_condition = "sex", 
        data_slot = "counts", 
        facet_plot = FALSE, 
        facet_by = "experiment",
        # y_limit = 3,
        umi_count_df = all_gene_max_sum__by_exp_together_wNormCont
    )
    plot(p1)
}

```






### Negative examples

```{r}
options(scipen = 1)
```


```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=6}

plot_sex_bias_across_celltypes_multi(
    object = meta_cb, 
    feature = "ninaE", 
    grouping_condition = "subcluster_clusters", 
    comparison_condition = "sex", 
    data_slot = "counts", 
    facet_plot = FALSE, 
    facet_by = "experiment",
    # y_limit = 3,
    umi_count_df = all_gene_max_sum__by_exp_together_wNormCont
)

```




```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=6}

plot_sex_bias_across_celltypes_multi(
    object = meta_cb, 
    feature = "Yp1", 
    grouping_condition = "subcluster_clusters", 
    comparison_condition = "sex", 
    data_slot = "counts", 
    facet_plot = FALSE, 
    facet_by = "experiment",
    # y_limit = 12,
    umi_count_df = all_gene_max_sum__by_exp_together_wNormCont
)

```




```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=6}

plot_sex_bias_across_celltypes_multi(
    object = meta_cb, 
    feature = "Nplp2", 
    grouping_condition = "subcluster_clusters", 
    comparison_condition = "sex", 
    data_slot = "counts", 
    facet_plot = FALSE, 
    facet_by = "experiment",
    # y_limit = 3,
    umi_count_df = all_gene_max_sum__by_exp_together_wNormCont
)

```


```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=6}

plot_sex_bias_across_celltypes_multi(
    object = meta_cb, 
    feature = "apolpp", 
    grouping_condition = "subcluster_clusters", 
    comparison_condition = "sex", 
    data_slot = "counts", 
    facet_plot = FALSE, 
    facet_by = "experiment",
    # y_limit = 3,
    umi_count_df = all_gene_max_sum__by_exp_together_wNormCont
)

```


```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=6}

plot_sex_bias_across_celltypes_multi(
    object = meta_cb, 
    feature = "Adh", 
    grouping_condition = "subcluster_clusters", 
    comparison_condition = "sex", 
    data_slot = "counts", 
    facet_plot = FALSE, 
    facet_by = "experiment",
    # y_limit = 4,
    umi_count_df = all_gene_max_sum__by_exp_together_wNormCont
)

```


```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=6}

plot_sex_bias_across_celltypes_multi(
    object = meta_cb, 
    feature = "Ilp2", 
    grouping_condition = "subcluster_clusters", 
    comparison_condition = "sex", 
    data_slot = "counts", 
    facet_plot = FALSE, 
    facet_by = "experiment",
    # y_limit = 4,
    umi_count_df = all_gene_max_sum__by_exp_together_wNormCont
)

```

```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=6}

plot_sex_bias_across_celltypes_multi(
    object = meta_cb, 
    feature = "RYa", 
    grouping_condition = "subcluster_clusters", 
    comparison_condition = "sex", 
    data_slot = "counts", 
    facet_plot = FALSE, 
    facet_by = "experiment",
    # y_limit = 4,
    umi_count_df = all_gene_max_sum__by_exp_together_wNormCont
)

```

```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=6}

plot_sex_bias_across_celltypes_multi(
    object = meta_cb, 
    feature = "lncRNA:CR44042", 
    grouping_condition = "subcluster_clusters", 
    comparison_condition = "sex", 
    data_slot = "counts", 
    facet_plot = FALSE, 
    facet_by = "experiment",
    # y_limit = 6,
    umi_count_df = all_gene_max_sum__by_exp_together_wNormCont
)

```


```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=6}

plot_sex_bias_across_celltypes_multi(
    object = meta_cb, 
    feature = "CG13924", 
    grouping_condition = "subcluster_clusters", 
    comparison_condition = "sex", 
    data_slot = "counts", 
    facet_plot = FALSE, 
    facet_by = "experiment",
    # y_limit = 6,
    umi_count_df = all_gene_max_sum__by_exp_together_wNormCont
)

```





### postive examples



```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=6}

plot_sex_bias_across_celltypes_multi(
    object = meta_cb, 
    feature = "elav", 
    grouping_condition = "subcluster_clusters", 
    comparison_condition = "sex", 
    data_slot = "counts", 
    facet_plot = FALSE, 
    facet_by = "experiment",
    # y_limit = 8,
    umi_count_df = all_gene_max_sum__by_exp_together_wNormCont
)

```

```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=6}

plot_sex_bias_across_celltypes_multi(
    object = meta_cb, 
    feature = "nSyb", 
    grouping_condition = "subcluster_clusters", 
    comparison_condition = "sex", 
    data_slot = "counts", 
    facet_plot = FALSE, 
    facet_by = "experiment",
    # y_limit = 6,
    umi_count_df = all_gene_max_sum__by_exp_together_wNormCont
)

```



```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=6}

plot_sex_bias_across_celltypes_multi(
    object = meta_cb, 
    feature = "VAChT", 
    grouping_condition = "subcluster_clusters", 
    comparison_condition = "sex", 
    data_slot = "counts", 
    facet_plot = FALSE, 
    facet_by = "experiment",
    # y_limit = 6,
    umi_count_df = all_gene_max_sum__by_exp_together_wNormCont
)

```


```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=6}

plot_sex_bias_across_celltypes_multi(
    object = meta_cb, 
    feature = "VGlut", 
    grouping_condition = "subcluster_clusters", 
    comparison_condition = "sex", 
    data_slot = "counts", 
    facet_plot = FALSE, 
    facet_by = "experiment",
    # y_limit = 6,
    umi_count_df = all_gene_max_sum__by_exp_together_wNormCont
)

```



```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=6}

plot_sex_bias_across_celltypes_multi(
    object = meta_cb, 
    feature = "Gad1", 
    grouping_condition = "subcluster_clusters", 
    comparison_condition = "sex", 
    data_slot = "counts", 
    facet_plot = FALSE, 
    facet_by = "experiment",
    # y_limit = 6,
    umi_count_df = all_gene_max_sum__by_exp_together_wNormCont
)

```





## Plot SDH


```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=6}

plot_sex_bias_across_celltypes_multi(
    object = meta_cb, 
    feature = "lncRNA:roX1", 
    grouping_condition = "subcluster_clusters", 
    comparison_condition = "sex", 
    data_slot = "counts", 
    facet_plot = FALSE, 
    facet_by = "experiment",
    # y_limit = 14,
    umi_count_df = all_gene_max_sum__by_exp_together_wNormCont
)

```




```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=6}

plot_sex_bias_across_celltypes_multi(
    object = meta_cb, 
    feature = "lncRNA:roX2", 
    grouping_condition = "subcluster_clusters", 
    comparison_condition = "sex", 
    data_slot = "counts", 
    facet_plot = FALSE, 
    facet_by = "experiment",
    # y_limit = 14,
    umi_count_df = all_gene_max_sum__by_exp_together_wNormCont
)

```




```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=6}

genes <- c(
    # "fru-plus", "fru-dP1", "fru-dP1-plus", "fruA", "fruB", "fruC",
    # "fruD1", "fruD2", "fruABint", "fruBCint", "fruCint", "fruPr1Ex1",
    "mle", "mof", "msl-1", "msl-2", "msl-3",
    "Sxl", "ssx", "tra", "tra2", "fru", "dsx"
)

for (i in seq_along(genes)) {
    p1 <- plot_sex_bias_across_celltypes_multi(
        object = meta_cb, 
        feature = genes[[i]], 
        grouping_condition = "subcluster_clusters", 
        comparison_condition = "sex", 
        data_slot = "counts", 
        facet_plot = FALSE, 
        facet_by = "experiment",
        # y_limit = 3,
        umi_count_df = all_gene_max_sum__by_exp_together_wNormCont
    )
    plot(p1)
}

```












# Session info

```{r session_info}
sessionInfo()
```



