---
title: "Allen_2025a_Analysis__Figure_3_Imp_v_dati_morphology_quantification"
description: description
author: "Aaron M. Allen"
date: 'Last update: `r date()`'
output:
  html_document:
    number_sections: true
    code_folding: show
    code_download: true
    theme: cerulean
    df_print: paged
    fig_width: 8.5
    fig_height: 5
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 6
---



```{css, echo = FALSE}
pre, code {white-space:pre !important; overflow-x:auto}
.tocify-item {white-space:pre}
```



```{r echo = FALSE, cache = FALSE}

```


```{r setup, echo = FALSE, cache = FALSE}
ggplot2::theme_set(ggplot2::theme_grey())
ggplot2::theme_update(panel.background = ggplot2::element_rect(fill = "transparent", colour = NA),
                      plot.background = ggplot2::element_rect(fill = "transparent", colour = NA))

rmd_name <- gsub(pattern = ".Rmd", replacement = "", x = knitr::current_input(), ignore.case = TRUE)
knitr::opts_chunk$set(
	dev = c("cairo_pdf"),
	# dev = c("png", "cairo_pdf"),
	dev.args=list(bg="transparent"),
	fig.align = "center",
	fig.height = 5,
	fig.width = 8.5,
	dpi = 100,
	pdf.options(encoding = "ISOLatin9.enc"),
	fig.path = paste0("../analyses/figures/", rmd_name, "-",
	                format(Sys.time(), "%Y%m%d_%H%M%S"), "/"),
	fig.retina = 1,
	warning = TRUE,
	message = TRUE)
```



# Setup


## Start time

```{r start-time}
set.seed(12021)
today <- format(Sys.time(), "%Y%m%d_%H%M%S")
total_start_time <- Sys.time()
total_start_time
```






## Parameters





## Libraries

```{r packages-etc, warning = FALSE, message = FALSE}
library(Seurat)
library(tidyverse)
library(colorspace)
library(cowplot)
```





# Load data


```{r load-dev-age}
meta_dev_age <- read_rds("../../proj136/analyses/rds_files/New_Meta_cbNeuron_PUBnofacs_dev_age_v1_alevin_LogNormalize_harmony_integrated_seurat_doublets_removed_diet.rds")
meta_dev_age
```


```{r}
dev_age_metadata <- FetchData(object = meta_dev_age, vars = "dev_age") %>% rownames_to_column("cell_id")
```

```{r}
rm(meta_dev_age)
gc()
```



```{r load-meta-brain}
meta_cb <- read_rds("../../proj136/analyses/rds_files/New_Meta_cbNeuron_PUBnofacs_noPolIII_v1_alevin_LogNormalize_harmony_integrated_seurat_doublets_removed_diet.rds")
meta_cb
```


```{r}
add_dev_age_metadata <- FetchData(object = meta_cb, vars = "annotation_broad_cell_type_ext_fix") %>% 
	rownames_to_column("cell_id") %>% 
	left_join(dev_age_metadata, by = "cell_id")
# add_dev_age_metadata
meta_cb <- AddMetaData(object = meta_cb, metadata = add_dev_age_metadata$dev_age, col.name = "dev_age")
```


```{r}
dev_age_filtered_cells <- FetchData(object = meta_cb, vars = c("dev_age")) %>% 
    rownames_to_column("cell_id") %>% 
	filter(!is.na(dev_age)) %>% 
    pull(cell_id)
meta_cb_filtered <- subset(x = meta_cb, cells = dev_age_filtered_cells)
meta_cb_filtered
```





# Set Embeddings and Clustering


```{r}
reduction_use <- commandArgs(trailingOnly=T)[4]
if (is.na(reduction_use)) {
	reduction_use <- "tsne_240pcs"
}
reduction_use
```


```{r}
clustering_use <- commandArgs(trailingOnly=T)[5]
if (is.na(clustering_use)) {
	clustering_use <- "subcluster_clusters"
}
clustering_use
```



```{r}
embeddings <- Embeddings(meta_cb_filtered, reduction = reduction_use)
clusters <- meta_cb_filtered@meta.data[[clustering_use]]
```



# Imp v dati morphology



## 1. Pairwise Distance Metrics
Compute the average distance between all points within a cluster.

```{r pairwise_distances}

pairwise_distances <- clusters %>%
  unique() %>%
  sapply(function(cluster) {
    points <- embeddings[clusters == cluster, ]
    mean(dist(points)) # Mean pairwise distance
  })

pairwise_distances

```


```{r}
cluster_by_pairwise_distances <- data.frame(subcluster_clusters = clusters %>% unique(),
		   subcluster_clusters_pairwise_distances = pairwise_distances,
		   log_pairwise_distances = log2(pairwise_distances))
# cluster_by_pairwise_distances

cluster_by_pairwise_distances_metadata <- FetchData(object = meta_cb_filtered, vars = "subcluster_clusters") %>% 
	rownames_to_column("cell_id") %>% 
	left_join(cluster_by_pairwise_distances, by = "subcluster_clusters")
# cluster_by_pairwise_distances_metadata

meta_cb_filtered <- AddMetaData(object = meta_cb_filtered, metadata = cluster_by_pairwise_distances_metadata$subcluster_clusters_pairwise_distances, col.name = "cluster_pairwise_distances")
meta_cb_filtered <- AddMetaData(object = meta_cb_filtered, metadata = cluster_by_pairwise_distances_metadata$log_pairwise_distances, col.name = "log_pairwise_distances")
```



```{r warning=FALSE, message=FALSE, fig.height=12, fig.width=20}
genes <- c("cluster_pairwise_distances")
plist <- FeaturePlot(
			object = meta_cb_filtered, 
			reduction = reduction_use, 
			features = genes,
			combine = FALSE,
			ncol = 1,
			pt.size = 0.4, 
			# min.cutoff = 10,
			# max.cutoff = 50,
			# slot = "counts",
			cols = c("lightsteelblue2", "black"), 
			raster = FALSE,
			raster.dpi = c(1024, 1024),
			order = TRUE,
			coord.fixed = TRUE)

for (i in seq_along(plist)) {
	plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}
```



```{r warning=FALSE, message=FALSE, fig.height=12, fig.width=20}
genes <- c("cluster_pairwise_distances")
plist <- FeaturePlot(
			object = meta_cb_filtered, 
			reduction = reduction_use, 
			features = genes,
			combine = FALSE,
			ncol = 1,
			pt.size = 0.4, 
			min.cutoff = 10,
			max.cutoff = 50,
			# slot = "counts",
			cols = c("lightsteelblue2", "black"), 
			raster = FALSE,
			raster.dpi = c(1024, 1024),
			order = TRUE,
			coord.fixed = TRUE)

for (i in seq_along(plist)) {
	plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}
```




```{r fig.width=4, fig.height=6}
dev_age_by_cluster_pairwise_distances <- FetchData(object = meta_cb_filtered, vars = c("dev_age", "cluster_pairwise_distances"))

dev_age_by_cluster_pairwise_distances %>%
	filter(!is.na(dev_age)) %>%                                                 # filtering by dev_age also removes KCs
	ggplot(aes(x = dev_age, y = cluster_pairwise_distances)) +
		geom_boxplot(outlier.size = -1) +
		theme_classic()
	
```



```{r}
early_cluster_pairwise_distances <- dev_age_by_cluster_pairwise_distances %>% filter(dev_age == "early_born") %>% pull(cluster_pairwise_distances)
late_cluster_pairwise_distances <- dev_age_by_cluster_pairwise_distances %>% filter(dev_age == "late_born") %>% pull(cluster_pairwise_distances)
```


```{r}
t.test(x = early_cluster_pairwise_distances, y = late_cluster_pairwise_distances, alternative = "two.sided")
```

```{r}
wilcox.test(x = early_cluster_pairwise_distances, y = late_cluster_pairwise_distances, alternative = "two.sided")
```




















## 2. Modularity or Community Detection
Treat the cluster as a graph, where points are nodes and edges connect nearby points.


```{r modularity_scores}
library(igraph)
modularity_scores <- unique(clusters) %>%
  sapply(function(cluster) {
    points <- embeddings[clusters == cluster, ]
    
    if (nrow(points) < 2) {
        return(NA)
    }
    dist_matrix <- as.matrix(dist(points))
    radius <- 1
    adj_matrix <- ifelse(dist_matrix < radius, 1, 0)
    diag(adj_matrix) <- 0
    graph <- graph_from_adjacency_matrix(adj_matrix, mode = "undirected", diag = FALSE)
    communities <- cluster_fast_greedy(graph)
    modularity(communities)
})
modularity_scores
```




```{r}
cluster_by_modularity_scores <- data.frame(subcluster_clusters = clusters %>% unique(),
		   subcluster_clusters_modularity_scores = modularity_scores
		   )
# cluster_by_modularity_scores

cluster_by_modularity_scores_metadata <- FetchData(object = meta_cb_filtered, vars = "subcluster_clusters") %>% 
	rownames_to_column("cell_id") %>% 
	left_join(cluster_by_modularity_scores, by = "subcluster_clusters")
# cluster_by_modularity_scores_metadata

meta_cb_filtered <- AddMetaData(object = meta_cb_filtered, metadata = cluster_by_modularity_scores_metadata$subcluster_clusters_modularity_scores, col.name = "modularity_scores")
```



```{r warning=FALSE, message=FALSE, fig.height=12, fig.width=20}
genes <- c("modularity_scores")
plist <- FeaturePlot(
			object = meta_cb_filtered, 
			reduction = reduction_use, 
			features = genes,
			combine = FALSE,
			ncol = 1,
			pt.size = 0.4, 
			# min.cutoff = 1,
			# max.cutoff = 3,
			# slot = "counts",
			cols = c("lightsteelblue2", "black"), 
			raster = FALSE,
			raster.dpi = c(1024, 1024),
			order = TRUE,
			coord.fixed = TRUE)

for (i in seq_along(plist)) {
	plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}
```




```{r warning=FALSE, message=FALSE, fig.height=12, fig.width=20}
genes <- c("modularity_scores")
plist <- FeaturePlot(
			object = meta_cb_filtered, 
			reduction = reduction_use, 
			features = genes,
			combine = FALSE,
			ncol = 1,
			pt.size = 0.4, 
			min.cutoff = 0.7,
			max.cutoff = 0.9,
			# slot = "counts",
			cols = c("lightsteelblue2", "black"), 
			raster = FALSE,
			raster.dpi = c(1024, 1024),
			order = TRUE,
			coord.fixed = TRUE)

for (i in seq_along(plist)) {
	plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}
```




```{r fig.width=4, fig.height=6}
dev_age_by_modularity_scores <- FetchData(object = meta_cb_filtered, vars = c("dev_age", "modularity_scores"))

dev_age_by_modularity_scores %>%
	filter(!is.na(dev_age)) %>%                                                  # filtering by dev_age also removes KCs
	ggplot(aes(x = dev_age, y = modularity_scores)) +
		geom_boxplot(outlier.size = -1) +
		# ylim(-4, 4) +
		theme_classic()
	
```



```{r}
early_modularity_scores <- dev_age_by_modularity_scores %>% filter(dev_age == "early_born") %>% pull(modularity_scores)
late_modularity_scores <- dev_age_by_modularity_scores %>% filter(dev_age == "late_born") %>% pull(modularity_scores)
```


```{r}
t.test(x = early_modularity_scores, y = late_modularity_scores, alternative = "two.sided")
```

```{r}
wilcox.test(x = early_modularity_scores, y = late_modularity_scores, alternative = "two.sided")
```
















## 3. Cluster Fragmentation
Identify subgroups within the cluster using DBSCAN.


```{r fragmentation_score}
library(dbscan)

fragmentation_score <- unique(clusters) %>%
  sapply(function(cluster) {
    points <- embeddings[clusters == cluster, ]
    db <- dbscan::dbscan(points, eps = 0.5, minPts = 5)
    length(unique(db$cluster))
  })

fragmentation_score
```


```{r}
cluster_by_fragmentation_score <- data.frame(subcluster_clusters = clusters %>% unique(),
		   subcluster_clusters_fragmentation_score = fragmentation_score
		   )
# cluster_by_fragmentation_score

cluster_by_fragmentation_score_metadata <- FetchData(object = meta_cb_filtered, vars = "subcluster_clusters") %>% 
	rownames_to_column("cell_id") %>% 
	left_join(cluster_by_fragmentation_score, by = "subcluster_clusters")
# cluster_by_fragmentation_score_metadata

meta_cb_filtered <- AddMetaData(object = meta_cb_filtered, metadata = cluster_by_fragmentation_score_metadata$subcluster_clusters_fragmentation_score, col.name = "fragmentation_score")
```



```{r warning=FALSE, message=FALSE, fig.height=12, fig.width=20}
genes <- c("fragmentation_score")
plist <- FeaturePlot(
			object = meta_cb_filtered, 
			reduction = reduction_use, 
			features = genes,
			combine = FALSE,
			ncol = 1,
			pt.size = 0.4, 
			# min.cutoff = 1,
			# max.cutoff = 3,
			# slot = "counts",
			cols = c("lightsteelblue2", "black"), 
			raster = FALSE,
			raster.dpi = c(1024, 1024),
			order = TRUE,
			coord.fixed = TRUE)

for (i in seq_along(plist)) {
	plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}
```


```{r warning=FALSE, message=FALSE, fig.height=12, fig.width=20}
genes <- c("fragmentation_score")
plist <- FeaturePlot(
			object = meta_cb_filtered, 
			reduction = reduction_use, 
			features = genes,
			combine = FALSE,
			ncol = 1,
			pt.size = 0.4, 
			min.cutoff = 10,
			max.cutoff = 80,
			# slot = "counts",
			cols = c("lightsteelblue2", "black"), 
			raster = FALSE,
			raster.dpi = c(1024, 1024),
			order = TRUE,
			coord.fixed = TRUE)

for (i in seq_along(plist)) {
	plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}
```


```{r warning=FALSE, message=FALSE, fig.height=12, fig.width=20}
genes <- c("fragmentation_score")
plist <- FeaturePlot(
			object = meta_cb_filtered, 
			reduction = reduction_use, 
			features = genes,
			combine = FALSE,
			ncol = 1,
			pt.size = 0.4, 
			min.cutoff = 10,
			max.cutoff = 40,
			# slot = "counts",
			cols = c("lightsteelblue2", "black"), 
			raster = FALSE,
			raster.dpi = c(1024, 1024),
			order = TRUE,
			coord.fixed = TRUE)

for (i in seq_along(plist)) {
	plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}
```


```{r fig.width=4, fig.height=6}
dev_age_by_fragmentation_score <- FetchData(object = meta_cb_filtered, vars = c("dev_age", "fragmentation_score"))

dev_age_by_fragmentation_score %>%
	filter(!is.na(dev_age)) %>%                                                  # filtering by dev_age also removes KCs
	ggplot(aes(x = dev_age, y = fragmentation_score)) +
		geom_boxplot(outlier.size = -1) +
		# ylim(-4, 4) +
		theme_classic()
	
```


```{r}
early_fragmentation_score <- dev_age_by_fragmentation_score %>% filter(dev_age == "early_born") %>% pull(fragmentation_score)
late_fragmentation_score <- dev_age_by_fragmentation_score %>% filter(dev_age == "late_born") %>% pull(fragmentation_score)
```


```{r}
t.test(x = early_fragmentation_score, y = late_fragmentation_score, alternative = "two.sided")
```

```{r}
wilcox.test(x = early_fragmentation_score, y = late_fragmentation_score, alternative = "two.sided")
```









#### scan radii


```{r fragmentation_score-scan}

radii <- c(seq(0.1, 2, 0.1), seq(3, 10, 1))
fragmentation_score_list <- list()

for (i in seq_along(radii)) {
	
	message(paste0("eps = ", radii[[i]]))
	fragmentation_score_list[[i]] <- unique(clusters) %>%
		sapply(function(cluster) {
			points <- embeddings[clusters == cluster, ]
			db <- dbscan::dbscan(points, eps = radii[[i]], minPts = 5)
			length(unique(db$cluster)) # Number of distinct subclusters
		})
	
}

names(fragmentation_score_list) <- paste0("eps", radii)

```




```{r}

cluster_by_fragmentation_score <- data.frame(subcluster_clusters = clusters %>% unique()) %>% 
	bind_cols(fragmentation_score_list)

cluster_by_fragmentation_score_metadata <- FetchData(object = meta_cb_filtered, vars = c("subcluster_clusters", "dev_age")) %>% 
	rownames_to_column("cell_id") %>% 
	left_join(cluster_by_fragmentation_score, by = "subcluster_clusters")
cluster_by_fragmentation_score_metadata

```



```{r}
calculate_ci <- function(x, conf_level = 0.95) {
	n <- length(x)
	se <- sd(x, na.rm = TRUE) / sqrt(n)
	margin <- qt((1 + conf_level) / 2, df = n - 1) * se
	mean_x <- mean(x, na.rm = TRUE)
	c(lower = mean_x - margin, upper = mean_x + margin)
}
```


```{r}

result <- cluster_by_fragmentation_score_metadata %>% 
	filter(!is.na(dev_age)) %>%
	dplyr::select(dev_age, starts_with("eps")) %>%
	pivot_longer(cols = starts_with("eps"), names_to = "eps_column", values_to = "value") %>%
	group_by(dev_age, eps_column) %>%
	summarise(
		mean = mean(value, na.rm = TRUE),
		median = median(value, na.rm = TRUE),
		sd = sd(value, na.rm = TRUE),
		se = sd(value, na.rm = TRUE) / sqrt(sum(!is.na(value))),
		cv = sd(value, na.rm = TRUE) / mean(value, na.rm = TRUE),
		ci_lower = calculate_ci(value)[1],
		ci_upper = calculate_ci(value)[2],
		.groups = "drop"
	)

result

```



```{r}
result %>% 
	mutate(eps_column = eps_column %>% str_remove("eps") %>% as.numeric()) %>% 
	ggplot(aes(x = eps_column, y = mean, color = dev_age, fill = dev_age)) +
		geom_point() +
		geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper)) +
		labs(title = "Varying eps affects on fragmentation score",
			 y = "mean fragmentation score",
			 x = "eps") +
		theme_classic()
```



#### scan minpts


```{r fragmentation_score-scan-2}

min_pts <- c(seq(1, 20, 1))
fragmentation_score_list <- list()

for (i in seq_along(min_pts)) {
	
	message(paste0("minPts = ", min_pts[[i]]))
	fragmentation_score_list[[i]] <- unique(clusters) %>%
		sapply(function(cluster) {
			points <- embeddings[clusters == cluster, ]
			db <- dbscan::dbscan(points, eps = 0.5, minPts = min_pts[[i]])
			length(unique(db$cluster)) # Number of distinct subclusters
		})
	
}

names(fragmentation_score_list) <- paste0("minPts", min_pts)

```




```{r}

cluster_by_fragmentation_score <- data.frame(subcluster_clusters = clusters %>% unique()) %>% 
	bind_cols(fragmentation_score_list)

cluster_by_fragmentation_score_metadata <- FetchData(object = meta_cb_filtered, vars = c("subcluster_clusters", "dev_age")) %>% 
	rownames_to_column("cell_id") %>% 
	left_join(cluster_by_fragmentation_score, by = "subcluster_clusters")
cluster_by_fragmentation_score_metadata

```



```{r}
calculate_ci <- function(x, conf_level = 0.95) {
	n <- length(x)
	se <- sd(x, na.rm = TRUE) / sqrt(n)
	margin <- qt((1 + conf_level) / 2, df = n - 1) * se
	mean_x <- mean(x, na.rm = TRUE)
	c(lower = mean_x - margin, upper = mean_x + margin)
}
```


```{r}

result <- cluster_by_fragmentation_score_metadata %>% 
	filter(!is.na(dev_age)) %>%
	dplyr::select(dev_age, starts_with("minPts")) %>%
	pivot_longer(cols = starts_with("minPts"), names_to = "minPts_column", values_to = "value") %>%
	group_by(dev_age, minPts_column) %>%
	summarise(
		mean = mean(value, na.rm = TRUE),
		median = median(value, na.rm = TRUE),
		sd = sd(value, na.rm = TRUE),
		se = sd(value, na.rm = TRUE) / sqrt(sum(!is.na(value))),
		cv = sd(value, na.rm = TRUE) / mean(value, na.rm = TRUE),
		ci_lower = calculate_ci(value)[1],
		ci_upper = calculate_ci(value)[2],
		.groups = "drop"
	)

result

```


```{r}
result %>% 
	mutate(minPts_column = minPts_column %>% str_remove("minPts") %>% as.numeric()) %>% 
	ggplot(aes(x = minPts_column, y = mean, color = dev_age, fill = dev_age)) +
		geom_point() +
		geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper)) +
		labs(title = "Varying minPts affects on fragmentation score",
			 y = "mean fragmentation score",
			 x = "minPts") +
		theme_classic()
```







# Session info

```{r session_info}
sessionInfo()
```

