---
title: "Allen_2025b_Analysis__Figure_2I_dsx_scenic_for_cytoscape"
description: description
author: "Aaron M. Allen"
date: 'Last update: `r date()`'
output:
  html_document:
    number_sections: true
    code_folding: show
    code_download: true
    theme: cerulean
    df_print: paged
    fig_width: 8.5
    fig_height: 5
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 6
---



```{css, echo = FALSE}
pre, code {white-space:pre !important; overflow-x:auto}
.tocify-item {white-space:pre}
```



```{r echo = FALSE, cache = FALSE}

```


```{r setup, echo = FALSE, cache = FALSE}
ggplot2::theme_set(ggplot2::theme_grey())
ggplot2::theme_update(panel.background = ggplot2::element_rect(fill = "transparent", colour = NA),
                      plot.background = ggplot2::element_rect(fill = "transparent", colour = NA))

rmd_name <- gsub(pattern = ".Rmd", replacement = "", x = knitr::current_input(), ignore.case = TRUE)
knitr::opts_chunk$set(dev = c("png", "cairo_pdf"),
                      dev.args=list(bg="transparent"),
                      fig.align = "center",
                      fig.height = 5,
                      fig.width = 8.5,
                      dpi = 300,
                      pdf.options(encoding = "ISOLatin9.enc"),
                      fig.path = paste0("../analyses/figures/", rmd_name, "-",
                                        format(Sys.time(), "%Y%m%d_%H%M%S"), "/"),
                      fig.retina = 1,
                      warning = TRUE,
                      message = TRUE)
```



# Setup


## Start time

```{r start-time}
set.seed(12021)
today <- format(Sys.time(), "%Y%m%d_%H%M%S")
total_start_time <- Sys.time()
total_start_time
```






## Parameters





## Libraries

```{r packages-etc, warning = FALSE, message = FALSE}
library(Seurat)
library(cowplot)
library(pheatmap)
library(colorspace)
library(tidyverse)
```

```{r}
dsx_scenic <- read_rds("../../proj136/analyses/scenic/scenic_multi_dsx/results.dir/aggregated.dir/raw.dir/dsx.dir/pyscenic_r.dir/scenic_seurat_raw.rds")
dsx_scenic
```


```{r}
dsx_scenic@assays$scenic %>% rownames()
```


```{r}
dsx_seu_NEW <- read_rds("../../proj136/analyses/rds_files/subclustering_dsx__seu--round3--500VarGene--NoReg--Harmony_ori_exp_cell.rds")
dsx_seu_NEW
```



```{r}
dsx_scenic <- subset(x = dsx_scenic, cells = colnames(dsx_seu_NEW))
dsx_scenic
```



# Scenic


## Some Plots

```{r}
dsx_scenic$cell_type %>% unique()
```

```{r}
colours_dsx <- c("#E32B1F", "#E96344", "#F4A98E", "#BDAB98", "#00AC99", "#60C1D8")
```


```{r fig.width=12, fig.height=12}

reductions_to_plot <- dsx_scenic@reductions %>% names() %>% str_subset("scenic_umap_")

for (i in seq_along(reductions_to_plot)) {
    plot(DimPlot(object = dsx_scenic, reduction = reductions_to_plot[[i]], group.by = "cell_type", cols = colours_dsx, pt.size = 2) + coord_fixed())
}

```



```{r fig.width=12, fig.height=12}

reductions_to_plot <- dsx_scenic@reductions %>% names() %>% str_subset("scenic_tsne_")

for (i in seq_along(reductions_to_plot)) {
    plot(DimPlot(object = dsx_scenic, reduction = reductions_to_plot[[i]], group.by = "cell_type", cols = colours_dsx, pt.size = 2) + coord_fixed())
}

```






```{r warning=FALSE, message=FALSE, fig.width=12, fig.height=12}

reductions_to_plot <- dsx_scenic@reductions %>% names() %>% str_subset("scenic_umap_")

for (i in seq_along(reductions_to_plot)) {
    plot(DimPlot(object = dsx_scenic, reduction = reductions_to_plot[[i]], group.by = "RNA_snn_res.1", pt.size = 2) + scale_colour_discrete_qualitative(palette = "Set 2") + coord_fixed())
}

```



```{r warning=FALSE, message=FALSE, fig.width=12, fig.height=12}

reductions_to_plot <- dsx_scenic@reductions %>% names() %>% str_subset("scenic_tsne_")

for (i in seq_along(reductions_to_plot)) {
    plot(DimPlot(object = dsx_scenic, reduction = reductions_to_plot[[i]], group.by = "RNA_snn_res.1", pt.size = 3) + scale_colour_discrete_qualitative(palette = "Set 2") + coord_fixed())
}

```





## Re-Run markers


```{r}

params <- list()

params$datatype <- "raw"

params$condition <- "sex"
params$celltype <- "cell_type"
params$celltype_condition <- "cell_type_sexed"

params$reference_condition <- "female"
params$latent_variables <- "experiment"
params$diff_exp_test <- "wilcox"
params$FDR_threshold <- 0.05
params$top_n <- 5

params$results_directory <- "../../proj136/analyses/scenic/scenic_multi_dsx/results.dir/aggregated.dir/raw.dir/dsx.dir/pyscenic_r.dir/"



```




#### Stats to identify regulons that are differentially-expressed between conditions or cell types or celltypes_conditions

```{r diff-reg-celltype}
# Latent variables for test
if (params$latent_variables == "None" | params$diff_exp_test == "wilcox") {
    latent_variables <- NULL
} else {
    latent_variables <- params$latent_variables
}

# Cell type - one vs. all
DefaultAssay(dsx_scenic) <- "scenic"
Idents(dsx_scenic) <- params$celltype

# if (params$parallelize) {
#     plan(params$futures_plan, workers = params$futures_n_cores)
#     options(future.globals.maxSize = params$futures_ram * 1024^3)
# }
markers_celltype <- FindAllMarkers(object = dsx_scenic,
                                   slot = "scale.data",
                                   only.pos = TRUE,
                                   logfc.threshold = 0,
                                   min.pct = 0,
                                   test.use = params$diff_exp_test,
                                   latent.vars = latent_variables,
                                   verbose = TRUE)
# if (params$parallelize) {
#     plan("sequential")
# }

markers_celltype <- markers_celltype %>%
    dplyr::rename(cell_type = cluster, regulon = gene) #%>% 
    # filter(p_val_adj < 0.05)

# write_csv(markers_celltype,
#           paste0(params$results_directory, params$diff_exp_test,
#                  "_celltype_onevsall_comparison.csv")
#           )
```



```{r}
markers_celltype %>% 
    dplyr::filter(p_val_adj < 0.05) %>% 
    dplyr::filter(regulon == "rn(+)")
```

```{r}
markers_celltype %>% 
    dplyr::filter(p_val_adj < 0.05) %>% 
    dplyr::filter(regulon == "tup(+)")
```


```{r}
markers_celltype %>% 
    dplyr::filter(p_val_adj < 0.05) %>% 
    dplyr::filter(regulon == "bab1(+)")
```


```{r}
markers_celltype %>% 
    dplyr::filter(p_val_adj < 0.05) %>% 
    dplyr::filter(regulon == "Dif(+)")
```


```{r}
markers_celltype %>% 
    dplyr::filter(p_val_adj < 0.05) %>% 
    dplyr::filter(regulon == "RunxA(+)")
```


```{r}
markers_celltype %>% 
    dplyr::filter(p_val_adj < 0.05) %>% 
    dplyr::filter(regulon == "luna(+)")
```







```{r}
markers_celltype %>% 
    filter(regulon == "otp(+)")
    # filter(cell_type == "pC1")
```

```{r}
markers_celltype %>% 
    # filter(regulon == "otp(+)") 
    filter(cell_type == "pC1")
```



Condition - one vs. all

```{r warning=FALSE}
if (params$condition != "None") {

    # if (params$parallelize) {
    #     plan(params$futures_plan, workers = params$futures_n_cores)
    #     options(future.globals.maxSize = params$futures_ram * 1024^3)
    # }

    markers_condition_by_celltype_list <- list()
    DefaultAssay(dsx_scenic) <- "scenic"
    Idents(dsx_scenic) <- params$celltype
    celltypes <- sort(as.character(unique(dsx_scenic@meta.data[[params$celltype]])))

    condition_options <- sort(unique(dsx_scenic@meta.data[[params$condition]]))
    object_metadata <- FetchData(object = dsx_scenic,
                                 vars = c(params$condition, params$celltype)) %>%
        dplyr::rename(condition = params$condition, celltype = params$celltype) %>%
        dplyr::count(condition, celltype, .drop = FALSE)

    for (ii in seq_along(celltypes)) {

        num_condition_1_cells <- object_metadata %>% filter(condition == condition_options[[1]] & celltype == celltypes[ii]) %>% pull(n)
        num_condition_2_cells <- object_metadata %>% filter(condition == condition_options[[2]] & celltype == celltypes[ii]) %>% pull(n)
        min_num_cells <- min(num_condition_1_cells, num_condition_2_cells)

        if (min_num_cells > 3) {
            markers <- FindMarkers(object = dsx_scenic,
                                   slot = "scale.data",
                                   ident.1 = condition_options[[1]],
                                   ident.2 = condition_options[[2]],
                                   group.by = params$condition,
                                   subset.ident = celltypes[ii],
                                   test.use = params$diff_exp_test,
                                   latent.vars = latent_variables,
                                   logfc.threshold = 0,
                                   min.pct = 0,
                                   only.pos = FALSE,
                                   verbose = TRUE
                                   ) %>%
                rownames_to_column(var = "regulon") %>%
                mutate(cell_type = celltypes[ii])
            markers_condition_by_celltype_list[[celltypes[ii]]] <- markers
            gc()
        }
    }
    markers_condition_by_celltype <- bind_rows(markers_condition_by_celltype_list) %>%
        dplyr::rename(!!paste0("pct_", condition_options[[1]]) := "pct.1") %>%
        dplyr::rename(!!paste0("pct_", condition_options[[2]]) := "pct.2") %>%
        filter(p_val_adj < 0.05)

    # if (params$parallelize) {
    #     plan("sequential")
    # }

    # write_csv(markers_condition_by_celltype,
    #           paste0(params$results_directory, "/", params$diff_exp_test,
    #                  "_condition_by_celltype_comparison.csv")
    #           )
}
```








### Top n regulons per group - average expression per cluster or per condition

```{r }
Idents(dsx_scenic) <- params$celltype
```

```{r}

top_n_celltype <- markers_celltype %>%
    filter(str_detect(string = regulon, pattern = fixed("(-)"), negate = TRUE)) %>% 
    group_by(cell_type) %>%
    filter(p_val_adj < params$FDR_threshold) %>%
    slice_max(order_by = avg_diff, n = params$top_n)

if (params$condition != "None") {
	top_n_cellsubtype_condition <- markers_condition_by_celltype %>%
        filter(str_detect(string = regulon, pattern = fixed("(-)"), negate = TRUE)) %>% 
	    group_by(cell_type) %>%
	    filter(p_val_adj < params$FDR_threshold) %>%
	    slice_max(order_by = abs(avg_diff), n = params$top_n)
}

```






```{r fig.width=6, fig.height=6}

Idents(dsx_scenic) <- params$celltype

p <- DotPlot(
		object = dsx_scenic, 
		assay = "scenic",
		features = unique(top_n_celltype$regulon),
		cols = "RdYlBu", 
		col.max = 3, 
		col.min = -3,
		cluster.idents = FALSE
		) +
    # RotatedAxis() +
    coord_flip() +
    FontSize(x.text = 9, y.text = 10, x.title = 8, y.title = 8)

p <- p +
    theme(legend.text = element_text(size = 12),
          legend.title = element_text(size = 12),
    	  axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
    	  axis.title.x = element_blank()
    	  )
plot(p)

```





```{r fig.width=10, fig.height=4}

Idents(dsx_scenic) <- params$celltype

p <- DotPlot(
		object = dsx_scenic, 
		assay = "scenic",
		features = unique(top_n_celltype$regulon),
		cols = "RdYlBu", 
		col.max = 3, 
		col.min = -3,
		cluster.idents = FALSE
		) +
    # RotatedAxis() +
    # coord_flip() +
    FontSize(x.text = 9, y.text = 10, x.title = 8, y.title = 8)

p <- p +
    theme(legend.text = element_text(size = 12),
          legend.title = element_text(size = 12),
    	  axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
    	  axis.title.x = element_blank()
    	  )
plot(p)

```







```{r fig.width=6, fig.height=6}

Idents(dsx_scenic) <- params$celltype

p <- DotPlot(
		object = dsx_scenic, 
		assay = "scenic",
		features = unique(top_n_cellsubtype_condition$regulon),
		split.by = "sex",
		cols = "RdYlBu", 
		col.max = 3, 
		col.min = -3,
		cluster.idents = FALSE
		) +
    # RotatedAxis() +
    coord_flip() +
    FontSize(x.text = 9, y.text = 10, x.title = 8, y.title = 8)

p <- p +
    theme(legend.text = element_text(size = 12),
          legend.title = element_text(size = 12),
    	  axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
    	  axis.title.x = element_blank()
    	  )
plot(p)

```


```{r fig.width=10, fig.height=4}

Idents(dsx_scenic) <- params$celltype

p <- DotPlot(
		object = dsx_scenic, 
		assay = "scenic",
		features = unique(top_n_cellsubtype_condition$regulon),
		split.by = "sex",
		cols = "RdYlBu", 
		col.max = 3, 
		col.min = -3,
		cluster.idents = FALSE
		) +
    # RotatedAxis() +
    # coord_flip() +
    FontSize(x.text = 9, y.text = 10, x.title = 8, y.title = 8)

p <- p +
    theme(legend.text = element_text(size = 12),
          legend.title = element_text(size = 12),
    	  axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
    	  axis.title.x = element_blank()
    	  )
plot(p)

```


```{r}
top_n_cellsubtype_condition
```










# Filter Regulons



```{r}
mast_top15 <- markers_celltype %>% 
	filter(p_val_adj < 0.0001) %>% 
	group_by(cell_type) %>% 
	top_n(n = 15, wt = "avg_diff")
mast_top15
```


```{r}
top_regulons <- markers_celltype %>% 
	filter(p_val_adj < 0.0001) %>% 
	group_by(cell_type) %>% 
	top_n(n = 15, wt = "avg_diff") %>%
    pull(regulon) %>%
	unique() %>% 
    sort()
top_regulons
```




```{r}

high_confidence_regulons <- read_csv(file = "../../proj136/analyses/scenic/scenic_multi_dsx/results.dir/aggregated.dir/raw.dir/dsx.dir/high_confidence_regulons.csv")
high_confidence_regulons

```




```{r}
cell_type_high_confidence_regulons <- high_confidence_regulons %>% 
    filter(!str_detect(string = regulon_name, pattern = fixed("(-)"))) %>% 
    filter(regulon_name %in% top_regulons)
cell_type_high_confidence_regulons
```



```{r}
cell_type_high_confidence_regulons <- cell_type_high_confidence_regulons %>% 
    mutate(regulon_name_new = str_remove(string = regulon_name, pattern = fixed("(+)")))
cell_type_high_confidence_regulons
```


```{r}
cell_type_high_confidence_regulons %>% 
    filter(gene_name == "sv")
```



```{r}
cell_type_high_confidence_regulons %>% 
    filter(regulon_name_new == "otp")
```


```{r}
cell_type_high_confidence_regulons %>% 
    filter(regulon_name_new == "bsh")
```


```{r}
cell_type_high_confidence_regulons %>% 
    filter(regulon_name_new == "TfAP-2")
```




```{r}
cell_type_high_confidence_regulons %>% 
    filter(regulon_name_new == "sv")
```



```{r}
cell_type_high_confidence_regulons %>% 
    group_by(regulon_name_new) %>% 
    tally() %>% 
    arrange(desc(n))
```






```{r}
dsx_cluster_markers <- read_csv(file = "../../proj136/analyses/markers/subclustering_dsx_seu--round1--500VarGene--NoReg--Harmony_ori_exp_cell--RNA_snn_res_2__cluster_markers.csv")
dsx_cluster_markers
```

```{r}
top_markers <- dsx_cluster_markers %>% 
    filter(p_val_adj < 0.05) %>%
    # filter(avg_log2FC > 1) %>% 
    pull(gene) %>% 
    unique() %>% 
    sort()
# top_markers
```


```{r}
cell_type_high_confidence_regulons <- cell_type_high_confidence_regulons %>% 
    filter(regulon_name_new %in% top_markers)
cell_type_high_confidence_regulons
```


```{r}
cell_type_high_confidence_regulons %>% 
    group_by(regulon_name_new) %>% 
    tally() %>% 
    arrange(desc(n))
```

```{r}
tfs <- read_tsv("../../../aallen/gene_lists/gene_sets/TFs__6_30.tsv")
tfs
```


```{r}

cell_type_high_confidence_regulons <- cell_type_high_confidence_regulons %>% 
    filter(regulon_name_new %in% tfs$gene)
cell_type_high_confidence_regulons

```


```{r}
cell_type_high_confidence_regulons %>% 
    group_by(regulon_name_new) %>% 
    tally() %>% 
    arrange(desc(n))
```


```{r}
cell_type_high_confidence_regulons_for_cytoscape <- cell_type_high_confidence_regulons %>% 
    dplyr::select(-regulon_name) %>% 
    dplyr::rename("source" = "regulon_name_new") %>% 
    dplyr::rename("target" = "gene_name") %>% 
    dplyr::relocate(source, target)
cell_type_high_confidence_regulons_for_cytoscape
```




```{r}
regulon_keep <- cell_type_high_confidence_regulons %>% 
    group_by(regulon_name_new) %>% 
    tally() %>% 
    filter(n < 100) %>% 
    pull(regulon_name_new)

cell_type_high_confidence_regulons_for_cytoscape_slim <- cell_type_high_confidence_regulons %>% 
    filter(regulon_name_new %in% regulon_keep) %>% 
    dplyr::select(-regulon_name) %>% 
    dplyr::rename("source" = "regulon_name_new") %>% 
    dplyr::rename("target" = "gene_name") %>% 
    dplyr::relocate(source, target)
cell_type_high_confidence_regulons_for_cytoscape_slim
```


```{r}
write_csv(x = cell_type_high_confidence_regulons_for_cytoscape, file = "../../proj136/analyses/scenic/scenic_multi_dsx/results.dir/aggregated.dir/raw.dir/dsx.dir/dsx_cell_type_high_confidence_regulons_for_cytoscape.csv")
write_csv(x = cell_type_high_confidence_regulons_for_cytoscape_slim, file = "../../proj136/analyses/scenic/scenic_multi_dsx/results.dir/aggregated.dir/raw.dir/dsx.dir/dsx_cell_type_high_confidence_regulons_for_cytoscape_slim.csv")
```






```{r}
max_umi <- apply(dsx_scenic@assays$RNA@counts, 1, function(x) max(x))
max_umi <- as.data.frame(max_umi)
max_umi <- rownames_to_column(max_umi)
max_umi
```



```{r}
sum_umi <- apply(dsx_scenic@assays$RNA@counts, 1, function(x) sum(x))
sum_umi <- as.data.frame(sum_umi)
sum_umi <- rownames_to_column(sum_umi)
sum_umi
```


```{r}
gene_exp <- dplyr::inner_join(max_umi, sum_umi, by = "rowname")
gene_exp
```


```{r}

genes <- c("TfAP-2", "otp", "opa", "Vsx1", "Vsx2", "bab1", "bab2", "drm", "Drgx", "dsx", "tup", "B-H1", "B-H2")

gene_exp %>% filter(rowname %in% genes)

```




```{r}
filtered_genes <- gene_exp %>% 
	filter(max_umi > 2) %>%
	filter(sum_umi > 80) %>%
	pull(rowname)
# filtered_genes
```



```{r}
cell_type_high_confidence_regulons
```


```{r}

cell_type_high_confidence_regulons__exp_genes <- cell_type_high_confidence_regulons %>% 
    filter(regulon_name_new %in% filtered_genes) %>% 
    filter(gene_name %in% filtered_genes)
cell_type_high_confidence_regulons__exp_genes

```


```{r}
cell_type_high_confidence_regulons__exp_genes %>% 
    group_by(regulon_name_new) %>% 
    tally() %>% 
    arrange(desc(n))
```


```{r}
cell_type_high_confidence_regulons__exp_genes_for_cytoscape <- cell_type_high_confidence_regulons__exp_genes %>% 
    dplyr::select(-regulon_name) %>% 
    dplyr::rename("source" = "regulon_name_new") %>% 
    dplyr::rename("target" = "gene_name") %>% 
    dplyr::relocate(source, target)
cell_type_high_confidence_regulons__exp_genes_for_cytoscape
```




```{r}
regulon_keep <- cell_type_high_confidence_regulons__exp_genes %>% 
    group_by(regulon_name_new) %>% 
    tally() %>% 
    filter(n < 100) %>% 
    pull(regulon_name_new)

cell_type_high_confidence_regulons__exp_genes_for_cytoscape_slim <- cell_type_high_confidence_regulons__exp_genes %>% 
    filter(regulon_name_new %in% regulon_keep) %>% 
    dplyr::select(-regulon_name) %>% 
    dplyr::rename("source" = "regulon_name_new") %>% 
    dplyr::rename("target" = "gene_name") %>% 
    dplyr::relocate(source, target)
cell_type_high_confidence_regulons__exp_genes_for_cytoscape_slim
```


```{r}
write_csv(x = cell_type_high_confidence_regulons__exp_genes_for_cytoscape, file = "../../proj136/analyses/scenic/scenic_multi_dsx/results.dir/aggregated.dir/raw.dir/dsx.dir/dsx_cell_type_high_confidence_regulons__exp_genes_for_cytoscape.csv")
write_csv(x = cell_type_high_confidence_regulons__exp_genes_for_cytoscape_slim, file = "../../proj136/analyses/scenic/scenic_multi_dsx/results.dir/aggregated.dir/raw.dir/dsx.dir/dsx_cell_type_high_confidence_regulons__exp_genes_for_cytoscape_slim.csv")
```








### otp regulon





```{r warning=FALSE, message=FALSE, fig.width=12, fig.height=12}

DimPlot(object = dsx_scenic, reduction = "scenic_tsne_30pcs", group.by = "cell_type", pt.size = 2, label = TRUE) + scale_colour_discrete_qualitative(palette = "Set 2") + coord_fixed()
DimPlot(object = dsx_scenic, reduction = "scenic_tsne_30pcs", group.by = "RNA_snn_res.2", pt.size = 2, label = TRUE) + scale_colour_discrete_qualitative(palette = "Set 2") + coord_fixed()

```

18,14,0,13,3,7,12,11


```{r}
pC1_markers <- dsx_cluster_markers %>% 
    filter(p_val_adj < 0.05) %>%
    filter(cluster %in% c(18,14,0,13,3,7,12,11)) %>% 
    # filter(avg_log2FC > 1) %>% 
    pull(gene) %>% 
    unique() %>% 
    sort()


```


```{r}
cell_type_high_confidence_regulons__exp_genes_for_cytoscape_slim %>% 
    filter(source == "otp") %>% 
    mutate(marker = if_else(target %in% pC1_markers, "yes", "no"))
```


top_markers

```{r}
top_markers <- dsx_cluster_markers %>% 
    filter(p_val_adj < 0.05) %>%
    # filter(avg_log2FC > 1) %>% 
    pull(gene) %>% 
    unique() %>% 
    sort()
# top_markers
```


```{r}
cell_type_high_confidence_regulons <- cell_type_high_confidence_regulons %>% 
    filter(regulon_name_new %in% top_markers)
cell_type_high_confidence_regulons
```













# pC1 focus





## Re-Run markers


```{r}

params <- list()

params$datatype <- "raw"

params$condition <- "sex"
params$celltype <- "RNA_snn_res.2"

params$reference_condition <- "female"
params$latent_variables <- "experiment"
params$diff_exp_test <- "wilcox"
params$FDR_threshold <- 0.05
params$top_n <- 5

params$results_directory <- "../../proj136/analyses/scenic/scenic_multi_dsx/results.dir/aggregated.dir/raw.dir/dsx.dir/pyscenic_r.dir/"



```




```{r warning=FALSE, message=FALSE, fig.width=12, fig.height=12}

DimPlot(object = dsx_scenic, reduction = "scenic_tsne_30pcs", group.by = "cell_type", pt.size = 2, label = TRUE) + scale_colour_discrete_qualitative(palette = "Set 2") + coord_fixed()
DimPlot(object = dsx_scenic, reduction = "scenic_tsne_30pcs", group.by = "RNA_snn_res.2", pt.size = 2, label = TRUE) + scale_colour_discrete_qualitative(palette = "Set 2") + coord_fixed()

```

```{r}
Idents(dsx_scenic) <- "RNA_snn_res.2"
pC1_scenic <- subset(x = dsx_scenic, idents = c(18,14,0,13,3,7,12,11))
pC1_scenic
```




#### Stats to identify regulons that are differentially-expressed between conditions or cell types or celltypes_conditions

```{r diff-reg-celltype}
# Latent variables for test
if (params$latent_variables == "None" | params$diff_exp_test == "wilcox") {
    latent_variables <- NULL
} else {
    latent_variables <- params$latent_variables
}

# Cell type - one vs. all
DefaultAssay(pC1_scenic) <- "scenic"
Idents(pC1_scenic) <- params$celltype

# if (params$parallelize) {
#     plan(params$futures_plan, workers = params$futures_n_cores)
#     options(future.globals.maxSize = params$futures_ram * 1024^3)
# }
markers_pC1_celltype <- FindAllMarkers(object = pC1_scenic,
                                   slot = "scale.data",
                                   only.pos = TRUE,
                                   logfc.threshold = 0,
                                   min.pct = 0,
                                   test.use = params$diff_exp_test,
                                   latent.vars = latent_variables,
                                   verbose = TRUE)
# if (params$parallelize) {
#     plan("sequential")
# }

markers_pC1_celltype <- markers_pC1_celltype %>%
    dplyr::rename(cell_type = cluster, regulon = gene) %>%
    filter(p_val_adj < 0.05)

# write_csv(markers_pC1_celltype,
#           paste0(params$results_directory, params$diff_exp_test,
#                  "_celltype_onevsall_comparison.csv")
#           )
```


```{r}
markers_pC1_celltype %>% 
    filter(regulon == "otp(+)")
    # filter(cell_type == "pC1")
```

```{r}
markers_pC1_celltype %>% 
    # filter(regulon == "otp(+)") 
    filter(cell_type == "pC1")
```




```{r}
if (params$condition != "None") {
    Idents(pC1_scenic) <- params$condition

    # if (params$parallelize) {
    #     plan(params$futures_plan, workers = params$futures_n_cores)
    #     options(future.globals.maxSize = params$futures_ram * 1024^3)
    # }
    markers_condition <-  FindAllMarkers(object = pC1_scenic,
                                         slot = "scale.data",
                                         only.pos = TRUE,
                                         logfc.threshold = 0,
                                         min.pct = 0,
                                         test.use = params$diff_exp_test,
                                         latent.vars = latent_variables,
                                         verbose = FALSE)
    # if (params$parallelize) {
    #     plan("sequential")
    # }

    markers_condition <- markers_condition %>%
        dplyr::rename(condition = cluster, regulon = gene) %>%
        filter(p_val_adj < 0.05)

    # write_csv(markers_condition,
    #           paste0(results_directory, "/", params$diff_exp_test,
    #                  "_condition_onevsall_comparison.csv")
    #           )
}
```


```{r warning=FALSE}
if (params$condition != "None") {

    # if (params$parallelize) {
    #     plan(params$futures_plan, workers = params$futures_n_cores)
    #     options(future.globals.maxSize = params$futures_ram * 1024^3)
    # }

    markers_pC1_condition_by_celltype_list <- list()
    DefaultAssay(pC1_scenic) <- "scenic"
    Idents(pC1_scenic) <- params$celltype
    celltypes <- sort(as.character(unique(pC1_scenic@meta.data[[params$celltype]])))

    condition_options <- sort(unique(pC1_scenic@meta.data[[params$condition]]))
    object_metadata <- FetchData(object = pC1_scenic,
                                 vars = c(params$condition, params$celltype)) %>%
        dplyr::rename(condition = params$condition, celltype = params$celltype) %>%
        dplyr::count(condition, celltype, .drop = FALSE)

    for (ii in seq_along(celltypes)) {

        num_condition_1_cells <- object_metadata %>% filter(condition == condition_options[[1]] & celltype == celltypes[ii]) %>% pull(n)
        num_condition_2_cells <- object_metadata %>% filter(condition == condition_options[[2]] & celltype == celltypes[ii]) %>% pull(n)
        min_num_cells <- min(num_condition_1_cells, num_condition_2_cells)

        if (min_num_cells > 3) {
            markers <- FindMarkers(object = pC1_scenic,
                                   slot = "scale.data",
                                   ident.1 = condition_options[[1]],
                                   ident.2 = condition_options[[2]],
                                   group.by = params$condition,
                                   subset.ident = celltypes[ii],
                                   test.use = params$diff_exp_test,
                                   latent.vars = latent_variables,
                                   logfc.threshold = 0,
                                   min.pct = 0,
                                   only.pos = FALSE,
                                   verbose = TRUE
                                   ) %>%
                rownames_to_column(var = "regulon") %>%
                mutate(cell_type = celltypes[ii])
            markers_pC1_condition_by_celltype_list[[celltypes[ii]]] <- markers
            gc()
        }
    }
    markers_pC1_condition_by_celltype <- bind_rows(markers_pC1_condition_by_celltype_list) %>%
        dplyr::rename(!!paste0("pct_", condition_options[[1]]) := "pct.1") %>%
        dplyr::rename(!!paste0("pct_", condition_options[[2]]) := "pct.2") %>%
        filter(p_val_adj < 0.05)

    # if (params$parallelize) {
    #     plan("sequential")
    # }

    # write_csv(markers_pC1_condition_by_celltype,
    #           paste0(params$results_directory, "/", params$diff_exp_test,
    #                  "_condition_by_celltype_comparison.csv")
    #           )
}
```








### Top n regulons per group - average expression per cluster or per condition

```{r }
Idents(pC1_scenic) <- params$celltype
```

```{r}

top_n_celltype <- markers_pC1_celltype %>%
    filter(str_detect(string = regulon, pattern = fixed("(-)"), negate = TRUE)) %>% 
    group_by(cell_type) %>%
    filter(p_val_adj < params$FDR_threshold) %>%
    slice_max(order_by = avg_diff, n = params$top_n)

if (params$condition != "None") {
	top_n_condition <- markers_condition %>%
	    group_by(condition) %>%
	    filter(p_val_adj < params$FDR_threshold) %>%
	    slice_max(order_by = avg_diff, n = params$top_n)
	
	top_n_cellsubtype_condition <- markers_pC1_condition_by_celltype %>%
        filter(str_detect(string = regulon, pattern = fixed("(-)"), negate = TRUE)) %>% 
	    group_by(cell_type) %>%
	    filter(p_val_adj < params$FDR_threshold) %>%
	    slice_max(order_by = abs(avg_diff), n = params$top_n)
}

```






```{r fig.width=6, fig.height=6}

Idents(pC1_scenic) <- params$celltype

p <- DotPlot(
		object = pC1_scenic, 
		assay = "scenic",
		features = unique(top_n_celltype$regulon),
		cols = "RdYlBu", 
		col.max = 3, 
		col.min = -3,
		cluster.idents = FALSE
		) +
    # RotatedAxis() +
    coord_flip() +
    FontSize(x.text = 9, y.text = 10, x.title = 8, y.title = 8)

p <- p +
    theme(legend.text = element_text(size = 12),
          legend.title = element_text(size = 12),
    	  axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
    	  axis.title.x = element_blank()
    	  )
plot(p)

```





```{r fig.width=10, fig.height=4}

Idents(pC1_scenic) <- params$celltype

p <- DotPlot(
		object = pC1_scenic, 
		assay = "scenic",
		features = unique(top_n_celltype$regulon),
		cols = "RdYlBu", 
		col.max = 3, 
		col.min = -3,
		cluster.idents = FALSE
		) +
    # RotatedAxis() +
    # coord_flip() +
    FontSize(x.text = 9, y.text = 10, x.title = 8, y.title = 8)

p <- p +
    theme(legend.text = element_text(size = 12),
          legend.title = element_text(size = 12),
    	  axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
    	  axis.title.x = element_blank()
    	  )
plot(p)

```







```{r fig.width=6, fig.height=6}

Idents(pC1_scenic) <- params$celltype

p <- DotPlot(
		object = pC1_scenic, 
		assay = "scenic",
		features = unique(top_n_cellsubtype_condition$regulon),
		split.by = "sex",
		cols = "RdYlBu", 
		col.max = 3, 
		col.min = -3,
		cluster.idents = FALSE
		) +
    # RotatedAxis() +
    coord_flip() +
    FontSize(x.text = 9, y.text = 10, x.title = 8, y.title = 8)

p <- p +
    theme(legend.text = element_text(size = 12),
          legend.title = element_text(size = 12),
    	  axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
    	  axis.title.x = element_blank()
    	  )
plot(p)

```


```{r fig.width=10, fig.height=4}

Idents(pC1_scenic) <- params$celltype

p <- DotPlot(
		object = pC1_scenic, 
		assay = "scenic",
		features = unique(top_n_cellsubtype_condition$regulon),
		split.by = "sex",
		cols = "RdYlBu", 
		col.max = 3, 
		col.min = -3,
		cluster.idents = FALSE
		) +
    # RotatedAxis() +
    # coord_flip() +
    FontSize(x.text = 9, y.text = 10, x.title = 8, y.title = 8)

p <- p +
    theme(legend.text = element_text(size = 12),
          legend.title = element_text(size = 12),
    	  axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
    	  axis.title.x = element_blank()
    	  )
plot(p)

```


```{r}
top_n_cellsubtype_condition
```






# Session Info

```{r}
sessionInfo()
```






