---
title: "Sexed_Paper_2_Analysis__Figure_3_supp_Kenyon_cell_re_subclustering"
description: description
author: "Aaron M. Allen"
date: 'Last update: `r date()`'
output:
  html_document:
    number_sections: true
    code_folding: show
    code_download: true
    theme: cerulean
    df_print: paged
    fig_width: 8.5
    fig_height: 5
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 6
---



```{css, echo = FALSE}
pre, code {white-space:pre !important; overflow-x:auto}
.tocify-item {white-space:pre}
```



```{r echo = FALSE, cache = FALSE}

```


```{r setup, echo = FALSE, cache = FALSE}
ggplot2::theme_set(ggplot2::theme_grey())
ggplot2::theme_update(panel.background = ggplot2::element_rect(fill = "transparent", colour = NA),
                      plot.background = ggplot2::element_rect(fill = "transparent", colour = NA))

rmd_name <- gsub(pattern = ".Rmd", replacement = "", x = knitr::current_input(), ignore.case = TRUE)
knitr::opts_chunk$set(dev = c("png", "cairo_pdf"),
                      dev.args=list(bg="transparent"),
                      fig.align = "center",
                      fig.height = 5,
                      fig.width = 8.5,
                      dpi = 300,
                      pdf.options(encoding = "ISOLatin9.enc"),
                      fig.path = paste0("../analyses/figures/", rmd_name, "-",
                                        format(Sys.time(), "%Y%m%d_%H%M%S"), "/"),
                      fig.retina = 1,
                      warning = TRUE,
                      message = TRUE)
```



# Setup


## Start time

```{r start-time}
set.seed(12021)
today <- format(Sys.time(), "%Y%m%d_%H%M%S")
total_start_time <- Sys.time()
total_start_time
```






## Parameters





## Libraries

```{r packages-etc, warning = FALSE, message = FALSE}

library(Seurat)
library(cowplot)
library(pheatmap)
library(colorspace)
library(tidyverse)

```



## Load Data





```{r load-meta-brain}
meta_cb <- read_rds("../../proj136/analyses/rds_files/New_Meta_cbNeuron_PUBnofacs_noPolIII_v1_alevin_LogNormalize_harmony_integrated_seurat_doublets_removed_diet.rds")
meta_cb
```




```{r}
KC_seu <- read_rds("../../proj136/analyses/rds_files/New_Meta_cbNeuron_PUBnofacs_noPolIII_v1--1000VarGene--NoReg--Harmony_ori_exp_cell--_alevin_LogNormalize_harmony_integrated_seurat_doublets_removed_Kenyon_cell_subclusting.rds")
KC_seu
```

```{r}
KC_seu@meta.data %>% colnames()
```



```{r fig.width=15, fig.height=10}
DimPlot(object = KC_seu, 
		group.by = "RNA_snn_res.1",
        reduction = "umap_20pcs", 
        label = TRUE, 
        label.box = FALSE) + 
    NoLegend() + 
    coord_fixed()
DimPlot(object = KC_seu, 
		group.by = "RNA_snn_res.4",
        reduction = "umap_20pcs", 
        label = TRUE, 
        label.box = FALSE) + 
    NoLegend() + 
    coord_fixed()
DimPlot(object = KC_seu, 
		group.by = "RNA_snn_res.10",
        reduction = "umap_20pcs", 
        label = TRUE, 
        label.box = FALSE) + 
    NoLegend() + 
    coord_fixed()
```

5,6,16,singleton
0,1,25,singleton


```{r fig.width=15, fig.height=10}
DimPlot(object = KC_seu, 
		group.by = "experiment",
        reduction = "umap_20pcs", 
        label = TRUE, 
        label.box = FALSE) + 
    NoLegend() + 
    coord_fixed()
```





```{r fig.width=10}
FetchData(object = KC_seu, vars = c("experiment","RNA_snn_res.1")) %>% 
    dplyr::count(experiment, RNA_snn_res.1) %>% 
    ggplot(aes(x = RNA_snn_res.1, y = n, fill = experiment)) +
        geom_bar(stat = "identity", position = "fill")
```






```{r fig.width=12, fig.height=12}

genes <- c("percent_mt", "percent_rRNA", "percent_rProt", "percent_hsp" )
plist <- FeaturePlot(
    object = KC_seu, 
    reduction = "umap_20pcs", 
    features = genes,
    combine = FALSE,
    pt.size = 1, 
    # min.cutoff = 0,
    # max.cutoff = 3,
    # slot = "counts",
    cols = c("lightsteelblue2", "black"), 
    raster = FALSE,
    raster.dpi = c(1024, 1024),
    order = TRUE,
    coord.fixed = TRUE
)

for (i in seq_along(plist)) {
	# plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}

```



```{r}
rp_genes <- KC_seu %>% rownames() %>% str_subset("Rp[S|L]")
rp_genes
```


```{r}
KC_seu <- AddModuleScore(object = KC_seu, features = list(rp_genes), name = "modulescore_rp")
```


```{r fig.width=12, fig.height=12}

genes <- c("modulescore_rp1")
plist <- FeaturePlot(
    object = KC_seu, 
    reduction = "umap_20pcs", 
    features = genes,
    combine = FALSE,
    pt.size = 1, 
    # min.cutoff = 0,
    # max.cutoff = 3,
    # slot = "counts",
    cols = c("lightsteelblue2", "black"), 
    raster = FALSE,
    raster.dpi = c(1024, 1024),
    order = TRUE,
    coord.fixed = TRUE
)

for (i in seq_along(plist)) {
	# plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}

```

```{r}
mt_genes <- KC_seu %>% rownames() %>% str_subset("^mt:")
mt_genes
```


```{r}
KC_seu <- AddModuleScore(object = KC_seu, features = list(mt_genes), name = "modulescore_mt")
```


```{r fig.width=12, fig.height=12}

genes <- c("modulescore_mt1")
plist <- FeaturePlot(
    object = KC_seu, 
    reduction = "umap_20pcs", 
    features = genes,
    combine = FALSE,
    pt.size = 1, 
    # min.cutoff = 0,
    # max.cutoff = 3,
    # slot = "counts",
    cols = c("lightsteelblue2", "black"), 
    raster = FALSE,
    raster.dpi = c(1024, 1024),
    order = TRUE,
    coord.fixed = TRUE
)

for (i in seq_along(plist)) {
	# plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}

```




```{r}
hsp_genes <- KC_seu %>% rownames() %>% str_subset("Hsp")
hsp_genes
```

```{r}
KC_seu <- AddModuleScore(object = KC_seu, features = list(hsp_genes), name = "modulescore_hsp")
```


```{r fig.width=12, fig.height=12}

genes <- c("modulescore_hsp1")
plist <- FeaturePlot(
    object = KC_seu, 
    reduction = "umap_20pcs", 
    features = genes,
    combine = FALSE,
    pt.size = 1, 
    # min.cutoff = 0,
    # max.cutoff = 3,
    # slot = "counts",
    cols = c("lightsteelblue2", "black"), 
    raster = FALSE,
    raster.dpi = c(1024, 1024),
    order = TRUE,
    coord.fixed = TRUE
)

for (i in seq_along(plist)) {
	# plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}

```



```{r fig.width=12, fig.height=12}

genes <- hsp_genes
plist <- FeaturePlot(
    object = KC_seu, 
    reduction = "umap_20pcs", 
    features = genes,
    combine = FALSE,
    pt.size = 1, 
    # min.cutoff = 0,
    # max.cutoff = 3,
    # slot = "counts",
    cols = c("lightsteelblue2", "black"), 
    raster = FALSE,
    raster.dpi = c(1024, 1024),
    order = TRUE,
    coord.fixed = TRUE
)

for (i in seq_along(plist)) {
	# plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}

```







```{r}
# 
# job::job({
#     library(future)
#     plan("multisession", workers = 8)
#     options(future.globals.maxSize = 8*1024*1024^2)
#     
#     Idents(KC_seu) <- "RNA_snn_res.1"
#     markers_KC_res1 <- FindAllMarkers(object = KC_seu, logfc.threshold = 0.5, only.pos = TRUE, verbose = TRUE)
#     
#     job::export(c(markers_KC_res1))
# }, import = c(KC_seu))

```






```{r}
# markers_KC_res1
```



```{r}
# markers_KC_res1 %>% 
#     filter(cluster == 12) %>% 
#     filter(p_val_adj < 0.05) %>% 
#     arrange(desc(avg_log2FC))
```



```{r fig.width=12, fig.height=12}

genes <- c("rn", "CG9008", "CG43795", "Apoltp", "CG43759", "kek2")
plist <- FeaturePlot(
    object = KC_seu, 
    reduction = "umap_20pcs", 
    features = genes,
    combine = FALSE,
    pt.size = 1, 
    # min.cutoff = 0,
    # max.cutoff = 3,
    # slot = "counts",
    cols = c("lightsteelblue2", "black"), 
    raster = FALSE,
    raster.dpi = c(1024, 1024),
    order = TRUE,
    coord.fixed = TRUE
)

for (i in seq_along(plist)) {
	# plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}

```




```{r fig.width=12, fig.height=12}

genes <- c("rn", "CG9008", "CG43795", "Apoltp")
plist <- FeaturePlot(
    object = meta_cb, 
    reduction = "tsne_240pcs", 
    features = genes,
    combine = FALSE,
    pt.size = 0.4, 
    min.cutoff = 0,
    max.cutoff = 3,
    # slot = "counts",
    cols = c("lightsteelblue2", "black"), 
    raster = FALSE,
    raster.dpi = c(1024, 1024),
    order = TRUE,
    coord.fixed = TRUE
)

for (i in seq_along(plist)) {
	plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}

```




```{r}
# markers_KC_res1 %>% 
#     filter(cluster == 17) %>% 
#     filter(p_val_adj < 0.05) %>% 
#     arrange(desc(avg_log2FC))
```



```{r fig.width=12, fig.height=12}

genes <- c("CG8641", "mamo", "DAT")
plist <- FeaturePlot(
    object = KC_seu, 
    reduction = "umap_20pcs", 
    features = genes,
    combine = FALSE,
    pt.size = 1, 
    # min.cutoff = 0,
    # max.cutoff = 3,
    # slot = "counts",
    cols = c("lightsteelblue2", "black"), 
    raster = FALSE,
    raster.dpi = c(1024, 1024),
    order = TRUE,
    coord.fixed = TRUE
)

for (i in seq_along(plist)) {
	# plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}

```


```{r}
# markers_KC_res1 %>% 
#     filter(cluster == 5) %>% 
#     filter(p_val_adj < 0.05) %>% 
#     arrange(desc(avg_log2FC))
```



```{r fig.width=12, fig.height=12}

genes <- c("CG17716", "side", "mbl", "ct", "Octbeta2R", "kek2", "SoxN", "bi", "beat-Ic")
plist <- FeaturePlot(
    object = KC_seu, 
    reduction = "umap_20pcs", 
    features = genes,
    combine = FALSE,
    pt.size = 1, 
    # min.cutoff = 0,
    # max.cutoff = 3,
    # slot = "counts",
    cols = c("lightsteelblue2", "black"), 
    raster = FALSE,
    raster.dpi = c(1024, 1024),
    order = TRUE,
    coord.fixed = TRUE
)

for (i in seq_along(plist)) {
	# plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}

```




```{r fig.width=12, fig.height=12}

genes <- c("CG17716", "side", "mbl", "ct", "Octbeta2R", "kek2", "SoxN", "bi", "beat-Ic")
plist <- FeaturePlot(
    object = meta_cb, 
    reduction = "tsne_240pcs", 
    features = genes,
    combine = FALSE,
    pt.size = 0.4, 
    min.cutoff = 0,
    max.cutoff = 3,
    # slot = "counts",
    cols = c("lightsteelblue2", "black"), 
    raster = FALSE,
    raster.dpi = c(1024, 1024),
    order = TRUE,
    coord.fixed = TRUE
)

for (i in seq_along(plist)) {
	plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}

```


```{r fig.width=12, fig.height=12}

genes <- c("br", "bi")
plist <- FeaturePlot(
    object = meta_cb, 
    reduction = "tsne_240pcs", 
    features = genes,
    combine = FALSE,
    pt.size = 0.4, 
    min.cutoff = 0,
    max.cutoff = 3,
    # slot = "counts",
    cols = c("lightsteelblue2", "black"), 
    raster = FALSE,
    raster.dpi = c(1024, 1024),
    order = TRUE,
    coord.fixed = TRUE
)

for (i in seq_along(plist)) {
	plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}

```







Kenyon_cell	comment to come later	ey	Pka-R2	Dop1R2	Hr51	CG34402	Gdap2	prt	CG1129	cora
ab_KC	alpha and beta lobe subset of kenyon cells	lbm	crb	Cks85A	Tsp42En	Eip93F
prime_KC	alpha' and beta' lobe subset of kenyon cells	CG8641	Gr77a	Lac	CG42322
y_KC	gamma lobe subset of kenyon cells	CG13055	Lgr3	GstD11	CAH2	CG32204	ab




```{r fig.width=12, fig.height=12}

genes <- c("ey", "Pka-R2", "Dop1R2", "Hr51", "CG34402", "Gdap2", "prt", "CG1129", "cora")
plist <- FeaturePlot(
    object = KC_seu, 
    reduction = "umap_20pcs", 
    features = genes,
    combine = FALSE,
    pt.size = 1, 
    # min.cutoff = 0,
    # max.cutoff = 3,
    # slot = "counts",
    cols = c("lightsteelblue2", "black"), 
    raster = FALSE,
    raster.dpi = c(1024, 1024),
    order = TRUE,
    coord.fixed = TRUE
)

for (i in seq_along(plist)) {
	# plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}

```




```{r fig.width=12, fig.height=12}

genes <- c("lbm", "crb", "Cks85A", "Tsp42En", "Eip93F")
plist <- FeaturePlot(
    object = KC_seu, 
    reduction = "umap_20pcs", 
    features = genes,
    combine = FALSE,
    pt.size = 1, 
    # min.cutoff = 0,
    # max.cutoff = 3,
    # slot = "counts",
    cols = c("lightsteelblue2", "black"), 
    raster = FALSE,
    raster.dpi = c(1024, 1024),
    order = TRUE,
    coord.fixed = TRUE
)

for (i in seq_along(plist)) {
	# plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}

```








```{r fig.width=12, fig.height=12}

genes <- c("CG8641", "Gr77a", "Lac", "CG42322")
plist <- FeaturePlot(
    object = KC_seu, 
    reduction = "umap_20pcs", 
    features = genes,
    combine = FALSE,
    pt.size = 1, 
    # min.cutoff = 0,
    # max.cutoff = 3,
    # slot = "counts",
    cols = c("lightsteelblue2", "black"), 
    raster = FALSE,
    raster.dpi = c(1024, 1024),
    order = TRUE,
    coord.fixed = TRUE
)

for (i in seq_along(plist)) {
	# plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}

```










```{r fig.width=12, fig.height=12}

genes <- c("CG13055", "Lgr3", "GstD11", "CAH2", "CG32204", "ab")
plist <- FeaturePlot(
    object = KC_seu, 
    reduction = "umap_20pcs", 
    features = genes,
    combine = FALSE,
    pt.size = 1, 
    # min.cutoff = 0,
    # max.cutoff = 3,
    # slot = "counts",
    cols = c("lightsteelblue2", "black"), 
    raster = FALSE,
    raster.dpi = c(1024, 1024),
    order = TRUE,
    coord.fixed = TRUE
)

for (i in seq_along(plist)) {
	# plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}

```




```{r fig.width=12, fig.height=12}

genes <- c("Imp", "dati", "br", "Eip93F")
plist <- FeaturePlot(
    object = KC_seu, 
    reduction = "umap_20pcs", 
    features = genes,
    combine = FALSE,
    pt.size = 1, 
    # min.cutoff = 0,
    # max.cutoff = 3,
    # slot = "counts",
    cols = c("lightsteelblue2", "black"), 
    raster = FALSE,
    raster.dpi = c(1024, 1024),
    order = TRUE,
    coord.fixed = TRUE
)

for (i in seq_along(plist)) {
	# plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}

```



```{r fig.width=12, fig.height=12}

genes <- c("Gad1", "VGlut", "Vmat", "repo")
plist <- FeaturePlot(
    object = KC_seu, 
    reduction = "umap_20pcs", 
    features = genes,
    combine = FALSE,
    pt.size = 1, 
    # min.cutoff = 0,
    # max.cutoff = 3,
    # slot = "counts",
    cols = c("lightsteelblue2", "black"), 
    raster = FALSE,
    raster.dpi = c(1024, 1024),
    order = TRUE,
    coord.fixed = TRUE
)

for (i in seq_along(plist)) {
	# plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}

```



```{r fig.width=12, fig.height=12}

genes <- c("hec", "Dh31-R")
plist <- FeaturePlot(
    object = KC_seu, 
    reduction = "umap_20pcs", 
    features = genes,
    combine = FALSE,
    pt.size = 1, 
    # min.cutoff = 0,
    # max.cutoff = 3,
    # slot = "counts",
    cols = c("lightsteelblue2", "black"), 
    raster = FALSE,
    raster.dpi = c(1024, 1024),
    order = TRUE,
    coord.fixed = TRUE
)

for (i in seq_along(plist)) {
	# plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}

```









5,6,16,singleton
0,1,25,singleton

```{r}
KC_cells <- FetchData(
        object = KC_seu, 
        vars = c("RNA_snn_res.1", "RNA_snn_res.4")) %>%
    filter(!RNA_snn_res.1 %in% c("5","6","16","17","singleton")) %>%
    filter(!RNA_snn_res.4 %in% c("0","1","25","30", "32","singleton")) %>%
	rownames()
head(KC_cells)
length(KC_cells)
```




```{r fig.width=14, fig.height=14}

p1 <- DimPlot(object = KC_seu,
			  cells.highlight = KC_cells,
			  sizes.highlight = 0.4, 
			  cols.highlight = "black", 
			  cols = "lightsteelblue2",
        reduction = "umap_20pcs",
        pt.size = 0.4,
        raster = FALSE,
        label = FALSE,
        label.box = FALSE,
        repel = TRUE) +
    # NoAxes() +
    NoLegend() +
    coord_fixed()

# p1[[1]]$layers[[1]]$aes_params$alpha = .1
plot(p1)

```



<!-- ## Cell Selector -->

<!-- ```{r} -->
<!-- dsx_region_1 <- CellSelector(plot = p1) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- dsx_region_2 <- CellSelector(plot = p1) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- dsx_region_3 <- CellSelector(plot = p1) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- dsx_region_4 <- CellSelector(plot = p1) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- dsx_region_5 <- CellSelector(plot = p1) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- dsx_region_6 <- CellSelector(plot = p1) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- cells_in_dsx_positive_regions <- c( -->
<!--     dsx_region_1, -->
<!--     dsx_region_2, -->
<!--     dsx_region_3, -->
<!--     dsx_region_4, -->
<!--     dsx_region_5, -->
<!--     dsx_region_6 -->
<!-- ) -->
<!-- length(cells_in_dsx_positive_regions) -->
<!-- ``` -->



<!-- ```{r} -->
<!-- # write_csv(x = data.frame(cell_id = dplyr::intersect(dsx_cells,cells_in_dsx_positive_regions)),  -->
<!-- #           file = "../../proj136/analyses/cell_lists/subclustering_dsx_metablob_cells_to_keep--round1.csv" -->
<!-- #         ) -->
<!-- # write_csv(x = data.frame(cell_id = cells_in_dsx_positive_regions),  -->
<!-- #           file = "../../proj136/analyses/cell_lists/subclustering_dsx_metablob_cells_in_dsx_positive_regions--round1.csv" -->
<!-- #         ) -->
<!-- cells_in_dsx_positive_regions <- read_csv(file = "../../proj136/analyses/cell_lists/subclustering_dsx_metablob_cells_in_dsx_positive_regions--round1.csv") %>% pull(cell_id) -->
<!-- ``` -->




<!-- ```{r fig.width=14, fig.height=14} -->

<!-- dsx_cells <- dplyr::intersect(dsx_cells,cells_in_dsx_positive_regions) -->

<!-- p1 <- DimPlot(object = meta_cb, -->
<!-- 			  cells.highlight = dsx_cells, -->
<!-- 			  sizes.highlight = 0.4,  -->
<!-- 			  cols.highlight = "black",  -->
<!-- 			  cols = "lightsteelblue2", -->
<!--         reduction = "tsne_240pcs", -->
<!--         pt.size = 0.4, -->
<!--         raster = FALSE, -->
<!--         label = FALSE, -->
<!--         label.box = FALSE, -->
<!--         repel = TRUE) + -->
<!--     NoAxes() + -->
<!--     NoLegend() + -->
<!--     labs(caption = paste0(length(dsx_cells), " cells")) + -->
<!--     coord_fixed() -->

<!-- # p1[[1]]$layers[[1]]$aes_params$alpha = .1 -->
<!-- plot(p1) -->

<!-- ``` -->



```{r}
KC_sub <- subset(x = KC_seu, cells = KC_cells)
KC_sub
```

```{r}
KC_list <- SplitObject(object = KC_sub, split.by = "experiment")
KC_list
```




```{r}
KC_list_clean <- list()
for (i in seq_along(KC_list)) {
    counts_to_use <- "RNA"
    KC_list_clean[[i]] <- CreateSeuratObject(counts = KC_list[[i]]@assays[[counts_to_use]])
	KC_list_clean[[i]] <- AddMetaData(object = KC_list_clean[[i]], metadata = KC_list[[i]]$orig.ident, col.name = "orig.ident")
	KC_list_clean[[i]] <- AddMetaData(object = KC_list_clean[[i]], metadata = KC_list[[i]]$experiment, col.name = "experiment")
	KC_list_clean[[i]] <- AddMetaData(object = KC_list_clean[[i]], metadata = KC_list[[i]]$tissue, col.name = "tissue")
	KC_list_clean[[i]] <- AddMetaData(object = KC_list_clean[[i]], metadata = KC_list[[i]]$cell_nuclei, col.name = "cell_nuclei")
	KC_list_clean[[i]] <- AddMetaData(object = KC_list_clean[[i]], metadata = KC_list[[i]]$availability, col.name = "availability")
	KC_list_clean[[i]] <- AddMetaData(object = KC_list_clean[[i]], metadata = KC_list[[i]]$sex, col.name = "sex")
	KC_list_clean[[i]] <- AddMetaData(object = KC_list_clean[[i]], metadata = KC_list[[i]]$percent_mt, col.name = "percent_mt")
	KC_list_clean[[i]] <- AddMetaData(object = KC_list_clean[[i]], metadata = KC_list[[i]]$percent_rRNA, col.name = "percent_rRNA")
	KC_list_clean[[i]] <- AddMetaData(object = KC_list_clean[[i]], metadata = KC_list[[i]]$percent_rProt, col.name = "percent_rProt")
	KC_list_clean[[i]] <- AddMetaData(object = KC_list_clean[[i]], metadata = KC_list[[i]]$percent_hsp, col.name = "percent_hsp")
	KC_list_clean[[i]] <- AddMetaData(object = KC_list_clean[[i]], metadata = KC_list[[i]]$modulescore_mt1, col.name = "modulescore_mt1")
	KC_list_clean[[i]] <- AddMetaData(object = KC_list_clean[[i]], metadata = KC_list[[i]]$modulescore_rp1, col.name = "modulescore_rp1")
	KC_list_clean[[i]] <- AddMetaData(object = KC_list_clean[[i]], metadata = KC_list[[i]]$modulescore_hsp1, col.name = "modulescore_hsp1")
}
names(KC_list_clean) <- names(KC_list)
KC_list_clean
```




### Remove Low Quality Cells

<!-- ```{r} -->

<!-- cells_to_keep_file <- "../../proj136/analyses/cell_lists/" -->

<!-- if (file.exists(cells_to_keep_file)) { -->
<!--     cells_to_keep <- read_csv(file = cells_to_keep_file) -->
<!--     for (i in seq_along(KC_list_clean)) { -->
<!--         KC_list_clean[[i]] <-subset(KC_list_clean[[i]], cells = cells_to_keep$cell_id) -->
<!--     } -->
<!--     KC_list_clean -->
<!-- } -->
<!-- ``` -->


### Save list

```{r}
# write_rds(x = KC_list_clean, file = "../../proj136/analyses/rds_files/subclustering_KC__seu_list_clean--round1--500VarGene--NoReg--Harmony_ori_exp_cell.rds")
```




# Normalize


```{r warning=FALSE, message=FALSE}

n_var_features <- 500

for (i in seq_along(KC_list_clean)) {
    KC_list_clean[[i]] <- NormalizeData(object = KC_list_clean[[i]],assay = "RNA")
    KC_list_clean[[i]] <- FindVariableFeatures(object = KC_list_clean[[i]],
                                                  selection.method = "mean.var.plot",     # "mean.var.plot" or "vst"
                                                  nfeatures = n_var_features,
                                                  mean.cutoff = c(0.01,Inf),
                                                  dispersion.cutoff = c(0.01, Inf))
}
```






# Integrate

```{r}
var_features <- SelectIntegrationFeatures(object.list = KC_list_clean, nfeatures = n_var_features)
iso_genes_remove <- c("fru-plus", "fru-dP1", "fru-dP1-plus", "fruA", "fruB", "fruC",
                      "fruD1", "fruD2", "fruABint", "fruBCint", "fruCint", "fruPr1Ex1",
                      "fruPr1Ex2", "fruCOM", "dsxM", "dsxF", "dsxCOM",
                      "EGFP", "GFP", "mCherry", "Gal4", "GAL4",
                      "lncRNA:roX1", "lncRNA:roX2"
                      )
var_features <- setdiff(var_features, iso_genes_remove)
```


```{r}
KC_seu <- merge(KC_list_clean[[1]],
		                unlist(KC_list_clean[2:length(KC_list_clean)]),
		                merge.data = TRUE)

VariableFeatures(KC_seu) <- var_features
KC_seu
```




```{r}
FetchData(object = KC_seu, vars = "experiment") %>% dplyr::count(experiment) %>% arrange(desc(n))
```



# Run PCA



```{r}
library(future)
# plan("multisession", workers = 8)
options(future.globals.maxSize = 8*1024*1024^2)

DefaultAssay(KC_seu) <- "RNA"
KC_seu <- ScaleData(
    object = KC_seu,
    features = rownames(KC_seu),
    assay = "RNA",
    vars.to.regress = c("percent_mt", "percent_hsp", "percent_rRNA", "percent_rProt")
)
```


```{r}
# plan("multisession", workers = 8)
KC_seu <- RunPCA(KC_seu, features = var_features, npcs = 100, approx = FALSE)

plan("sequential")
invisible(gc())
```


# Run Harmony

```{r}
library(harmony)
KC_seu <- RunHarmony(KC_seu,
                    group.by.vars = c("orig.ident", "experiment", "cell_nuclei"),
					lambda = c(1,1,1),
                    assay.use = "RNA")
invisible(gc())
```








# Visualize

## Plot PCA



<!-- ```{r fig.width=8, fig.height=6} -->
<!-- ElbowPlot(KC_seu, reduction = "harmony", ndims = 35) + -->
<!--     geom_vline(xintercept = seq(5,35,5), linetype = "dashed", color = "grey40") + -->
<!--     ylim(0,10) -->
<!-- ``` -->




```{r fig.width=12, fig.height=20, warning=FALSE, message=FALSE}
library(colorspace)
library(cowplot)
p_list <- list()
for (i in 1:10) {
    p_list[[i]] <- DimHeatmap(KC_seu,
                            reduction = "harmony",
                            dims = i,
                            cells = 500,
                            balanced = TRUE,
                            fast = FALSE) +
        scale_fill_gradientn(colors = diverging_hcl(10, palette = "Vik"))
}
plot_grid(plotlist = p_list, ncol = 2)
rm(p_list)
invisible(gc())
```


## Run UMAP



```{r}

# test_pcs <- 60
test_pcs <- seq(5, 35, 5)

for (i in seq_along(test_pcs)) {
    KC_seu <- RunUMAP(KC_seu, 
		                    reduction = "harmony", 
		                    dims = 1:test_pcs[[i]],
		                    n.neighbors = 50,
		                    n.epochs = 1000,
		                    min.dist = 0.01,
		                    spread = 5,
		                    fast = FALSE,
		                    verbose = TRUE,
		                    reduction.name = paste0("umap_", test_pcs[[i]], "pcs"),
		                    reduction.key = paste0("umap", test_pcs[[i]], "pcs_"))
}

KC_seu@reductions[["umap"]] <- KC_seu@reductions[["umap_20pcs"]]
# write_rds(x = KC_seu, file = "../../proj136/analyses/rds_files/subclustering_KC__seu--round1--500VarGene--NoReg--Harmony_ori_exp_cell.rds")
# KC_seu <- read_rds("../../proj136/analyses/rds_files/subclustering_KC__seu--round1--500VarGene--NoReg--Harmony_ori_exp_cell.rds")

```


```{r}

# for (i in seq_along(test_pcs)) {
#     KC_seu <- RunTSNE(KC_seu,
# 		                    reduction = "harmony",
# 		                    num_threads = 24,
# 		                    verbose = F,
# 		                    dims = 1:test_pcs[[i]],
# 		                    do.fast = FALSE,
# 		                    check_duplicates = F,
# 		                    max_iter = 10000,
# 		                    theta = 0.05,
# 		                    reduction.name = paste0("tsne_", test_pcs[[i]], "pcs"),
# 		                    reduction.key = paste0("tsne", test_pcs[[i]], "pcs_"))
# }
# 
# KC_seu@reductions[["tsne"]] <- KC_seu@reductions[["tsne_20pcs"]]
# # write_rds(x = KC_seu, file = "../../proj136/analyses/rds_files/subclustering_KC__seu--round1--500VarGene--NoReg--Harmony_ori_exp_cell.rds")
# # KC_seu <- read_rds("../../proj136/analyses/rds_files/subclustering_KC__seu--round1--500VarGene--NoReg--Harmony_ori_exp_cell.rds")

```



## Plot Embeddings


```{r fig.width=8, fig.height=8}
for (test_pc in test_pcs) {
    plot(DimPlot(object = KC_seu, reduction = paste0("umap_", test_pc, "pcs"), group.by = "experiment") + coord_fixed())
}
```



```{r fig.width=8, fig.height=8}
# for (test_pc in test_pcs) {
#     plot(DimPlot(object = KC_seu, reduction = paste0("tsne_", test_pc, "pcs"), group.by = "experiment") + coord_fixed())
# }
```



```{r fig.width=10, fig.height=10}
col_list <- scales::hue_pal()(length(unique(KC_seu$experiment)))
test_pc <- 20
p <- list()
for (i in 1:length(unique(KC_seu@meta.data[["experiment"]]))) {
  p[[i]] <- DimPlot(object = KC_seu,
                    reduction = paste0("umap_", test_pc, "pcs"),
                    group.by = "experiment",
                    pt.size = 0.4,
                    order = c(unique(KC_seu@meta.data[["experiment"]])[i]),
                    cols = c(rep("lightgrey",length(unique(KC_seu$experiment)) - 1),col_list[[i]])) +
                NoLegend() +
                NoAxes() +
                coord_fixed() +
                ggtitle(unique(KC_seu@meta.data[["experiment"]])[i])
}
plot_grid(plotlist = p, ncol = 3)
```




```{r fig.width=8, fig.height=8}
# col_list <- scales::hue_pal()(length(unique(KC_seu$experiment)))
# test_pc <- 20
# p <- list()
# for (i in 1:length(unique(KC_seu@meta.data[["experiment"]]))) {
#   p[[i]] <- DimPlot(object = KC_seu,
#                     reduction = paste0("tsne_", test_pc, "pcs"),
#                     group.by = "experiment",
#                     pt.size = 0.4,
#                     order = c(unique(KC_seu@meta.data[["experiment"]])[i]),
#                     cols = c(rep("lightgrey",length(unique(KC_seu$experiment)) - 1),col_list[[i]])) +
#                 NoLegend() +
#                 NoAxes() +
#                 coord_fixed() +
#                 ggtitle(unique(KC_seu@meta.data[["experiment"]])[i])
# }
# plot_grid(plotlist = p, ncol = 3)
```


```{r}
FetchData(object = KC_seu, vars = "experiment") %>% 
    dplyr::count(experiment)
```



```{r fig.width=16, fig.height=8}
genes <- c("nFeature_RNA","nCount_RNA")
FeaturePlot(object = KC_seu, 
           features = genes,
           ncol = 2,
           pt.size = 0.4, 
           cols = c("lightsteelblue2", "black"),
           max.cutoff = 10000,
           reduction = "umap_20pcs", 
           raster = FALSE, 
           order = TRUE,
           coord.fixed = TRUE)
```


```{r fig.width=16, fig.height=16}
genes <- c("Imp", "dati", "br", "pdm3")
FeaturePlot(object = KC_seu, 
           features = genes,
           # split.by = "sex",
			# min.cutoff = 4,
           pt.size = 1, 
           cols = c("lightsteelblue2", "black"),
           reduction = "umap_20pcs", 
           raster = FALSE, 
           order = TRUE,
           coord.fixed = TRUE)
```



```{r fig.width=16, fig.height=16}
genes <- c("hec", "Lgr3")
FeaturePlot(object = KC_seu, 
           features = genes,
           # split.by = "sex",
			# min.cutoff = 4,
           pt.size = 1, 
           cols = c("lightsteelblue2", "black"),
           reduction = "umap_20pcs", 
           raster = FALSE, 
           order = TRUE,
           coord.fixed = TRUE)
```






```{r fig.width=6, fig.height=6}

genes <- c("Fer2", "TfAP-2", "sv", "unc-4", "bsh", "inv")
plist <- FeaturePlot(
    object = KC_seu, 
    reduction = "umap_20pcs", 
    features = genes,
    combine = FALSE,
    pt.size = 0.4, 
    min.cutoff = 0,
    max.cutoff = 2,
    # slot = "counts",
    cols = c("lightsteelblue2", "black"), 
    raster = FALSE,
    raster.dpi = c(1024, 1024),
    order = TRUE,
    coord.fixed = TRUE
)

for (i in seq_along(plist)) {
	# plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}

```



```{r fig.width=6, fig.height=6}

genes <- c("Fer2", "Optix", "salr", "salm", "Hr51", "otp")
plist <- FeaturePlot(
    object = KC_seu, 
    reduction = "umap_20pcs", 
    features = genes,
    combine = FALSE,
    pt.size = 0.4, 
    min.cutoff = 0,
    max.cutoff = 2,
    # slot = "counts",
    cols = c("lightsteelblue2", "black"), 
    raster = FALSE,
    raster.dpi = c(1024, 1024),
    order = TRUE,
    coord.fixed = TRUE
)

for (i in seq_along(plist)) {
	# plist[[i]]$layers[[1]]$aes_params$alpha = .1
	plot(plist[[i]])
}

```






# Clustering

```{r}
KC_seu <- FindNeighbors(KC_seu,
                       reduction = "harmony",
                       dims = 1:20,
                       k.param = 30,
                       force.recalc = TRUE,
                       verbose = TRUE)

# in_conda <- R.home() %>% str_detect("R4_cbrg")
# if (in_conda) {
#     algorithm_use <- 4
# } else {
    algorithm_use <- 1
# }

KC_seu <- FindClusters(KC_seu,
                        algorithm = algorithm_use,
                        resolution = c(0.1,0.2,0.5,1,1.5,2,3,4,5,6,8,10),
                        verbose = TRUE)
```



```{r}
write_rds(x = KC_seu, file = "../../proj136/analyses/rds_files/subclustering_KC__seu--round1--500VarGene--NoReg--Harmony_ori_exp_cell.rds")
# KC_seu <- read_rds("../../proj136/analyses/rds_files/subclustering_KC__seu--round1--500VarGene--NoReg--Harmony_ori_exp_cell.rds")
```



```{r fig.width=12, fig.height=26}
p_list <- list()
res <- colnames(KC_seu@meta.data) %>% str_subset("RNA_snn_res") %>% head(n = 10)
for (i in seq_along(res)) {
    p_list[[i]] <- DimPlot(object = KC_seu, group.by = res[[i]],
                           reduction = "umap_20pcs", 
                           label = FALSE, 
                           label.box = TRUE) + 
        NoLegend() + 
        coord_fixed()
}
plot_grid(plotlist = p_list,  ncol = 2)
```


```{r fig.width=12, fig.height=26}
p_list <- list()
res <- colnames(KC_seu@meta.data) %>% str_subset("RNA_snn_res") %>% head(n = 10)
for (i in seq_along(res)) {
    p_list[[i]] <- DimPlot(object = KC_seu, group.by = res[[i]],
                           reduction = "umap_20pcs", 
                           label = TRUE, 
                           label.box = FALSE) + 
        NoLegend() + 
        coord_fixed()
}
plot_grid(plotlist = p_list,  ncol = 2)
```















```{r fig.width=15, fig.height=10}
DimPlot(object = KC_seu, 
		group.by = "RNA_snn_res.4",
		split.by = "sex",
        reduction = "umap_20pcs", 
        label = FALSE, 
        label.box = TRUE) + 
    NoLegend() + 
    coord_fixed()
```




```{r fig.width=12, fig.height=26}
p_list <- list()
res <- colnames(KC_seu@meta.data) %>% str_subset("RNA_snn_res") %>% head(n = 10)
for (i in seq_along(res)) {
    p_list[[i]] <- DimPlot(object = KC_seu, group.by = res[[i]],
                           reduction = "umpa_20pcs", 
                           label = FALSE, 
                           label.box = TRUE) + 
        NoLegend() + 
        coord_fixed()
}
plot_grid(plotlist = p_list,  ncol = 2)
```




```{r fig.width=12, fig.height=26}
p_list <- list()
res <- colnames(KC_seu@meta.data) %>% str_subset("RNA_snn_res") %>% head(n = 10)
for (i in seq_along(res)) {
    p_list[[i]] <- DimPlot(object = KC_seu, group.by = res[[i]],
                           reduction = "umpa_20pcs", 
                           label = TRUE, 
                           label.box = FALSE) + 
        NoLegend() + 
        coord_fixed()
}
plot_grid(plotlist = p_list,  ncol = 2)
```




```{r fig.width=10}
FetchData(object = KC_seu, vars = c("experiment","RNA_snn_res.4")) %>% 
    dplyr::count(experiment, RNA_snn_res.4) %>% 
    ggplot(aes(x = RNA_snn_res.4, y = n, fill = experiment)) +
        geom_bar(stat = "identity", position = "stack")
```



```{r fig.width=10}
FetchData(object = KC_seu, vars = c("experiment","RNA_snn_res.4")) %>% 
    dplyr::count(experiment,RNA_snn_res.4) %>% 
    ggplot(aes(x = RNA_snn_res.4, y = n, fill = experiment)) +
        geom_bar(stat = "identity", position = "fill") #+
        #geom_hline(yintercept = 0.5, colour = "white")
```



```{r fig.width=10}
FetchData(object = KC_seu, vars = c("experiment", "cell_nuclei", "RNA_snn_res.4")) %>% 
    dplyr::count(cell_nuclei, RNA_snn_res.4) %>%
    # mutate(region = if_else(experiment %in% c("aging", "cocaine"),"WB","CB")) %>%  
    ggplot(aes(x = RNA_snn_res.4, y = n, fill = cell_nuclei)) +
        geom_bar(stat = "identity", position = "fill") +
        geom_hline(yintercept = 0.5, colour = "white")
```



```{r fig.width=10}
FetchData(object = KC_seu, vars = c("experiment", "sex", "RNA_snn_res.4")) %>% 
    dplyr::count(sex, RNA_snn_res.4) %>%
    ggplot(aes(x = RNA_snn_res.4, y = n, fill = sex)) +
        geom_bar(stat = "identity", position = "fill") +
        geom_hline(yintercept = 0.5, colour = "white")
```


```{r}
FetchData(object = KC_seu, vars = c("experiment", "sex", "RNA_snn_res.4")) %>% 
    dplyr::count(RNA_snn_res.4, sex) %>% 
    spread("sex", "n")
```





```{r}
FetchData(object = KC_seu, vars = c("experiment", "sex", "RNA_snn_res.4")) %>% 
    dplyr::count(experiment, sex) %>% 
    spread("sex", "n")
```





















# Cell Remove


<!-- ## Round 1 -->





<!-- ```{r fig.width=15, fig.height=10} -->
<!-- DimPlot(object = KC_seu,  -->
<!-- 		group.by = "RNA_snn_res.4", -->
<!-- 		# split.by = "sex", -->
<!--         reduction = "umap_20pcs",  -->
<!--         label = TRUE,  -->
<!--         label.box = FALSE) +  -->
<!--     NoLegend() +  -->
<!--     coord_fixed() -->
<!-- DimPlot(object = KC_seu,  -->
<!-- 		group.by = "RNA_snn_res.5", -->
<!-- 		# split.by = "sex", -->
<!--         reduction = "umap_20pcs",  -->
<!--         label = TRUE,  -->
<!--         label.box = FALSE) +  -->
<!--     NoLegend() +  -->
<!--     coord_fixed() -->
<!-- DimPlot(object = KC_seu,  -->
<!-- 		group.by = "RNA_snn_res.6", -->
<!-- 		# split.by = "sex", -->
<!--         reduction = "umap_20pcs",  -->
<!--         label = TRUE,  -->
<!--         label.box = FALSE) +  -->
<!--     NoLegend() +  -->
<!--     coord_fixed() -->
<!-- DimPlot(object = KC_seu,  -->
<!-- 		group.by = "RNA_snn_res.8", -->
<!-- 		# split.by = "sex", -->
<!--         reduction = "umap_20pcs",  -->
<!--         label = TRUE,  -->
<!--         label.box = FALSE) +  -->
<!--     NoLegend() +  -->
<!--     coord_fixed() -->
<!-- DimPlot(object = KC_seu,  -->
<!-- 		group.by = "RNA_snn_res.10", -->
<!-- 		# split.by = "sex", -->
<!--         reduction = "umap_20pcs",  -->
<!--         label = TRUE,  -->
<!--         label.box = FALSE) +  -->
<!--     NoLegend() +  -->
<!--     coord_fixed() -->
<!-- ``` -->



<!-- ```{r} -->
<!-- cell_remove <- FetchData( -->
<!--         object = KC_seu,  -->
<!--         vars = c("RNA_snn_res.4", "RNA_snn_res.5", "RNA_snn_res.6", -->
<!--                  "RNA_snn_res.8", "RNA_snn_res.10")) %>% -->
<!--     rownames_to_column("cell_id") %>% -->
<!--     filter(RNA_snn_res.4 == 16 | -->
<!--            RNA_snn_res.5 %in% c(16,17,19) | -->
<!--            RNA_snn_res.6 %in% c(19,20,21) | -->
<!--            RNA_snn_res.8 %in% c(21,22,23,24,25) | -->
<!--            RNA_snn_res.10 %in% c(22,24,37,44,43,32) -->
<!--            ) -->
<!-- ``` -->





<!-- ```{r fig.width=6, fig.height=6} -->

<!-- DimPlot(object = KC_seu, -->
<!-- 			  cells.highlight = cell_remove$cell_id, -->
<!-- 			  sizes.highlight = 0.4, -->
<!-- 			  cols.highlight = "black", -->
<!-- 			  cols = "lightsteelblue2", -->
<!--         reduction = "umap_20pcs", -->
<!--         pt.size = 0.4, -->
<!--         raster = FALSE, -->
<!--         label = FALSE, -->
<!--         label.box = FALSE, -->
<!--         repel = TRUE) + -->
<!--     # NoAxes() + -->
<!--     NoLegend() + -->
<!--     coord_fixed() -->


<!-- ``` -->


<!-- ```{r} -->
<!-- cells_to_keep <- FetchData(object = KC_seu, vars = c("RNA_snn_res.1")) %>% -->
<!--     rownames_to_column("cell_id") %>% -->
<!--     filter(!cell_id %in% cell_remove$cell_id) -->
<!-- cells_to_keep -->
<!-- ``` -->



<!-- ```{r} -->
<!-- write_csv(x = cells_to_keep, file = "../../proj136/analyses/cell_lists/subclustering_KC_metablob_cells_to_keep--round1.csv") -->
<!-- ``` -->











# Save


```{r}
getwd()
```


```{r}
write_rds(x = KC_seu, file = "../../proj136/analyses/rds_files/subclustering_KC__seu--round1--500VarGene--NoReg--Harmony_ori_exp_cell.rds")
# KC_seu <- read_rds("../../proj136/analyses/rds_files/subclustering_KC__seu--round1--500VarGene--NoReg--Harmony_ori_exp_cell.rds")
# KC_seu
```








# Markers


```{r}
Idents(KC_seu) <- "RNA_snn_res.2"
markers_pC1 <- FindAllMarkers(object = KC_seu, assay = "RNA", test.use = "wilcox",only.pos = TRUE)
markers_pC1
```


```{r}
write_csv(x = markers_pC1, file = "../../proj136/analyses/markers/subclustering_pC1--round1--500VarGene--NoReg--Harmony_ori_exp_cell--RNA_snn_res_2__cluster_markers.csv")
# markers_pC1 <- read_csv(file = "../../proj136/analyses/markers/subclustering_pC1--round1--500VarGene--NoReg--Harmony_ori_exp_cell--RNA_snn_res_2__cluster_markers.csv")
markers_pC1
```









```{r fig.width=10, fig.height=10}
markers_pC1 %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1) %>%
    slice_head(n = 10) %>%
    ungroup() -> top10
DoHeatmap(KC_seu, features = top10$gene) + 
    scale_fill_continuous_diverging(palette = "Vik", rev = FALSE,
                                    limits = c(-2,2), oob = scales::squish,
                                    na.value = "white") +
    NoLegend()
```









# Session Info

```{r}
sessionInfo()
```




